### 2017-07-12 response/response-jde-1.tex
Initial commit

\documentclass[letterpaper,12pt]{article}

% Xetex preamble
\usepackage{fontspec}
\setromanfont[Ligatures=TeX]{TeX Gyre Pagella}
\usepackage{unicode-math}
\setmathfont{TeX Gyre Pagella Math}

\usepackage[longnamesfirst]{natbib}
%\usepackage[tabhead,nolists,tablesfirst]{endfloat}
\usepackage[flushleft]{threeparttable}
\usepackage{booktabs}
\usepackage{rotating}  
\usepackage{amsmath}
\usepackage{caption} 
\usepackage{dcolumn} 
\usepackage{setspace}
%\usepackage{flafter} 
%\usepackage{longtable}
%\usepackage[pdftex]{graphicx}
\usepackage[xetex,colorlinks=true,linkcolor=black,citecolor=black]{hyperref} 
\usepackage[margin=1.0in]{geometry} 
\usepackage{multirow}

%opening
\title{} \author{}

\doublespacing

\begin{document}

\begin{center} \textbf{\large Income Shocks, Contraceptive Use, and
Timing of Fertility} \end{center}

\begin{center} Response to Referee Comments \end{center}

\noindent Please find attached the revised version of our paper,
``Income Shocks, Contraceptive Use, and Timing of Fertility.''
We would like to thank the two referees for the very useful comments and
suggestions,
which we believe has improved the paper.

[Additional discussion of overall things we have changed---if any]

We list below our responses to the individual comments and suggestions.

\newpage

\section*{Reviewer 1}

% This interesting paper analyzes the relationship between income shocks
% (as proxied by self-reported crop loss) and marital fertility, using a
% small Tanzanian panel survey from the early 1990s. A key feature of the
% paper is that the authors are able to observe use of both traditional
% and modern birth control techniques. In regression specifications with
% and without community or woman fixed effects, the authors find that crop
% loss is associated with decreases in the probability of conceiving, with
% a corresponding increase in use of traditional birth control techniques
% but not modern contraceptives.

The paper could benefit from focusing its motivation somewhat more. An
important point (that the authors do not emphasize enough) is that use
of modern contraception in the sample is extremely low, so the lack of a
relationship between shocks and use of modern methods is not surprising.
The more noteworthy result is that use of traditional methods responds
to shocks. In this sense, the paper makes an important contribution to
understanding how populations control fertility before widespread
adoption of modern contraception. For instance, it is well known that
marital fertility in Europe declined largely before modern contraception
became available, but the methods couples used to achieve this reduction
are not well documented (see, for example, Guinnane JEL 2011). The paper
would be even more compelling if the authors focused the paper's
motivation on the questions of whether and how couples modify fertility
to respond to changing economic conditions before the advent of
modern contraception. That motivation would also help justify their
choice of dataset, which otherwise seems somewhat small and out of date
relative to other available panel datasets from developing countries.

[RESPONSE:]

Specific comments:

\begin{description}

\item The authors interpret their results as postponement, implying they
represent a change in the timing of fertility but not its lifetime
cumulation. That is likely to be the correct interpretation, but can
they really tell? It seems to me that the current fixed effect model
does not tell us. Perhaps they could include 1-2 lags in the crop shock
variable to shed some light on it.

[RESPONSE:]

\item The authors argue that the increase in modern contraceptive use is
intentional, but the evidence for this interpretation is weaker than
they would like. As they do note, while the survey asks about using
abstinence specifically to prevent births, there is much scope for
misreporting, and many women may claim that reductions in coital
frequency were related to fertility planning even if they were not. As
their main evidence against this concern, the authors show that crop
loss has small effects on labor supply. But this finding is only
relevant for the very specific theory that all unintentional abstinence
is related to a work-sex tradeoff. Couples may have less sex during hard
times because they are stressed, for example.

[RESPONSE:]

\item The authors analyze the fertility of women who were married at the
start of the sample period. As a separate analysis, it would be
interesting to know whether marriage hazards change with crop loss in
the parents' home, although perhaps the authors do not have the sample
size to study this question.

[RESPONSE:]

\item The survey measures crop loss continuously, and the indicator for
crop loss greater than 200 TZS is the authors' creation. They show in
the appendix that they obtain similar but less significant results if
they use the continuous measure (in levels or logs) or an indicator for
any crop loss. The 200 TZS seems arbitrary and possibly ex post, since
it leads to the most attractive results of the three sets of results
provided. It would be useful to more transparently show readers the
underlying variation and possible non-linearities, perhaps with a binned
version of the variable.

[RESPONSE:]

\item The authors use causal language throughout the paper, describing
their results as effects. Of course, crop loss may be far from randomly
assigned, so this language is not appropriate.

[RESPONSE:]

\item In the regression specifications, the authors express the survey
wave fixed effects as $D_{i}'\alpha$, where $D_{i}$ is a vector of survey
wave dummies. The subscript is incorrect, and it would be much more
natural to just write $\alpha_{t}$.

[RESPONSE:]

\item What other time-varying covariates are available in the survey? It
would be interesting to consider the extent to which the crop loss
results are explained by other household-level idiosyncratic variation.

[RESPONSE:]

\item The authors report some nonsensical coefficients and standard
errors like -0.000 (0.000). All coefficients and standard errors should
be written to two significant digits.

[RESPONSE:]

\item Table 6 regresses pregnancy on contraceptive use, crop loss, and
their interaction. As the authors note, this regression is quite
difficult to interpret, given the endogeneity of contraceptive use. I
think this table adds little to the paper.

[RESPONSE:]

\item Several appendix tables interact crop loss with initial assets in
Round 1 of the survey. This exercise is interesting and worthy of doing,
given that one might expect asset-poor households to have particular
difficulty in smoothing consumption. But the authors should leave Round
1 outcomes out of this regression.

[RESPONSE:]

\end{description}

% Overall, this is a simple but interesting paper that sheds light on
% important questions in the economics of fertility decisions before
% widespread adoption of modern contraception. It will make an incremental
% but solid contribution to the literature.

\newpage

\section*{Reviewer 2}

% This manuscript explores the effects of "unexpected" crop loss on
% fertility issues. The data come from four survey waves conducted in
% Tanzania, where there are repeated observations for the same individual.
% There are three main outcomes: births, pregnancies, and contraceptive
% use. The main estimation strategy relies on family fixed effects to
% address fixed differences (in wealth, education, etc.) that could be
% related to propensity to experience a crop loss. The crucial identifying
% assumption is that the timing of crop loss for a family is random and
% not related to other factors that might affect fertility.

The manuscript has a big positive: Fertility issues in developing
countries likely play an important role in economic growth, especially
via total completed fertility. Even if the shock in question did not
affect total completed fertility (i.e. individuals only postpone having
children), this paper would still be important since timing of births
matter. For example, being born in a period of low "resources" (food,
income, parental time) could irreparably harm a child's long-term
potential. To my knowledge, we know very little about the main
determinants of birth timing in a developing country context.

I have laid out some comments below in the interest of helping the
authors improve on a promising manuscript.

COMMENTS

\begin{description}

\item Omitted variables

Even with family fixed effects, crop loss is potentially tied with
omitted variables. Crop loss is subjectively measured, so we'd need to
rely on the claim of the survey respondent that the loss was for
"uncontrollable" circumstances. The use of family fixed effects does
address fixed differences in wealth or education that might be related
to families at greater risk of crop loss. But, there could still be time
varying factors. We can't be certain that the crop loss is not the
result of some disease epidemic (or conflict) that also could affect
fertility.

Suggestion: The manuscript should be a little more forthcoming about
this potential problem. I would like to see an exploration into the
potential mechanisms behind the crop loss. Is it weather?

[RESPONSE:]

\item Trend

I am concerned by the fact that crop loss is much higher in the first
and second waves of the survey (see Table 1 on page 6). So, there's a
chance that the estimates are driven by some time trend that affects
family types that are more likely to experience crop losses. (The survey
wave fixed effects won't deal with this if there are differential trends
in the type of families who experience crop loss.)

Suggestion: Control for family characteristics (age, proxies for
education, proxies for wealth) at the first wave interacted with survey
wave. At the very least, this will help acknowledge the identification
threat for the reader.

[RESPONSE:]

\item Spillovers

Neighbors may be affected an individuals crop loss. For one, price of
food goes up. This could be a positive income shock for families with
crops, but bad for families working in other areas.

Suggestion: Aggregate crop loss to the community level and use that as
the treatment variable.

[RESPONSE:]

\item Births analysis

The births analysis automatically drops first wave of survey due to the
longer lag in reported crop loss (Table 3). However, I did not see this
mentioned anywhere.

[RESPONSE:]

\end{description}

\newpage
\bibliographystyle{econometrica}
\bibliography{../paper/incomeShocks-jde-r1}
\addcontentsline{toc}{section}{References}

\end{document}

### 2017-07-12 paper/incomeShocks-jde-r1.tex
Initial commit

% Income Shocks, Contraceptive Use, and Timing of Fertility
% 2016-08-18 - crop loss.tex
% Begun.: 05/01/12
% Edited: 2016-08-18

% Revisions based on Demography referee reports

\documentclass[letterpaper,12pt]{article}

\usepackage[title]{appendix}
\usepackage{fontspec}
\setromanfont[Ligatures=TeX]{TeX Gyre Pagella}
\usepackage{unicode-math}
\setmathfont{TeX Gyre Pagella Math}

\usepackage[margin=1.0in]{geometry}
\usepackage[figuresleft]{rotating}
\usepackage[longnamesfirst]{natbib}
\usepackage{dcolumn}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage[flushleft]{threeparttable}
\usepackage{setspace}
\usepackage[justification=centering]{caption}
\usepackage[font=scriptsize]{subfig}
\usepackage[xetex,colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black]{hyperref}
\usepackage{tikz}
\usetikzlibrary{decorations}
\usepackage{xfrac}

% \bibpunct{(}{)}{;}{a}{}{,}
\newcommand{\mco}[1]{\multicolumn{1}{c}{#1}}
\newcommand{\mct}[1]{\multicolumn{2}{c}{#1}}
\newcommand{\X}{$\times$ }
\newcommand{\hs}{\hspace{15pt}}

% Attempt to squeeze more floats in
\renewcommand\floatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.1}
\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

%------------------------------------------------------------------------

\title{Income Shocks, Contraceptive Use,\\ and Timing of Fertility\thanks{%
We would like to thank 
Alfredo Burlando,
Zo\"e McLaren, 
Erik Plug,
Lawrence Wu, 
and seminar participants at 
Agricultural Economics Society Annual Conference,
European Society for Population Economics Annual Conference,
IUSSP International Population Conference,
Montana State University, 
NEUDC 2015,
Norwegian School of Economics, 
Pacific Conference for Development Economics, 
the seventh annual PopPov Conference, 
Population Reference Bureau,
University of Oregon,
University of Washington,
and the 
Western Economic Association International Conference
for helpful comments and suggestions.
Partial support for this research came from a Eunice Kennedy Shriver National
Institute of Child Health and Human Development research infrastructure grant,
5R24HD042828, to the Center for Studies in Demography and Ecology at the
University of Washington.
}}

\author{Shamma Adeeb Alam\\
	Department of International Studies\\
	Dickinson College\\
	28 N College St\\
	Carlisle, PA 17013\\
	\href{mailto:alams@dickinson.edu}{\texttt{alams@dickinson.edu}}\\
\and
	Claus C P\"ortner\\
    Department of Economics\\
    Albers School of Business and Economics\\
    Seattle University, P.O. Box 222000\\
    Seattle, WA 98122\\
    \href{mailto:cportner@seattleu.edu}{\texttt{cportner@seattleu.edu}}\\
    \href{http://www.clausportner.com}{\texttt{www.clausportner.com}}\\
    \& \\
    Center for Studies in Demography and Ecology \\
    University of Washington\\ 
    }

\date{July 2017}

\begin{document}

% Need this for tables copied directly from estout
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}

\setcounter{page}{0}
\maketitle
\thispagestyle{empty}

\newpage
\setcounter{page}{0}
\thispagestyle{empty}

\doublespacing

\begin{abstract} 
% JDE abstract
\noindent This paper examines the relationship between income 
shocks and fertility decisions. 
Using panel data from Tanzania, we estimate the impact of agricultural shocks 
on 
pregnancies, 
births, 
and 
contraception use.
The likelihood of 
pregnancies and childbirth are significantly lower for households that experience
a crop shock.
Furthermore, women significantly increase their contraception use in response 
to crop losses. 
The increase in contraceptive use comes almost entirely from traditional 
contraceptive methods, such as abstinence, withdrawal, and the rhythm method.
We argue that these changes in behavior are the result of deliberate decisions of
the households rather than the shocks' effects on other factors that influence
fertility, such as women's health status, the absence or migration of a spouse, 
the dissolution of partnerships, or the number of hours worked. 
We also show that, although traditional contraceptives have low overall
efficacy, households with a strong incentive to postpone fertility
are very effective at using them.

\noindent Keywords: Tanzania; family planning; shocks; timing of fertility

\noindent JEL codes: J1, 01, J2
\end{abstract}

\newpage

\section{Introduction}

How shocks affect household decisions on fertility in developing countries has 
received little attention in the literature.
There is evidence that shocks can affect the timing of births, but it is unclear
whether the cause is households' intentional planning or unintended consequences 
of the shocks \citep{pitt98b,lindstrom99,Evans2010,Portner2014}.%
\footnote{
There is also a related literature on the effects of weather, especially temperature,
on timing of fertility \citep[see, for example,][]{Barreca2015}.
%  and TK Josh paper.
In a different vein, a monthlong blackout in Zanzibar increased 
births 8 to 10 months later \citep{Burlando2014}.
There is also a large literature examining how households in developing countries
cope with shocks in areas such as labor supply, crop choice, and education.
} 
Furthermore, standard models of fertility focus mainly on the effects of time
invariant household and individual characteristics on fertility outcomes, and 
tend to ignore the dynamic aspects of fertility decisions.%
\footnote{
For an early exception see \cite{Newman1984}.
}

We seek to answer two key questions in this paper.
First, do income shocks, measured by accidental crop loss in Tanzania, affect timing of 
fertility?
Second, do changes in fertility timing, if any, arise from an intentional decision
to postpone fertility?

There are three reasons why a household may want to delay fertility in the
event of a shock.
First, children are costly in the short run.
Having more children may eventually contribute to the household's production 
and help it overcome shocks, but the short-term impact on availability 
of resources is almost certainly negative \citep{Portner2014}.
Secondly, diverting time away from market work to child care will be even more costly
if households need to respond to a shock by working more \citep{kochar99};
although it is also possible that a shock will temporarily lower women's cost 
of time, making it more attractive to have a birth now.
Finally, the household may realize that children born following a shock 
are more likely to be malnourished and have worse health outcomes, and therefore 
decide to postpone having the next child \citep{Portner2010}.

Observing a decline in pregnancies and birth rates following a shock is, however, 
not evidence of a conscious decision to change the timing of fertility.
A shock may lead one or more members of the household to migrate 
in search of better economic opportunities.
If either the husband or the wife is gone for extended periods, this 
will lower the probability of conception.
Furthermore, a shock may reduce intercourse frequency---either
from psychological effects or from increased hours of work.
We would also observe a decline in births if people delay marriage or if there
is an increase in the dissolution of marriage.
Finally, severe shocks can lead to starvation and health problems \citep{lindstrom99}.
Malnourished pregnant mothers are more likely to have stillbirths,%
\footnote{
See \cite{Hernandez-Julian2014} for a review of the literature and an 
analysis of the effect of famine on stillbirths.
}
and non-pregnant women may see an increased incidence of secondary sterility
from a reduction in age of menopause or famine induced amenorrhea.

Understanding the relation between shocks and fertility decisions is important for
three reasons.
First, as many households in developing countries experience a large number of shocks, 
it is important to understand the factors that help or hinder households' recovery 
after a shock.
Second, children born immediately after a shock fare worse than other children do.
In the short- and medium-term, they have worse health as measured by height-for-age 
Z-scores \citep{Portner2010}, and
long-term effects include worse self-reported health, lower schooling, and less
wealth \citep{Maccini2009}.
% This could be better explained
Finally, parents do not directly control fertility and its timing, but rather 
control type and intensiveness of contraceptive use and the frequency and timing of 
sexual activity.
Hence, it helps us understand if, and how, households regulate their fertility.
This, in turn, can inform policies on availability and targeting
of family planning services.

Most of the prior literature on the impact of economic and other shocks on fertility has
used historic data on what are now developed countries.
Data from Rouen, France, over the period 1681-1787, show that increases in the
price of wheat led to a dramatic fertility decline for the urban poor, 
but fertility of the urban wealthy was unaffected \citep{Galloway1987}.
Similarly, English data from 1542 to 1800 show that mortality shocks led to short-term
fertility declines, with the largest decline typically occurring  the year following 
the shock \citep{Bailey1998}.%
\footnote{
This data also indicated that an increase in real wage led to an increase in short-term 
fertility.
Increases in wage rate and declines in child mortality explain a significant part of the 
long-term fertility decline in Sweden during the period 1736 to 1946 \citep{Eckstein1999a}.
In another study on Sweden, increases in the value of women's time relative to men's 
time---arising from increases in the price of butter---led to a decline in fertility during 
the  1860-1910 period \citep{Schultz1985}. 
}

In pre-industrial Sweden, a positive crop shock, positive weather shock, and positive wage 
shock increased fertility for a five- to ten-year period following the shocks \citep{Eckstein1985}.
However, the increase in birth rate changes only the timing of births and has no cumulative 
effect on long-term fertility rate.
Similarly, the increasing price of rye in Southern Sweden led landless and semi-landless 
families to reduce their fertility, whereas landed farmers showed no effect 
\citep{bengtsson06}.

Using present-day data, there is evidence that famine and domestic/regional military
attacks in Ethiopia led to short-term decreases in the likelihood of conception
\citep{lindstrom99}.
Closer to our paper, positive rainfall shocks in Senegal increased 
fertility \citep{pitt98b}.
None of these paper can, however, directly address whether the observed changes 
arise from deliberate control of fertility.

Many studies examine how a variety of factors affect contraception use, such 
as, 
schooling, 
focused information campaigns, and
participation in savings or credit group 
\citep{Ainsworth1996,Feyisetan1996,Steele2001,Chen2003}.
To our knowledge, only one study attempts to show the effect of shocks on 
contraception use \citep{Hernandez-Correa2010}. 
Cross-sectional data from Madagascar show that households that experience adverse events 
are more likely to use contraception.\footnote{%
Adverse events include economic and environmental aspect, such as: rise in input 
cost, rise in cost of goods, difficulty finding buyers, difficulty finding inputs, 
floods, late rains, early rains, droughts, pest problems, etc. 
} 
However, households may be able to anticipate many of these events before they 
occur and adjust their behavior accordingly.
Combined with possible correlation between the probability of
experiencing an adverse event and unobserved individual and household
characteristics, this may bias the estimates, and the cross-sectional
data allows for neither household nor individual level fixed effects
estimates.

We use longitudinal data from the Kagera region in Tanzania with information on accidental 
crop loss, fertility, and contraceptive use to estimate individual level fixed 
effects models of the effects of income shocks on both timing of fertility and 
contraceptive use.
We show that income shocks decrease the likelihood of pregnancies and births.
At the same time, shocks significantly increase use of contraceptives, 
especially traditional contraceptives.
We argue that the postponement of fertility in households experiencing income 
shocks is the result of the increased contraception use rather than a biological 
response or separation of spouses.

The paper makes two main contributions to the literature. 
First, this is the first paper to show that households respond to income shocks 
through family planning, which, in turn, affects the timing of child bearing. 
Our study provides evidence that the postponement of fertility occurs through a 
planned decision process rather than as an unplanned consequence.
Secondly, the paper contributes to the consumption smoothing literature as households 
affected by income shock use family planning as a mechanism to smooth their consumption. 
In the short run, bearing a child means diverting scarce resources away from other
useful purposes to the birth and maintenance of the child. 
Farmers can, therefore, partly smooth their consumption by delaying births during 
times of income shock.

\section{Data and Estimation Strategy}

We use the Kagera Health and Development Survey (KHDS) conducted by the World Bank 
and the University of Dar es Salaam in the Kagera region in Tanzania. 
The survey covers four rounds from 1991 through 1994 and over 800 households, 
drawn from 51 communities (49 villages) in the six districts of Kagera.%
\footnote{
The sample selection was based on a variable probability sampling procedure 
(a two-stage, randomized stratified procedure) based on expected mortality.
In the first stage, based on the 1988 Tanzanian census, the census clusters were randomly 
selected after stratifying them according to mortality rates and agro-climatic zones. 
Households were then divided into ``high-risk'' and ``low-risk'' groups in the 
second stage, based on illness and death of households in the 12 months 
before the enumeration process.
Finally, households were randomly sampled from the groups.
For further details on the sample selection, please see World Bank (2004).
}
The average interval between each of the survey round was between six and seven months, 
and interviews of households took place on a rolling basis throughout the year. 
The Kagera region is in the northwestern corner of Tanzania, west of Lake Victoria, and 
borders Uganda, Rwanda, and Burundi as shown in Figure \ref{fig:kagera_map}. 
The population of the area---1.3 million in 1988---is 
overwhelmingly rural and primarily engaged in agriculture.

\begin{figure}[ht] 
 \centering
 \includegraphics[width=0.6\textwidth]{../map/kagera.pdf}
 \caption{Map of Kagera region}
  \label{fig:kagera_map}
\end{figure}

The data contain information on individual and household level demographic 
and socio-economic characteristics. 
In addition, questions about fertility and birth control were asked of all 
married women, regardless of age, and unmarried women 14 years and older. 
We restrict our sample to married or partnered women 18-45 years of age who we 
observe in all 4 rounds of the survey.
This ensures that we capture women who are at risk of conception and for whom 
the available observations cover sufficient time for shocks to have an effect
on pregnancies and births.
The downside is that the sample size is reduced.
We discuss possible non-random attrition below.

\begin{table}
\centering
\footnotesize
\begin{threeparttable}
\caption{Descriptive statistics for Crop loss and Outcomes}
\label{tab:desc_stat_croploss}
\begin{tabular}{l  D{.}{.}{2,3} D{.}{.}{2,3} D{.}{.}{2,3} D{.}{.}{2,3} D{.}{.}{2,3}} \toprule
                                  & \multicolumn{4}{c}{Wave} & \mco{Average} \\ \cmidrule(lr){2-5}
                                  &\multicolumn{1}{c}{1} &\multicolumn{1}{c}{2}   &\multicolumn{1}{c}{3} &\multicolumn{1}{c}{4} &   \\ \midrule
Dummy crop loss (1-7 months) $\geq$ TZS 200            &        0.71&        0.26&        0.02&        0.03&        0.25\\
                                                       &      (0.45)&      (0.44)&      (0.13)&      (0.18)&      (0.44)\\ 
\addlinespace 
Currently pregnant                                     &        0.17&        0.23&        0.15&        0.14&        0.17\\
                                                       &      (0.37)&      (0.42)&      (0.36)&      (0.35)&      (0.38)\\
Gave birth since last survey                           &        0.19&        0.16&        0.22&        0.12&        0.17\\
                                                       &      (0.40)&      (0.37)&      (0.41)&      (0.32)&      (0.38)\\
\addlinespace 
Contraceptive use                                      &        0.15&        0.09&        0.09&        0.10&        0.11\\
                                                       &      (0.36)&      (0.29)&      (0.29)&      (0.30)&      (0.31)\\
Contraceptive use -- Traditional\tnote{a}              &        0.11&        0.06&        0.07&        0.04&        0.07\\
                                                       &      (0.32)&      (0.24)&      (0.25)&      (0.21)&      (0.26)\\
Contraceptive use -- Modern\tnote{b}                   &        0.04&        0.02&        0.02&        0.05&        0.03\\
                                                       &      (0.20)&      (0.14)&      (0.14)&      (0.22)&      (0.18)\\ 
\bottomrule
\end{tabular}
\begin{tablenotes} 
\scriptsize
\item[a] Traditional contraceptives include abstinence and rhythm method.
\item[b] Modern contraceptives include condom, diaphragm, pill, IUD, injection, female and 
male sterilization.
\end{tablenotes}
\end{threeparttable}
\end{table}

Table \ref{tab:desc_stat_croploss} shows the main explanatory variable of interest 
and the 5 outcomes that we examine.
The main explanatory variable is a dummy for whether the household experienced 
an accidental loss of crops, which covers any crop lost due to insects, 
rodents, fire, rotting, or other calamities, greater than TZS 200 since the 
prior survey round (or prior 1 year for the first round).
% We create a dummy for accidental crop losses greater than 200 TZS,
% which is our preferred measure of income shock.
The TZS 200 cut-off is equivalent to approximately 5th 
percentile of all household savings.%
% That period is approximately 6 to 7 months for the last three surveys, but covers 
% the previous 12 months for the first survey. 
% [MAY HAVE TO EXPLAIN MORE ABOUT WHAT THIS MEANS FOR OUR ESTIMATES]
There is substantial variation in the number of households that experience a 
crop loss between the 4 survey rounds.
In the 12 months prior to the first survey round just over 70\% of households 
lost crops, whereas just over 25\% lost crops between the 1st and 2nd survey round.
That falls even further for the 3rd and 4th rounds, with less than 5\% 
reporting crop loss of more than TZS 200 for both.
Our results are robust to defining crop loss as any crop loss, with the 
coefficients slightly smaller and slightly less statistically 
significant as expected. 
% Crop losses between the valuation of 0-200 TZS are not coded as an income shock,
% because of the small magnitude (equal to  of savings). 
% and unlikely to affect fertility decisions of households. 
Furthermore, our results are also qualitatively similar if we use a log of the amount of 
crop lost or a linear specification of crop loss instead of the dummy variable.
\footnote{
Results using these alternative specifications of crop loss are available in the Appendix.
}

On average across the four survey rounds, 17\% of women report being pregnant 
at the time of the interviews and 17\% of women report giving birth since the 
prior survey round.
In total, just below 70\% of women in the sample report giving birth in
the covered period (3\sfrac{1}{2} years).
Similarly, 70\% report that they are currently pregnant at one point during 
the 4 rounds.

We capture contraceptive use at the time of the survey, and further divide
contraceptives into traditional and modern types. 
Traditional contraceptives include abstinence, rhythm, and withdrawal, 
and modern contraceptives include condom, diaphragm, pill, IUD, injection,
and female and male sterilization.%
\footnote{
The relevant questions are "Some couples use contraception methods to avoid
pregnancy or to space births. Are you currently using a method of contraception?".
If the woman answers yes, she is then asked "What contraceptive method are you
and your partner using at present?"
}
A small subset of women report using both modern and traditional contraceptives;
these women are included in both categories.
If a woman is currently pregnant, we code her as not using contraceptives.
We remove sterilized women from the sample, as sterilization is typically a 
permanent procedure.

Contraceptive prevalence is low; at the time of the first survey round, 15\% 
of women report using any type of contraceptive---split between 11 percent that use 
traditional contraceptives and 4\% that use modern contraceptives.
In subsequent survey rounds, the contraceptive use rate hovers around 10\%, 
with 4 to 7\% using traditional contraceptives and 2 to 5\% 
using modern contraceptives.

% [Descriptive statistics and discussion here]

\begin{table}
\centering
\footnotesize
\begin{threeparttable}
\caption{Wave 1 Descriptive Statistics for Women}
\label{tab:desc_stat_women}
\begin{tabular}{l  D{.}{.}{3,2} D{.}{.}{3,2} D{.}{.}{3,2} D{.}{.}{3,2}} \toprule
                            	 & \mco{Mean} 	 & 	\mco{St Dev}	  \\ \midrule
Age 18-22                                              &        0.23&        0.42\\
Age 23-27                                              &        0.17&        0.38\\
Age 28-32                                              &        0.23&        0.42\\
Age 33-37                                              &        0.21&        0.41\\
Age 38-45                                              &        0.16&        0.37\\ \addlinespace 
No education                                           &        0.29&        0.45\\
1 - 6 years of education                               &        0.26&        0.44\\
7 plus years of education                              &        0.45&        0.50\\ \addlinespace 
Assets per capita in wave 1 (10,000 TZS)\tnote{a}      &        7.67&       12.75\\ \addlinespace 
Number of women                                        &     \mct{247}            \\
\bottomrule
\end{tabular}
\begin{tablenotes} 
\scriptsize
\item[a] Assets capture self-reported values of land, livestock, business assets, durable
goods, and savings.
\end{tablenotes}
\end{threeparttable}
\end{table}

Table \ref{tab:desc_stat_women} shows descriptive statistics for basic
demographic and household characteristics at wave 1.
% This survey suffers from two data quality issues: misreporting of 
% age and misreporting of education level.
As the time span of the surveys is too short to confidently identify age effects,
we use women's age at the time of the first survey round.
There is substantial measurement error in age and heaping of reported ages 
around numbers ending in 0 and 5.%
\footnote{
Some women report the same age in all survey rounds despite that there are
21 months between the first and last survey round, whereas other women's reported 
age goes up by 1 year between each survey round despite that the surveys are only 
6 to 7 months apart.
}
% If her reported age vary more than the length of interval between each survey 
% round, we use her median reported age rounded down.
We use respondent's median reported age, rounded down, and 5-year age groups 
centered around ages ending in 0 and 5 to take account of heaping.
Only married/partnered women are in the sample, which explains the 
equal age distribution despite population growth in the area.

A notable characteristic is the high level of reported education for a poor 
rural area, although there also appears to be errors in 
reported education levels.%
\footnote{
For many women, the reported education level actually decreases in certain survey 
rounds, or increases by 2 years or more in consecutive survey rounds, which 
conflicts with the 6 to 7 months survey intervals. 
Additionally, individuals are supposed to complete primary school by the age of 15,
so changes in primary schooling levels in our sample is likely from reporting 
errors. 
For women who report different primary school education levels, we 
assign the modal education level if possible, or, if there is no modal 
observation, the median years of education.
We use the same approach for household heads' education
when we use household heads' education as a proxy for wealth.
}
We divide education levels into three groups: no education, 
some primary (1-6 years of education), and primary school 
or above (7 or more years of education).
Only 29\% of the women received no education, 26\% received some 
primary education, and 45\% completed at least primary education.
The high reported education level is most likely the result of the 1974 
Universal Primary Education Movement,
which increased accessibility of primary education and enrollment rates, 
even though there are reports that the quality of education was very low.
In addition, the crisis Tanzania experienced in the 1980s further lowered the 
quality and enrollments declined significantly.%
\footnote{See \cite{Galabawa2001} and \cite{Wedgwood2005} for detailed discussions 
of the development of education in Tanzania.
}
Hence, it is unclear to what extent reported education levels reflect 
women's actual human capital.

To capture wealth status of the household, we use total reported assets of the 
household.
These include self-reported values of land, livestock, business assets, durable
goods, and savings.
We measure assets per capita in 10,000 Tanzanian Shilling (TZS).

Finally, the average level of assets in wave 1 was just shy of TZS 80,000.%
\footnote{%
In 1991 the official exchange rate was TZS 200 to the dollar, while it
was TZS 440 to the dollar in the parallel market.
}
There was, however, wide dispersion in asset levels, with the minimum reported 
wave 1 asset level around TZS 510 and the maximum almost TZS 19,000,000.
The 25-percentile asset level is around TZS 19,000, median is approximately 
TZS 35,000, and the 75-percentile asset level is only slightly higher than the 
mean at TZS 86,000.
In other words, the asset levels for a majority of households in the area is very low,
with a few relatively wealthy households.

A potential issue is that some women may be left out of our sample because of
non-random attrition.
In Appendix Tables \ref{tab:attrition_desc_stat_croploss} and \ref{tab:attrition_desc_stat_women}
we show descriptive statistics for women who fulfilled the requirements for being in the 
sample in wave 1, but were not included in the main sample.
A total of 207 women were dropped because they were not observed in all four waves 
and/or failed the criteria in at least a one of the later waves, for example 
because they divorced or became widows.%
\footnote{
Divorce, separation, and widowhood accounted for 28 women, whereas not observed in all
4 waves for accounted for 188 women. 
The two numbers do not add to 207 because some women experienced both a change
in marital status and were not observed in all 4 waves.
}
In each period, women who were dropped from the sample were less likely to have been 
affected by a shock.
Furthermore, they were less likely to be pregnant and to have given birth, and,
correspondingly, a larger percentage use contraceptives.
Compared to our main sample the women dropped were more likely to be at the ends of
the included age samples, with 6 percentage points more women in the very youngest
and the very oldest age groups.
They were also more educated and had a slightly higher level of asset per capita.
Hence, it appears that we are not dropping women who were more likely to 
experience shocks and less likely to cope with them, which in turn suggests that 
non-random attrition is not a substantial problem.
However, to further investigate whether there is non-random attrition, we 
estimate the association between shocks and partner migration and between
shocks and dissolution of marriage as described below.

\subsection{Estimation Strategy}

An important concern when estimating
the effects of shocks on fertility and contraception decisions 
is that there may be unobserved factors that affect both likelihood of 
shocks and outcomes.
A household with lower (unobserved) ability may, for example, simultaneously be more 
likely to experience a shock and less likely to afford or have knowledge about 
contraceptives.
This would bias OLS estimates of shocks' effects towards zero.
To address the problem of unobserved heterogeneity and endogeneity, our
preferred model is an individual level fixed effect model. 
This approach allows us to control for all time-invariant mother characteristics when 
estimating the impact of crop loss on contraceptive use and fertility outcomes.
The main downside is that fixed effects tend to exacerbate measurement errors.

We begin by estimating OLS and community and individual level fixed effects 
versions of the following equation:
\begin{align}
Y_{i,t} &=  \beta_1 S_{i,t}  +\boldsymbol{D_{i}}'\boldsymbol{\alpha} 
+ \mu_i + \varepsilon_{i,t},   \label{eq:fixed_effects} 
\end{align}
where $Y_{i,t}$ is the outcome of interest,  
$S_{i,t}$ is whether the household experienced a shock in the relevant period,
$\boldsymbol{D_{i}}$ is a vector of survey wave dummies, 
and $\mu_i$ is time-invariant individual specific characteristics.
With the outcomes of interest all dichotomous, the estimated equations 
are linear probability models. 
For the outcome variables, currently pregnant and contraceptive use, we use 
the dummy variable for crop loss that occurred in the period immediately 
prior to the interview, which is 6 to 7 months long, as our main explanatory 
variable of interest. 
When we use birth as the outcome variable, 
we use the dummy variable for crop loss that occurred in the period between 
the prior two survey rounds.%
\footnote{
In the tables we indicate this difference by referring to the period 
immediately before the survey as 1-7 months and the period between the
prior survey and the survey before that as 7-14 months.
}

To further investigate whether the response to shock vary by household
characteristics, we also estimate individual level fixed effects versions 
of the following equation:
\begin{align}
Y_{i,t} &=  \beta_1 S_{i,t}  +  \beta_2 S_{i,t} \times X_{i,1} +
\boldsymbol{D_{i}}'\boldsymbol{\alpha} + \mu_i + \varepsilon_{i,t},   
\label{eq:fixed_effects} 
\end{align}
where $X_{i,1}$ are either assets holdings at the time of the first
survey or education levels of the respondent or household head.

\section{Results}

% [BASELINE estimations]

First, we address whether agricultural shocks affect timing of fertility at all.
Table \ref{tab:birth} shows OLS, community fixed effects, and woman level individual 
fixed effects estimates of the effect of crop loss on the likelihood of pregnancy 
and childbirth.
The three left columns show the effects of crop loss in the 7 months prior to the
survey on a woman's likelihood of being pregnant at the time of the interview.
The three right columns show the effects of crop loss on the likelihood of having given 
birth since the last survey round. 
For a birth to have occurred in the 7 months prior to the interview, a woman would have
had to conceive in the period before the prior survey round, and we therefore use shocks
that occurred 7 to 14 months prior to current round.

% health effect explain the difference between births and pregnancy.
% We would also expect pregnancy to be a little lower because some women may not
% realize that they are pregnant. 
% 95\% pregnancy: -.1774091   -.0084771
% 95\% birth: -.2570566   -.0607087

\begin{table}[htbp]
\begin{center}
\begin{threeparttable}
\caption{The Effects of Crop Loss on Pregnancy and Births}
\label{tab:birth}
\begin{tabular}{@{} l D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5}  @{}}
\toprule
                                                       &\multicolumn{3}{c}{Currently pregnant}&\multicolumn{3}{c}{Birth since last survey}\\ \cmidrule(lr){2-4} \cmidrule(lr){5-7}
                                                       &                     & \mct{Fixed Effects}                       &                     & \mct{Fixed Effects}                       \\ \cmidrule(lr){3-4} \cmidrule(lr){6-7}
                                                       & \mco{OLS}           & \mco{Community}     & \mco{Woman}         & \mco{OLS}           & \mco{Community}     & \mco{Woman}         \\ \midrule
Crop loss - 1-7 months                                 &      -0.053         &      -0.074\sym{*}  &      -0.093\sym{**} &                     &                     &                     \\
\hs (200 TZS or above)                                 &     (0.038)         &     (0.039)         &     (0.043)         &                     &                     &                     \\
Crop loss - 7-14 months                                &                     &                     &                     &      -0.103\sym{***}&      -0.115\sym{***}&      -0.159\sym{***}\\
\hs (200 TZS or above)                                 &                     &                     &                     &     (0.040)         &     (0.042)         &     (0.050)         \\
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Community fixed effects                                &    \mco{No}         &    \mco{Yes}        &    \mco{No}         &    \mco{No}         &    \mco{Yes}        &    \mco{No}         \\
Woman fixed effects                                    &    \mco{No}         &    \mco{No}         &    \mco{Yes}        &    \mco{No}         &    \mco{No}         &    \mco{Yes}        \\
Observations                                           &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{741}        &    \mco{741}        &    \mco{741}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule    
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
All models are linear probability models.
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
% Assets are per capita and measured in 10,000 TZS.
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

Two main results stand out.
First, there are statistically significant and negative effects of crop loss on the
likelihood of both pregnancies and births.
Our preferred estimates---woman level fixed effects---show that experiencing 
crop loss decreases the likelihood of being pregnant by 9 percentage points 
and the likelihood of a birth by 16 percentage points.
These effects are substantial; on average across the four survey rounds, approximately 
17\% of women were pregnant and 17\% had given birth in the preceding period. 
Second, the absolute values of the community fixed effects point estimates are 
larger than the OLS estimates for both outcomes and the individual level
fixed effects estimates are larger than the community fixed effects estimates.%
\footnote{
Full results for the OLS and community fixed effects models are in shown in 
Table \ref{tab:ols_fertility}.
Table \ref{tab:com_cluster_birth} shows the results with clustering at community level
to account for potential spacial correlation in shocks.
Clustering at community level leads to slightly less statistically significance for 
pregnancies and slightly higher statistically significance for births.
}
This suggests that unobservable characteristics, at both community
and individual level, correlate with a higher likelihood of experiencing 
a shock and lower likelihood of using contraceptives.

\begin{table}[htbp]
\begin{center}
\begin{threeparttable}
\caption{The Effects of Crop Loss on Contraceptive\\ Use Overall and By Type}
\label{tab:contraceptive}
\begin{tabular}{@{} l D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
\toprule
                                                       &                     & \mct{Fixed Effects}                       \\ \cmidrule(lr){3-4} 
                                                       & \mco{OLS}           & \mco{Community}     & \mco{Woman}         \\ \midrule
                                                       &\multicolumn{3}{c}{Any contraception}  \\ \cmidrule(lr){2-4}
Crop loss - 1-7 months (200 TZS or above)              &       0.061\sym{**} &       0.059\sym{**} &       0.070\sym{**} \\
                                                       &     (0.027)         &     (0.027)         &     (0.033)         \\
\addlinespace 
                                                       &\multicolumn{3}{c}{Traditional contraception}  \\ \cmidrule(lr){2-4}
Crop loss - 1-7 months (200 TZS or above)              &       0.052\sym{**} &       0.048\sym{**} &       0.054\sym{**} \\
                                                       &     (0.023)         &     (0.022)         &     (0.024)         \\
\addlinespace 
                                                       &\multicolumn{3}{c}{Modern contraception}  \\ \cmidrule(lr){2-4}
Crop loss - 1-7 months (200 TZS or above)              &      -0.001         &       0.004         &       0.003         \\
                                                       &     (0.014)         &     (0.013)         &     (0.020)         \\
\addlinespace                                                        
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Community fixed effects                                &    \mco{No}         &    \mco{Yes}        &    \mco{No}         \\
Woman fixed effects                                    &    \mco{No}         &    \mco{No}         &    \mco{Yes}        \\
Observations                                           &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
All models are linear probability models.
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
% Assets are per capita and measured in 10,000 TZS.
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

Our results in Table \ref{tab:birth} are consistent with the prior literature,
with shocks associated with statistically significant declines in pregnancies 
and births.
But, households do not control these two outcomes directly.
Instead, they have control over some of the proximate determinants of fertility, 
such as use of contraceptives and coital frequency.
Our second question is to what extent households are intentionally delaying 
the conception and birth of their next child. 
To answer this question, we examine the effects of crop loss in the 7 months
prior to the survey on contraceptive use at the time of the interview. 
Table \ref{tab:contraceptive} presents OLS, community fixed effects, and woman 
level individual fixed effects estimates of  effect of crop loss on overall 
contraceptive use and use by type.%
\footnote{
Full results for the OLS and community fixed effects models are in shown in 
Table \ref{tab:ols_contraceptive}.
Table \ref{tab:com_cluster_contraceptive} show results with clustering at
community level, which leads to little change in the level of
statistically significance.
}

The main result is that crop loss leads to a statistically significant 
increase in the overall contraceptive use.
Our preferred model shows that crop loss is associated with
a 7 percentage points increase in contraceptive use.%
\footnote{
The difference between the OLS, community, and individual level fixed effects
estimates are consistent, but less clear-cut, with what we found for pregnancies 
and births.
The woman fixed effects estimate is again the largest.
}
The effect on overall contraceptive use is substantial considering that the 
average use rates across the four rounds vary between 9 and 15\%.
The statistically significant increase in contraceptive use following
a crop loss indicates that households are reacting to crop 
losses by increasing contraceptive use.

Splitting contraceptive use into traditional and modern shows that almost
all of the increase in contraceptive use comes from traditional contraceptives.
The individual fixed effects specification shows that a
crop loss leads to a statistically significant increase in
traditional contraceptive use by 5.4 percentage points, whereas
the effect on use of modern contraception is close to zero
and not statistically significant. 
Given the low levels of use for modern contraceptives in the area
this result is not surprising.

In sum, crop loss is strongly associated with a decline in
both the likelihood of being currently pregnant and the likelihood
of having given birth in the period before the survey round.
Furthermore, women who experienced a crop loss are substantially
more likely to use contraceptives, especially traditional methods.
These results are consistent with the idea that households 
deliberately postpone births in response to adverse shocks.%
\footnote{
To examine whether the effects of crop loss are an artifact of our
chosen specifications we also tried alternative specifications
as shown in Appendix Tables \ref{tab:logit}, \ref{tab:ln_croploss}, 
and \ref{tab:continuous}.
The results are qualitatively the same, especially for the Logit model
and for the log of crop loss.
Furthermore, Appendix Table \ref{tab:agegroups} shows the effects
of crop losses by age groups.
The effect of crop losses are concentrated for women younger than 32
years of age, although as discussed above there appears to be 
substantial measurement error in self-reported age.
}

\subsection{Conscious Decision or Unintended Consequence?}

Our main results point to households making a conscious decision
to postpone fertility in the face of a shock.
Three questions arise from our results.
First, 
why are there differences in the point estimates for
the effects of crop across the three main outcomes: 
births, pregnancy, and contraceptive use?
Second, 
are we picking up behavior changes brought on by
crop loss, such as decrease in coital frequency from substantial
increase in hours worked or poorer health, that is later 
rationalized or reported as increased use of traditional 
contraception methods such as abstinence?
Third, 
how can the effect on postponement of fertility be so
strong if traditional contraceptives have low efficacy?
We attempt to answer these questions in the remainder of the paper.

The first question is why the point estimates for the three main 
outcomes differ.
If deliberate timing of fertility was the only explanation, 
we should observe similar absolute effects of crop loss across all three outcomes.
But, the point estimate is largest for births, second largest for pregnancy, 
and smallest for contraceptive use.
Part of the explanation for this could be simple random variation.
Indeed, we cannot reject that the absolute values of the point 
estimates for births and contraceptive use and pregnancy and contraceptive use are
the same, although we can reject at the 5\% significance level that the
point estimates for births and pregnancy are the same.%
\footnote{
The tests were done using Stata's ``suest'' command, which allows for correlation 
of errors across the three main fixed effects models,  with clustering at
household level. 
}

The pattern of largest point estimate for births, followed by
pregnancy and contraceptive use also matches what we would 
expect to arise from measurement error.
As mentioned above, fixed effect models exacerbate measurement error, 
which bias the estimated coefficients toward zero.
Of the three outcomes, births should have the least amount of 
measurement errors.
Women know whether they have given birth, 
although it is possible that some women report a birth that occurred 
before the prior survey and others fail to report births that 
occurred between the prior survey and the current one.
Pregnancy is associated with more measurement error;
some women may believe themselves to be pregnant when they are
not and some women may not realize that they are pregnant.
Contraceptive use likely has the most measurement 
error---especially in relation to how we measure births and pregnancies.
The surveys asked only about current use, so some women may
have used contraceptives intensively since last survey and 
just recently stopped, or only recently begun.
Obviously, we cannot verify that measurement error is the explanation.
It is, however, consistent with crop loss having the same underlying impact on 
all outcomes, with measurement error biasing the estimated effects most 
for contraceptives and least for births.
What also points in this direction is that the OLS point estimates 
are even less statistically significantly different from each other
than the fixed effects point estimates.%
\footnote{
Of course, the problem with the OLS estimates is that they
do not adequately control for endogeneity, although we see
no strong reason why this should be related to differences
in the degree of measurement error across the three outcomes.
}

Finally, it is possible that the differences in point estimates 
could reflect effects of crop loss other than conscious 
delay of fertility.
There are two main possible culprits: a decline in health
or the absence of the husband or the wife.
Worsening health has two effects.
First, more women will be unable to conceive, which would make
the effect of crop loss on pregnancy larger than the effect
on contraceptive use.
Second, those women who do conceive will be less likely to carry their 
pregnancy to term, which would make the effect on births 
larger than the effect on pregnancy.
Similarly, the absence of a partner will lower the risk of
pregnancy and birth for a given level of contraceptive use.
Table \ref{tab:health} shows the impacts of crop loss on body mass index 
(BMI) and self-reported illness in the last 4 weeks for the women, 
outmigration of partner, and risk of marriage dissolution.

\begin{table}[htbp]
\begin{center}
\begin{threeparttable}
\caption{The Effects of Crop Loss on Women's Health, Absence of Partner, and Marriage Dissolution}
\label{tab:health}
\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
\toprule
                                                       & \mct{Respondent} \\
													   & \mco{BMI}           &  \mco{Illness}  	       \\ \midrule
Crop loss - 1-7 months (200 TZS or above)              &      -0.014         &       0.023         \\
                                                       &     (0.016)         &     (0.047)         \\
Pregnant                                               &       0.083\sym{***}&                     \\
                                                       &     (0.013)         &                     \\
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        \\
Observations                                           &    \mco{988}        &    \mco{988}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        \\ 
\addlinespace 
                                                       &\mco{Absence/migration}&\mco{Dissolution of}\\
                                                       &\mco{of partner}     &\mco{ of marriage}\\ \midrule
Crop loss - 1-7 months (200 TZS or above)              &       0.011         &      -0.001         \\
                                                       &     (0.011)         &     (0.012)         \\
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}       \\
Observations                                           &    \mco{988}        &  \mco{1,293}        \\
Number of women                                        &    \mco{247}        &    \mco{578}        \\
\addlinespace 
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
The sample for the dissolution of marriage includes women who are dropped
from the main sample because they do not show up in all four survey rounds
or who are dropped before they are not partnered in all four survey rounds.
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
All values of crop lost and assets are per capita and measure in 1,000 TZS.
Variables not shown: Age of mother and number of births.
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

Both health outcomes have the expected signs, but the effects are small
and not statistically significant.
Similarly, Table \ref{tab:health} shows that shocks do not lead to a
significant increase in migration or a significant increase in dissolution of
marriages.
The coefficient for migration has the expected positive sign, but the
effect is too small to explain the reduction in pregnancies and births. 
In contrast, the coefficient for dissolution of marriage has the
opposite sign, and the magnitude is near zero. 
Hence, neither health nor the absence of a partner seem to drive the
postponement of fertility, suggesting that the effect of crop loss 
works predominately through the increased use of contraceptives.

% - hours worked and abstinence  / also shows that the shocks are "real"
The second question is whether the amount worked increases so substantially 
that coital frequency declines and the respondent reports this as abstinence.%
\footnote{
We are grateful to Lawrence Wu for suggesting this mechanism.
}
We examine this possibility in two ways.
First, for this explanation to hold the amount worked must increase 
dramatically.%
\footnote{
As we mentioned in the introduction it is possible that crop loss could
decrease labor supply, especially if households mainly provide labor
to their own land and there are imperfect labor markets.
A decrease in labor supply is, however, likely associated with an
increase in fertility.
}
We estimated the effect of both the dummy for crop loss and the log
of crop lost on hours worked per week for women in our sample.%
\footnote{
Appendix Table \ref{tab:hours} shows the results.
}
A shock does, indeed, increase the hours worked per week, but the 
effect is only slightly above 1 hour for the dummy measure of crop loss, 
and that effect is not statistically significant.
Using log of crop loss, the effect is statistically significant
although the effect is still relatively small; 
a 1\% increase in the size of a crop loss leads to a 0.08 hours
increase in amount worked per week.

Second, we can disentangle the different types of contraceptives.
Given the small sample size this is, however, at best suggestive.
Excluding abstinence, the effects of crop loss on contraceptive use
are smaller and no longer statistically significant.%
\footnote{
Appendix Table \ref{tab:excludeAbs} shows the results.
}
With the relatively low use rates of modern contraceptives, this
is not surprising given that abstinence is clearly the only 100\%
effective traditional contraceptive.
In sum, combined with an average work week of just below 44 hours a week,
it is unlikely that a substantial amount of the reduction in 
fertility is explained by lower coital frequency brought on by
high work load.

% Third, 
% how can the effect on postponement of fertility be so
% strong if traditional contraceptives are generally considered 
% to be of low efficacy?

The final question is how the effect on postponement of fertility 
is so strong when there is almost no use of modern contraceptives.
In Table \ref{tab:effectiveness}, we examine to what extent shocks cause
a delay in pregnancy through the increased use of contraceptives by
estimating the following: 
the effects of crop loss between the prior survey and the
survey before that, contraceptive use at the prior survey round, and their
interaction on whether the respondent is currently pregnant.
Contraceptive use is clearly endogenous, but this can provide 
suggestive evidence on whether shocks do delay
pregnancy through increased use of contraceptives.

\begin{table}[htbp]
\begin{center}
\begin{threeparttable}
\caption{Contraceptive Use and Pregnancy}
\label{tab:effectiveness}
\begin{tabular}{@{} l  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
\toprule
                                                       & \multicolumn{3}{c}{Currently Pregnant} \\ \cmidrule(lr){2-4}
                                                       &\mco{Any}            & \mco{Traditional}   & \mco{Modern}        \\ \midrule
Contraceptives (7 months)                              &       0.174\sym{**} &       0.161\sym{**} &       0.300\sym{*}  \\
                                                       &     (0.067)         &     (0.066)         &     (0.167)         \\
Crop loss (7-14 months) \X Any Contraceptive (7 months)&      -0.253\sym{**} &      -0.254\sym{**} &      -0.314         \\
                                                       &     (0.110)         &     (0.113)         &     (0.235)         \\
Crop loss - 7-14 months (200 TZS or above)             &       0.047         &       0.039         &       0.031         \\
                                                       &     (0.051)         &     (0.052)         &     (0.049)         \\
\addlinespace 
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Observations                                           &    \mco{741}        &    \mco{741}        &    \mco{741}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
% All values of crop lost and assets are per capita and measure in 1,000 TZS.
% Variables not shown: Age of mother and number of births.
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

Crop loss, by itself, has a small, not statistically significant,
and positive association with pregnancy.
Using contraceptives is associated with large and statistically 
significant, but \emph{higher} likelihood of pregnancy.
The interaction term of crop loss and contraceptive use is,
however, negative, large, and statistically significant. 
With the caveat that contraceptive use is endogenous, this suggests
that women using contraceptives following a crop loss are significantly 
less likely to be pregnant. 
In order words, once there is a strong enough incentive to avoid
becoming pregnant, traditional contraceptives have a surprisingly high
effectiveness rate.%
\footnote{
The statistically significant and large positive effect of 
contraceptive use by itself it likely the result of capturing
that these women are also the most likely to be sexually active.
}
This also suggests that the crop loss is affecting conceptions primarily 
through contraceptive use, and not through some other channel.

\subsection{Robustness of Results}

Finally, we present a number of robustness checks for our results.
We first turn to household wealth.
There are three, potentially opposite, effects of wealth.
First,
wealthier households should be able to smooth consumption more easily,
which will mute their fertility and contraceptive response to the shock.
Second, 
wealthier households likely have better access to and knowledge
of contraceptives.
Third, 
wealthier households are, on average, healthier, which means that any 
effects on postponement of fertility and contraceptive use for wealthier
households is more likely to be the result of intentional decisions 
rather than worse health resulting from the shock.

As crop loss may affect asset levels, we should ideally use assets owned 
prior to any observed shocks.
Unfortunately, we do not have data available on assets prior to the first 
survey round.
We therefore use assets owned in the first round, even though 
shocks that occurred before the first round likely affect these.%
\footnote{
Appendix Table \ref{tab:assets} expands the main results above by adding 
the interaction of crop loss and assets.
}
In all cases, the coefficients for the interaction of assets in round 1 and 
crop loss are small and not statistically significant,
suggesting that wealth level may not be a determining
factor regarding the contraceptive use decision following shocks. 
There are three potential reasons why we do not find an effect of wealth.
First, the opposite predicted effects of wealth cancel each other out.
Second, even ``wealthy'' households in this area are  poor and
there may be too little variation in wealth level to capture any effects.
Third, there is likely measurement error in the asset levels. 
There are significant variations in reported household asset levels between 
periods, and it is possible that part of these variations are a result of 
reporting errors rather than actual variation in wealth.

We next turn to education.
Education is of interest for two reasons.
First, there is a positive correlation between education and wealth and 
education is unlikely to change over time in response to shocks.%
\footnote{
As we discuss above, there is still the potential for substantial
measurement error, where the level of reported education changes
from survey round to survey round.
None of the reported changes appear to be because of people going 
back to school.
}
Second, education, especially for women, might affect the efficacy of 
traditional contraceptives \citep{Rosenzweig1989}.%
\footnote{
Appendix Table \ref{tab:education} presents the effect of crop loss on
likelihood of pregnancy, birth, and contraceptive use for women
by the three levels of education discussed above 
}

For women with 7 or more years of education---which is the largest
group in the sample---the estimates are consistent with the 
main results, with statistically significant effects of crop loss 
across all outcomes, except use of modern contraceptives.
The reductions in likelihoods of being pregnant or having a birth
following crop loss are now close to each other at 17 and 19 percentage
points, and the increase in use of any contraception is 10 percentage
points.
For the other two education groups, the effects are mostly insignificant.

Using the household head's education level leads to broadly 
similar results as for the respondents' education. 
A crop loss leads to reductions in pregnancy and birth by
15 and 19 percentage points and an increase of 8 percentage points
in the use of any contraception.
For the middle-education group there is a statistically significant
effect on birth, but the impacts on the other outcomes are close to zero and
not statistically significant.
The no-education group shows a large and statistically significant
effect of contraceptive use following a crop loss, but the
effects on pregnancy and birth are about half the size and
not statistically significant.

Taken together, these results make it difficult to draw strong
conclusions on the impacts of wealth or education on responses to shocks.
The best-educated households have results that are
most closely in line with our main results above, but 
the no-education households also show effects that are
consistent with the main results.
If education is a proxy for wealth, we would 
expect that better-educated households should show less
of an effect of crop loss, but instead they show the largest.
In addition, for these households the difference between
the effects on pregnancy and births and the effects on
contraceptive use are similar to what we find for the main results.
Furthermore, if worse health after a crop loss was responsible for the
effects on birth and pregnancy, we should expect exactly
the opposite result.
As we discuss above, a potential problem with relying on 
reported years of schooling is that Tanzania's education reform and the
subsequent roll-back led to a lower correlation between years of
schooling and human capital accumulation than what we see in other
countries.

% [robustness checks]

% - Something else changes in the village
It is also possible that households respond to changed circumstances
in their local area.
We therefore used the fraction of households that experienced a crop
loss in the community---over both the last 7 months and the period 
between the two prior surveys.%
\footnote{
This fraction is calculated on the complete sample of households
rather than the sample of women used here.
Results are shown in Appendix Table \ref{tab:vill_croploss}.
}
The effect on pregnancy remains the same, whereas the effect on
births is slightly larger, although not statistically significantly so.
There is no change in the effect on the different measures of
contraceptive use.
Furthermore, none of the village level coefficients are statistically 
significant.
Another way to look at this is that the spillovers from other
households' exposure to shocks are not large enough to force 
not exposed households to change their fertility and contraceptive
behavior.

% - preemptive behavior (pregnant / contraceptive use predict shocks)

Finally, households may anticipate shocks and
change their behavior in response to the expected shocks.
One way to examine this is to reverse the estimations.
We therefore estimate whether contraceptive use at the time of the last 
survey was associated with observing a crop loss between the last survey
and the current one and similarly for pregnancy at prior survey and 
birth in the period before the prior survey.%
\footnote{
Appendix Table \ref{tab:reverse} shows the results.
We examine both the dummy for crop loss and the log of crop loss.
}
There are no statistically significant direct effects of 
prior behavior on current crop loss, measured as either a
dummy or log of amount of crop loss.
The small effects and lack of statistical significance 
indicate that crop losses are, indeed, unanticipated 
shocks.

\section{Conclusion}

We estimate the relationship between household income shocks and decisions
on timing of fertility. 
To account for potential correlation between unobservable household
characteristics and both shocks and fertility decisions, we employ a fixed effect model.
Our estimates show that both the likelihood of pregnancy and birth decrease in
response to a crop loss and that contraceptive use, especially use of traditional 
methods like abstinence and rhythm, significantly increases.

We examine a number of potential explanations for these results.
Crop loss shocks do not worsen women's health enough to 
reduce pregnancies and childbirths to the extent that we find.
Furthermore, with no significant change in the likelihood of migration or 
dissolution of partnership in response to crop loss, there is no evidence 
that the reduction is due to physical separation of spouses or partner.
As expected, the number of hours worked increases for women after a shock, 
but the effect is unlikely to be large enough to affect fertility 
through a change in coital frequency.

Based on this, our main conclusion is that households consciously delay 
fertility through adjustment in contraceptive use to cope 
with income shocks. 
This is especially of interest since there is little use of modern
contraceptives in this setting.
If traditional contraceptives have low efficacy, a natural question is 
how the substantial reductions in likelihood of births and pregnancy 
come about.
We present suggestive evidence that show that only those households
with a strong incentive, i.e. those exposed to crop loss, are 
effective users of traditional contraceptives.%
\footnote{
This is in line with the results in \cite{Rosenzweig1989},
where women with the highest opportunity cost of children
are also the most effective user of traditional contraceptives,
although the setting is very different.
}

There are two potential caveats to our results.
First, the sample size is small and the survey rounds
could have been further apart, but the longitudinal nature
of the data means that we can control for unobservable 
time-invariant individual and household characteristics.
Second, not surprisingly there appears to be measurement 
errors.
Reported age and education levels are not always consistent
across rounds, which is an indication that we probably also
have measurement errors in the other variables, although it is
harder to verify.
To the extent that these measurement errors are classical 
measurement errors, this biases our results towards zero,
making it harder to show an effect of crop loss.%
\footnote{
The problems with measurement error, combined with the small
sample size, may also explain why we do not find clear
differences in the response to crop loss across education
and wealth levels.
}

The data unfortunately do not allow us to establish whether there 
are long-term impacts on fertility outcomes.
Results from Guatemala indicate that only shocks that occur 
toward the end of a woman's reproductive life is likely to have an 
impact on final fertility \citep{Portner2014}.
There is, therefore, little reason to believe that temporary shocks,
like the crop losses studied here, lead to over-all changes in fertility.

Hence, an important unanswered question is what the implications
are for our understanding of people's ability to control long-term
fertility when there is limited access to modern contraceptives.
As we have shown, it is possible for households to, at least temporarily, 
limit fertility using traditional methods once the incentives are
strong enough.
What we do not know is whether the psychic cost of using traditional 
contraceptives is so high that although households are willing to use 
them to deal with short-term shocks, they become difficult to use for 
long-run moderation of fertility.
That studies of the introduction of modern family planning, using individual
level data, rarely find reductions in fertility larger than one child
is an indication that households were already able to limit long-term 
fertility using traditional contraceptives.%
\footnote{
See, for example, \cite{Portner2014a} and the literature surveyed therein.
}

This does not mean that the ability to postpone birth is unimportant
or that there is not an important role for modern contraceptives.
Children born after shocks are likely to fare worse than children born 
during better times.
The mother is also likely to fare better if she does not give birth 
immediately following an economic shock that may impact her ability 
to recover and care for the new child.
Our results imply that providing better access to contraceptives 
during times of economic shocks could improve the ability of households 
to postpone births.
In turn, this could lead to better mother and child health outcomes.

\clearpage

\newpage
\bibliographystyle{econometrica}
\bibliography{incomeShocks-jde-r1}
\addcontentsline{toc}{section}{References}

% \newpage

\clearpage

\appendix
%\renewcommand\thesection{Appendix \Alph{section}:}

\renewcommand\thetable{A-\arabic{table}}
\renewcommand\thefigure{A-\arabic{figure}}
\setcounter{table}{0}
\setcounter{figure}{0}

\section{Appendix}

% Attrition descriptive statistics

\begin{table}[htbp]
\centering
\footnotesize
\begin{threeparttable}
\caption{Descriptive statistics for Crop loss and Outcomes}
\label{tab:attrition_desc_stat_croploss}
\begin{tabular}{l  D{.}{.}{2,3} D{.}{.}{2,3} D{.}{.}{2,3} D{.}{.}{2,3} D{.}{.}{2,3}} \toprule
                                  & \multicolumn{4}{c}{Wave} & \mco{Average} \\ \cmidrule(lr){2-5}
                                  &\multicolumn{1}{c}{1} &\multicolumn{1}{c}{2}   &\multicolumn{1}{c}{3} &\multicolumn{1}{c}{4} &   \\ \midrule
Dummy crop loss (1-7 months) $\geq$ TZS 200            &        0.58&        0.22&        0.01&        0.00&        0.34\\
                                                       &      (0.49)&      (0.42)&      (0.12)&      (0.00)&      (0.48)\\ 
\addlinespace 
Currently pregnant                                     &        0.14&        0.10&        0.08&        0.00&        0.11\\
                                                       &      (0.34)&      (0.30)&      (0.27)&      (0.00)&      (0.31)\\
Gave birth since last survey                           &        0.19&        0.12&        0.12&        0.07&        0.15\\
                                                       &      (0.39)&      (0.33)&      (0.33)&      (0.27)&      (0.36)\\
\addlinespace 
Contraceptive use                                      &        0.12&        0.17&        0.15&        0.19&        0.14\\
                                                       &      (0.33)&      (0.37)&      (0.36)&      (0.40)&      (0.35)\\
Contraceptive use -- Traditional\tnote{a}              &        0.08&        0.10&        0.05&        0.11&        0.08\\
                                                       &      (0.27)&      (0.30)&      (0.23)&      (0.32)&      (0.28)\\
Contraceptive use -- Modern\tnote{b}                   &        0.05&        0.07&        0.09&        0.07&        0.06\\
                                                       &      (0.21)&      (0.26)&      (0.29)&      (0.27)&      (0.25)\\ 
\bottomrule
\end{tabular}
\begin{tablenotes} 
\scriptsize
\item[a] Traditional contraceptives include abstinence and rhythm method.
\item[b] Modern contraceptives include condom, diaphragm, pill, IUD, injection, female and 
male sterilization.
\end{tablenotes}
\end{threeparttable}
\end{table}

\begin{table}[htbp]
\centering
\footnotesize
\begin{threeparttable}
\caption{Wave 1 Descriptive Statistics for Women}
\label{tab:attrition_desc_stat_women}
\begin{tabular}{l  D{.}{.}{3,2} D{.}{.}{3,2} D{.}{.}{3,2} D{.}{.}{3,2}} \toprule
                            	 & \mco{Mean} 	 & 	\mco{St Dev}	  \\ \midrule
Age 18-22                                              &        0.29&        0.45\\
Age 23-27                                              &        0.17&        0.38\\
Age 28-32                                              &        0.18&        0.38\\
Age 33-37                                              &        0.14&        0.35\\
Age 38-45                                              &        0.22&        0.41\\ 
\addlinespace 
No education                                           &        0.23&        0.42\\
1 - 6 years of education                               &        0.24&        0.43\\
7 plus years of education                              &        0.53&        0.50\\ 
\addlinespace 
Assets per capita in wave 1 (10,000 TZS)\tnote{a}      &        9.24&       21.44\\
\addlinespace 
Number of women                                        &     \mct{207}           \\
\bottomrule
\end{tabular}
\begin{tablenotes} 
\scriptsize
\item[a] Assets capture self-reported values of land, livestock, business assets, durable
goods, and savings.
\end{tablenotes}
\end{threeparttable}
\end{table}

% OLS estimates

\begin{table}[htbp]
\begin{center}
\begin{small}
\begin{threeparttable}
\caption{Full OLS and Community Fixed Effects Results for Fertility Outcomes}
\label{tab:ols_fertility}
\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} @{}}
\toprule
                                                       &\multicolumn{2}{c}{Currently pregnant}&\multicolumn{2}{c}{Birth since last survey}\\ \cmidrule(lr){2-3} \cmidrule(lr){4-5}
                                                       &                     & \mco{Community}     &                     & \mco{Community}      \\ 
                                                       & \mco{OLS}           & \mco{Fixed Effects} & \mco{OLS}           & \mco{Fixed Effects}  \\ 
\midrule
Crop loss - 1-7 months (200 TZS or above)              &      -0.053         &      -0.074\sym{*}  &                     &                     \\
                                                       &     (0.038)         &     (0.039)         &                     &                     \\
Crop loss - 7-14 months (200 TZS or above)             &                     &                     &      -0.103\sym{***}&      -0.115\sym{***}\\
                                                       &                     &                     &     (0.040)         &     (0.042)         \\
Assets per capita in wave 1 (10,000 TZS)               &      -0.001         &      -0.002\sym{**} &       0.000         &       0.001         \\
                                                       &     (0.001)         &     (0.001)         &     (0.001)         &     (0.001)         \\
1 - 6 years of education                               &      -0.011         &      -0.005         &       0.006         &       0.006         \\
                                                       &     (0.027)         &     (0.027)         &     (0.029)         &     (0.032)         \\
7 plus years of education                              &      -0.016         &       0.002         &       0.026         &       0.042         \\
                                                       &     (0.028)         &     (0.029)         &     (0.029)         &     (0.029)         \\
Age 23-27                                              &       0.017         &       0.022         &      -0.036         &      -0.027         \\
                                                       &     (0.033)         &     (0.036)         &     (0.033)         &     (0.032)         \\
Age 28-32                                              &      -0.037         &      -0.033         &      -0.060\sym{*}  &      -0.059\sym{*}  \\
                                                       &     (0.031)         &     (0.034)         &     (0.034)         &     (0.034)         \\
Age 33-37                                              &      -0.079\sym{**} &      -0.068\sym{**} &      -0.055         &      -0.057         \\
                                                       &     (0.032)         &     (0.034)         &     (0.035)         &     (0.036)         \\
Age 38-45                                              &      -0.174\sym{***}&      -0.156\sym{***}&      -0.184\sym{***}&      -0.194\sym{***}\\
                                                       &     (0.034)         &     (0.036)         &     (0.032)         &     (0.035)         \\
Constant                                               &       0.270\sym{***}&       0.293\sym{***}&       0.278\sym{***}&       0.237\sym{**} \\
                                                       &     (0.048)         &     (0.091)         &     (0.048)         &     (0.103)         \\
\addlinespace 
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Community fixed effects                                &    \mco{No}         &    \mco{Yes}        &    \mco{No}         &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{No }        &    \mco{No }        &    \mco{No }        &    \mco{No }        \\
Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
Assets are per capita and measured in 10,000 TZS.
\end{tablenotes}
\end{threeparttable}
\end{small}
\end{center}
\end{table}

\begin{table}[htbp]
\begin{center}
\begin{footnotesize}
\begin{threeparttable}
\caption{Full OLS and Community Fixed Effects Results for Contraceptive Use}
\label{tab:ols_contraceptive}
\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
\toprule
                                                       &                    \mct{Any}              &  \mct{Traditional}                        & \mct{Modern}        \\  \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
                                                       &                     & \mco{Community}     &                     & \mco{Community}     &                     & \mco{Community}      \\ 
                                                       & \mco{OLS}           & \mco{Fixed Effects} & \mco{OLS}           & \mco{Fixed Effects} & \mco{OLS}           & \mco{Fixed Effects}  \\

\midrule
Crop loss - 1-7 months                                 &       0.061\sym{**} &       0.059\sym{**} &       0.052\sym{**} &       0.048\sym{**} &      -0.001         &       0.004         \\
\hs (200 TZS or above)                                 &     (0.027)         &     (0.027)         &     (0.023)         &     (0.022)         &     (0.014)         &     (0.013)         \\
Assets per capita in wave 1                            &       0.001         &       0.001         &       0.001         &       0.001         &       0.001         &       0.000         \\
\hs (10,000 TZS)                                       &     (0.001)         &     (0.001)         &     (0.001)         &     (0.001)         &     (0.001)         &     (0.001)         \\
1 - 6 years of education                               &       0.025         &      -0.015         &       0.012         &      -0.020         &       0.020         &       0.003         \\
                                                       &     (0.022)         &     (0.023)         &     (0.019)         &     (0.020)         &     (0.013)         &     (0.015)         \\
7 plus years of education                              &       0.123\sym{***}&       0.095\sym{***}&       0.076\sym{**} &       0.056\sym{*}  &       0.058\sym{***}&       0.035\sym{**} \\
                                                       &     (0.035)         &     (0.036)         &     (0.030)         &     (0.030)         &     (0.019)         &     (0.016)         \\
Age 23-27                                              &       0.013         &       0.016         &       0.005         &       0.024         &       0.021         &      -0.000         \\
                                                       &     (0.037)         &     (0.038)         &     (0.024)         &     (0.028)         &     (0.030)         &     (0.023)         \\
Age 28-32                                              &       0.074\sym{*}  &       0.060         &       0.059\sym{*}  &       0.066\sym{*}  &       0.032         &      -0.012         \\
                                                       &     (0.039)         &     (0.044)         &     (0.033)         &     (0.039)         &     (0.021)         &     (0.019)         \\
Age 33-37                                              &       0.072\sym{*}  &       0.075\sym{*}  &       0.061\sym{*}  &       0.084\sym{**} &       0.048\sym{*}  &       0.018         \\
                                                       &     (0.040)         &     (0.043)         &     (0.035)         &     (0.036)         &     (0.029)         &     (0.023)         \\
Age 38-45                                              &       0.045         &       0.044         &       0.033         &       0.028         &       0.024         &       0.010         \\
                                                       &     (0.047)         &     (0.061)         &     (0.038)         &     (0.047)         &     (0.021)         &     (0.022)         \\
Constant                                               &      -0.005         &      -0.014         &       0.007         &      -0.002         &      -0.021         &      -0.017         \\
                                                       &     (0.040)         &     (0.065)         &     (0.035)         &     (0.056)         &     (0.018)         &     (0.024)         \\
\addlinespace 
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Community dummies                                      &    \mco{No}         &    \mco{Yes}        &    \mco{No}         &    \mco{Yes}        &    \mco{No}         &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{No }        &    \mco{No }        &    \mco{No }        &    \mco{No }        &    \mco{No }        &    \mco{No }        \\
Observations                                           &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \tiny
\item \hspace*{-0.5em} \textbf{Note.}
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
Assets are per capita and measured in 10,000 TZS.
\end{tablenotes}
\end{threeparttable}
\end{footnotesize}
\end{center}
\end{table}

% Community level clustering for two main tables

\begin{table}[htbp]
\begin{center}
\begin{threeparttable}
\caption{The Effects of Crop Loss on Pregnancy and Births}
\label{tab:com_cluster_birth}
\begin{tabular}{@{} l D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5}  @{}}
\toprule
                                                       &\multicolumn{3}{c}{Currently pregnant}&\multicolumn{3}{c}{Birth since last survey}\\ \cmidrule(lr){2-4} \cmidrule(lr){5-7}
                                                       &                     & \mct{Fixed Effects}                       &                     & \mct{Fixed Effects}                       \\ \cmidrule(lr){3-4} \cmidrule(lr){6-7}
                                                       & \mco{OLS}           & \mco{Community}     & \mco{Woman}         & \mco{OLS}           & \mco{Community}     & \mco{Woman}         \\ \midrule
Crop loss - 1-7 months                                 &      -0.053         &      -0.074         &      -0.093\sym{*}  &                     &                     &                     \\
\hs (200 TZS or above)                                 &     (0.041)         &     (0.045)         &     (0.049)         &                     &                     &                     \\
Crop loss - 7-14 months                                &                     &                     &                     &      -0.103\sym{***}&      -0.115\sym{***}&      -0.159\sym{***}\\
\hs (200 TZS or above)                                 &                     &                     &                     &     (0.036)         &     (0.039)         &     (0.046)         \\
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Community fixed effects                                &    \mco{No}         &    \mco{Yes}        &    \mco{No}         &    \mco{No}         &    \mco{Yes}        &    \mco{No}         \\
Woman fixed effects                                    &    \mco{No}         &    \mco{No}         &    \mco{Yes}        &    \mco{No}         &    \mco{No}         &    \mco{Yes}        \\
Observations                                           &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{741}        &    \mco{741}        &    \mco{741}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule    
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
All models are linear probability models.
Robust standard errors clustered at community level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
% Assets are per capita and measured in 10,000 TZS.
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{table}[htbp]
\begin{center}
\begin{threeparttable}
\caption{The Effects of Crop Loss on Contraceptive\\ Use Overall and By Type}
\label{tab:com_cluster_contraceptive}
\begin{tabular}{@{} l D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
\toprule
                                                       &                     & \mct{Fixed Effects}                       \\ \cmidrule(lr){3-4} 
                                                       & \mco{OLS}           & \mco{Community}     & \mco{Woman}         \\ \midrule
                                                       &\multicolumn{3}{c}{Any contraception}  \\ \cmidrule(lr){2-4}
Crop loss - 1-7 months (200 TZS or above)              &       0.061\sym{***}&       0.059\sym{**} &       0.070\sym{**} \\
                                                       &     (0.022)         &     (0.024)         &     (0.032)         \\
\addlinespace 
                                                       &\multicolumn{3}{c}{Traditional contraception}  \\ \cmidrule(lr){2-4}
Crop loss - 1-7 months (200 TZS or above)              &       0.052\sym{**} &       0.048\sym{**} &       0.054\sym{**} \\
                                                       &     (0.021)         &     (0.022)         &     (0.025)         \\
\addlinespace 
                                                       &\multicolumn{3}{c}{Modern contraception}  \\ \cmidrule(lr){2-4}
Crop loss - 1-7 months (200 TZS or above)              &      -0.001         &       0.004         &       0.003         \\
                                                       &     (0.013)         &     (0.013)         &     (0.021)         \\
\addlinespace                                                        
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Community fixed effects                                &    \mco{No}         &    \mco{Yes}        &    \mco{No}         \\
Woman fixed effects                                    &    \mco{No}         &    \mco{No}         &    \mco{Yes}        \\
Observations                                           &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
All models are linear probability models.
Robust standard errors clustered at community level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
% Assets are per capita and measured in 10,000 TZS.
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

% Logit versions of main dummy set-up

\begin{table}[htbp]
\begin{center}
\begin{small}
\begin{threeparttable}
\caption{The Effects of Crop Loss -- Logit Fixed Effects Estimates}
\label{tab:logit}
\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
\toprule
                                                       & \mco{}              &\mco{}               &\multicolumn{3}{c}{Contraceptive Use}\\ \cmidrule(lr){4-6}
                                                       & \mco{Pregnant}      &\mco{Birth}          &\mco{Any}            &\mco{Traditional}    & \mco{Modern}        \\ \midrule
 & \multicolumn{5}{l}{Without interaction with assets} \\
Crop loss - 1-7 months                                 &      -0.607\sym{**} &                     &       0.804\sym{**} &       0.764\sym{*}  &       0.299         \\
                                                       &     (0.274)         &                     &     (0.390)         &     (0.444)         &     (0.691)         \\
Crop loss - 7-14 months                                &                     &      -0.946\sym{***}&                     &                     &                     \\
                                                       &                     &     (0.294)         &                     &                     &                     \\
\addlinespace 
 & \multicolumn{5}{l}{With interaction with assets} \\
Crop loss - 1-7 months                                 &      -0.570\sym{*}  &                     &       1.107\sym{**} &       0.934\sym{*}  &       1.379         \\
                                                       &     (0.335)         &                     &     (0.449)         &     (0.500)         &     (0.945)         \\
Crop loss - 1-7 months                                 &      -0.006         &                     &      -0.027         &      -0.013         &      -0.070         \\
\hs \X initial assets (10,000 TZS)                                                       &     (0.032)         &                     &     (0.021)         &     (0.019)         &     (0.045)         \\
Crop loss - 7-14 months                                &                     &      -0.952\sym{***}&                     &                     &                     \\
                                                       &                     &     (0.337)         &                     &                     &                     \\
Crop loss - 7-14 months                                &                     &       0.001         &                     &                     &                     \\
\hs \X initial assets (10,000 TZS)                     &                     &     (0.020)         &                     &                     &                     \\
\addlinespace 
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Observations                                           &    \mco{560}        &    \mco{360}        &    \mco{300}        &    \mco{228}        &     \mco{88}        \\
Number of women                                        &    \mco{140}        &    \mco{120}        &     \mco{75}        &     \mco{57}        &     \mco{22}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
Assets are per capita and measured in 10,000 TZS.
\end{tablenotes}
\end{threeparttable}
\end{small}
\end{center}
\end{table}

% Log of crop loss

\begin{table}[htbp]
\begin{center}
\begin{small}
\begin{threeparttable}
\caption{The Effects of Log Crop Loss}
\label{tab:ln_croploss}
\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
\toprule
                                                       & \mco{}              &\mco{}               &\multicolumn{3}{c}{Contraceptive Use}\\ \cmidrule(lr){4-6}
                                                       & \mco{Pregnant}      &\mco{Birth}          &\mco{Any}            &\mco{Traditional}    & \mco{Modern}        \\ \midrule
 & \multicolumn{5}{c}{Without interaction with assets} \\
Log crop loss - 1-7 months                             &      -0.102\sym{*}  &                     &       0.322\sym{***}&       0.333\sym{***}&      -0.013         \\
                                                       &     (0.060)         &                     &     (0.062)         &     (0.060)         &     (0.025)         \\
Log crop loss - 7-14 months                            &                     &      -0.119\sym{*}  &                     &                     &                     \\
                                                       &                     &     (0.061)         &                     &                     &                     \\
\addlinespace 
 & \multicolumn{5}{c}{With interaction with assets} \\
Log crop loss - 1-7 months                             &      -0.135\sym{*}  &                     &       0.360\sym{***}&       0.353\sym{***}&       0.012         \\
                                                       &     (0.070)         &                     &     (0.086)         &     (0.083)         &     (0.023)         \\
Log crop loss - 1-7 months                             &       0.002\sym{*}  &                     &      -0.002         &      -0.001         &      -0.002         \\
\hs \X initial assets (10,000 TZS)                     &     (0.001)         &                     &     (0.005)         &     (0.004)         &     (0.001)         \\
Log crop loss - 7-14 months                            &                     &      -0.149\sym{**} &                     &                     &                     \\
                                                       &                     &     (0.075)         &                     &                     &                     \\
Log crop loss - 7-14 months                            &                     &       0.002         &                     &                     &                     \\
\hs \X initial assets (10,000 TZS)                     &                     &     (0.002)         &                     &                     &                     \\
\addlinespace 
 & \multicolumn{5}{c}{With interaction with log assets} \\
Log crop loss - 1-7 months                             &      -0.401         &                     &       1.684\sym{***}&       1.518\sym{**} &       0.283\sym{*}  \\
                                                       &     (0.278)         &                     &     (0.601)         &     (0.597)         &     (0.145)         \\
Log crop loss - 1-7 months                             &       0.027         &                     &      -0.122\sym{**} &      -0.106\sym{*}  &      -0.027\sym{*}  \\
\hs \X log initial assets (10,000 TZS)                 &     (0.025)         &                     &     (0.056)         &     (0.056)         &     (0.014)         \\
Log crop loss - 7-14 months                            &                     &      -0.469         &                     &                     &                     \\
                                                       &                     &     (0.309)         &                     &                     &                     \\
Log crop loss - 7-14 months                            &                     &       0.031         &                     &                     &                     \\
\hs \X log initial assets (10,000 TZS)                 &                     &     (0.026)         &                     &                     &                     \\
\addlinespace 
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is log per capita crop loss plus 1.
Assets are assets per capita in round 1 of the survey and are measured in 10,000 TZS.
Log of assets are taken off assets per capita in TZS.
\end{tablenotes}
\end{threeparttable}
\end{small}
\end{center}
\end{table}

% Linear crop loss

\begin{table}[htbp]
\begin{center}
\begin{small}
\begin{threeparttable}
\caption{The Effects of Crop Loss Using Continuous Measure}
\label{tab:continuous}
\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
\toprule
                                                       & \mco{}              &\mco{}               &\multicolumn{3}{c}{Contraceptive Use}\\ \cmidrule(lr){4-6}
                                                       & \mco{Pregnant}      &\mco{Birth}          &\mco{Any}            &\mco{Traditional}    & \mco{Modern}        \\ \midrule
 & \multicolumn{5}{c}{Without interaction with assets} \\
Crop loss - 1-7 months (10,000 TZS)                    &      -0.022         &                     &       0.102\sym{***}&       0.114\sym{***}&      -0.013         \\
                                                       &     (0.018)         &                     &     (0.030)         &     (0.023)         &     (0.008)         \\
Crop loss - 7-14 months (10,000 TZS)                   &                     &      -0.023         &                     &                     &                     \\
                                                       &                     &     (0.017)         &                     &                     &                     \\
\addlinespace 
 & \multicolumn{5}{c}{With interaction with assets} \\
Crop loss - 1-7 months (10,000 TZS)                    &      -0.041\sym{*}  &                     &       0.124\sym{***}&       0.127\sym{***}&      -0.003         \\
                                                       &     (0.024)         &                     &     (0.038)         &     (0.033)         &     (0.008)         \\
Crop loss - 1-7 months                                 &       0.001         &                     &      -0.001         &      -0.001         &      -0.001         \\
\hs \X initial assets (10,000 TZS)                     &     (0.001)         &                     &     (0.002)         &     (0.002)         &     (0.000)         \\
Crop loss - 7-14 months (10,000 TZS)                   &                     &      -0.039         &                     &                     &                     \\
                                                       &                     &     (0.025)         &                     &                     &                     \\
Crop loss - 7-14 months                                &                     &       0.001         &                     &                     &                     \\
\hs \X initial assets (10,000 TZS)                     &                     &     (0.001)         &                     &                     &                     \\
\addlinespace 
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a per capita crop loss and measured in 10,000 TZS.
Assets are assets per capita in round 1 of the survey and are measured in 10,000 TZS.
\end{tablenotes}
\end{threeparttable}
\end{small}
\end{center}
\end{table}

% Age groups and crop loss effects

\begin{table}[htbp]
\begin{center}
\begin{small}
\begin{threeparttable}
\caption{The Effects of Crop Loss by Age}
\label{tab:agegroups}
\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
\toprule
                                                       & \mco{}              &\mco{}               &\multicolumn{3}{c}{Contraceptive Use}\\ \cmidrule(lr){4-6}
                                                       & \mco{Pregnant}      &\mco{Birth}          &\mco{Any}            &\mco{Traditional}    & \mco{Modern}        \\ \midrule
Crop loss - 1-7 months \X age 18-22                    &      -0.160\sym{**} &                     &       0.096\sym{*}  &       0.048         &      -0.003         \\
                                                       &     (0.075)         &                     &     (0.052)         &     (0.044)         &     (0.016)         \\
Crop loss - 1-7 months \X age 23-27                    &      -0.170\sym{**} &                     &       0.004         &      -0.017         &       0.009         \\
                                                       &     (0.072)         &                     &     (0.066)         &     (0.063)         &     (0.019)         \\
Crop loss - 1-7 months \X age 28-32                    &      -0.153\sym{**} &                     &       0.149\sym{**} &       0.126\sym{**} &       0.029         \\
                                                       &     (0.065)         &                     &     (0.058)         &     (0.055)         &     (0.031)         \\
Crop loss - 1-7 months \X age 33-37                    &       0.031         &                     &       0.048         &       0.038         &       0.001         \\
                                                       &     (0.072)         &                     &     (0.070)         &     (0.055)         &     (0.044)         \\
Crop loss - 1-7 months \X age 38-45                    &      -0.021         &                     &       0.020         &       0.050         &      -0.034\sym{*}  \\
                                                       &     (0.043)         &                     &     (0.054)         &     (0.048)         &     (0.020)         \\
Crop loss - 7-14 months \X age 18-22                   &                     &      -0.235\sym{***}&                     &                     &                     \\
                                                       &                     &     (0.087)         &                     &                     &                     \\
Crop loss - 7-14 months \X age 23-27                   &                     &      -0.129         &                     &                     &                     \\
                                                       &                     &     (0.094)         &                     &                     &                     \\
Crop loss - 7-14 months \X age 28-32                   &                     &      -0.225\sym{***}&                     &                     &                     \\
                                                       &                     &     (0.073)         &                     &                     &                     \\
Crop loss - 7-14 months \X age 33-37                   &                     &      -0.127         &                     &                     &                     \\
                                                       &                     &     (0.082)         &                     &                     &                     \\
Crop loss - 7-14 months \X age 38-45                   &                     &      -0.042         &                     &                     &                     \\
                                                       &                     &     (0.050)         &                     &                     &                     \\
\addlinespace 
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
Age groups is based on age at the first round of the survey as described in the text.
\end{tablenotes}
\end{threeparttable}
\end{small}
\end{center}
\end{table}

% Hours worked and crop loss

\begin{table}[htbp]
\begin{center}
\begin{small}
\begin{threeparttable}
\caption{Crop loss and Hours Worked per Week for Women}
\label{tab:hours}
\begin{tabular}{@{} l D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}   @{}}
\toprule
                                                       &\multicolumn{5}{c}{Hours worked by woman} \\ \midrule
Crop loss - 1-7 months                                 &       1.391         &       1.097         &                     &                     &                     \\
\hs (200 TZS or above)                                 &     (2.319)         &     (2.543)         &                     &                     &                     \\
Crop loss - 1-7 months                                 &                     &       0.037         &                     &                     &                     \\
\hs  \X initial assets (10,000 TZS)                    &                     &     (0.103)         &                     &                     &                     \\
\addlinespace 
Log crop loss - 1-7 months                             &                     &                     &       8.013\sym{**} &      10.073\sym{**} &      15.213         \\
                                                       &                     &                     &     (4.041)         &     (4.526)         &    (25.934)         \\
Log crop loss - 1-7 months                             &                     &                     &                     &      -0.132         &                     \\
\hs \X initial assets (10,000 TZS)                     &                     &                     &                     &     (0.113)         &                     \\
Log crop loss - 1-7 months                             &                     &                     &                     &                     &      -0.647         \\
\hs \X log initial assets                              &                     &                     &                     &                     &     (2.371)         \\
\addlinespace 
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Observations                                           &    \mco{982}        &    \mco{982}        &    \mco{982}        &    \mco{982}        &    \mco{982}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
Assets are per capita and measured in 10,000 TZS.
Log crop loss is log per capita crop loss plus 1.
Log of assets are taken off assets per capita in TZS.
\end{tablenotes}
\end{threeparttable}
\end{small}
\end{center}
\end{table}

% Excluding abstinence

\begin{table}[htbp]
\begin{center}
\begin{small}
\begin{threeparttable}
\caption{Crop loss and Abstinence as Contraceptives}
\label{tab:excludeAbs}
\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}   @{}}
\toprule
                                                       & \multicolumn{4}{c}{Contraceptive Use}\\ \cmidrule(lr){2-5}
                                                       & \mct{Any, except abstinence}              & \mct{Rhythm and withdrawal}               & \mct{Only abstinence}                     \\ \midrule
Crop loss - 1-7 months                                 &       0.036         &       0.039         &       0.018         &       0.002         &       0.021         &       0.025         \\
\hs (200 TZS or above)                                 &     (0.030)         &     (0.031)         &     (0.021)         &     (0.026)         &     (0.019)         &     (0.027)         \\
Crop loss - 1-7 months                                 &                     &      -0.000         &                     &       0.002         &                     &      -0.000         \\
\hs  \X initial assets (10,000 TZS)                    &                     &     (0.003)         &                     &     (0.002)         &                     &     (0.003)         \\
\addlinespace 
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Observations                                           &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
Assets are per capita and measured in 10,000 TZS.
\end{tablenotes}
\end{threeparttable}
\end{small}
\end{center}
\end{table}

\begin{table}[htbp]
\begin{center}
\begin{small}
\begin{threeparttable}
\caption{The Effects of Crop Loss on Contraceptives Use by Assets}
\label{tab:assets}
\begin{tabular}{@{} l D{.}{.}{1.5} D{.}{.}{1.5}  D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5}  @{}}
\toprule
                                                       & \mco{}              &\mco{}               &\multicolumn{3}{c}{Contraceptive Use}\\ \cmidrule(lr){4-6}
                                                       & \mco{Pregnant}      &\mco{Childbirth}     &\mco{Any}            &\mco{Traditional}    & \mco{Modern}        \\ \midrule
Crop loss - 1-7 months (200 TZS or above)              &      -0.095\sym{**} &                     &       0.093\sym{**} &       0.065\sym{**} &       0.021         \\
                                                       &     (0.045)         &                     &     (0.037)         &     (0.032)         &     (0.015)         \\
Crop loss - 1-7 months                                 &       0.000         &                     &      -0.003         &      -0.001         &      -0.002         \\
\hs \X initial assets (10,000 TZS)                     &     (0.001)         &                     &     (0.003)         &     (0.003)         &     (0.002)         \\
Crop loss - 7-14 months (200 TZS or above)             &                     &      -0.164\sym{***}&                     &                     &                     \\
                                                       &                     &     (0.053)         &                     &                     &                     \\
Crop loss - 7-14 months                                &                     &       0.001         &                     &                     &                     \\
\hs \X initial assets (10,000 TZS)                     &                     &     (0.002)         &                     &                     &                     \\
\addlinespace 
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
Assets are per capita and measured in 10,000 TZS.
\end{tablenotes}
\end{threeparttable}
\end{small}
\end{center}
\end{table}

\begin{table}[htbp]
\begin{center}
\begin{small}
\begin{threeparttable}
\caption{The Effects of Crop Loss on Contraceptives Use by Education Level}
\label{tab:education}
\begin{tabular}{@{} l D{.}{.}{1.5} D{.}{.}{1.5}  D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5}  @{}}
\toprule
                                                       & \mco{}              &\mco{}               &\multicolumn{3}{c}{Contraceptive Use}\\ \cmidrule(lr){4-6}
                                                       & \mco{Pregnant}      &\mco{Birth}          &\mco{Any}            &\mco{Traditional}    & \mco{Modern}        \\ \midrule
                                                       & \multicolumn{5}{c}{Respondent's education level} \\
Crop loss \X no education                              &      -0.069         &                     &       0.080         &       0.041         &       0.032         \\
                                                       &     (0.062)         &                     &     (0.051)         &     (0.041)         &     (0.029)         \\
Crop loss \X 1-6 years of education                    &      -0.006         &                     &       0.019         &       0.036         &      -0.033         \\
                                                       &     (0.063)         &                     &     (0.051)         &     (0.040)         &     (0.031)         \\
Crop loss \X 7 plus years of education                 &      -0.167\sym{***}&                     &       0.098\sym{**} &       0.073\sym{*}  &       0.009         \\
                                                       &     (0.050)         &                     &     (0.043)         &     (0.039)         &     (0.018)         \\
Crop loss (7-14 months)                                &                     &      -0.165\sym{**} &                     &                     &                     \\
\hs \X no education                                    &                     &     (0.064)         &                     &                     &                     \\
Crop loss (7-14 months)                                &                     &      -0.115         &                     &                     &                     \\
\hs \X 1-6 years of education                          &                     &     (0.072)         &                     &                     &                     \\
Crop loss (7-14 months)                                &                     &      -0.185\sym{***}&                     &                     &                     \\
\hs \X 7 plus years of education                       &                     &     (0.062)         &                     &                     &                     \\
% \addlinespace 
% Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
% Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\addlinespace 
                                                       & \multicolumn{5}{c}{Head's education level} \\
Crop loss \X no education                              &      -0.071         &                     &       0.168\sym{**} &       0.149\sym{**} &       0.017         \\
                                                       &     (0.074)         &                     &     (0.080)         &     (0.069)         &     (0.040)         \\
Crop loss \X 1-6 years of education                    &      -0.024         &                     &       0.017         &       0.006         &      -0.015         \\
                                                       &     (0.061)         &                     &     (0.049)         &     (0.039)         &     (0.030)         \\
Crop loss \X 7 plus years of education                 &      -0.147\sym{***}&                     &       0.083\sym{**} &       0.062\sym{*}  &       0.013         \\
                                                       &     (0.048)         &                     &     (0.040)         &     (0.036)         &     (0.017)         \\
Crop loss (7-14 months)                                &                     &      -0.083         &                     &                     &                     \\
\hs \X no education                                    &                     &     (0.095)         &                     &                     &                     \\
Crop loss (7-14 months)                                &                     &      -0.137\sym{**} &                     &                     &                     \\
\hs \X 1-6 years of education                          &                     &     (0.066)         &                     &                     &                     \\
Crop loss (7-14 months)                                &                     &      -0.194\sym{***}&                     &                     &                     \\
\hs \X 7 plus years of education                       &                     &     (0.057)         &                     &                     &                     \\
% \addlinespace 
% Observations                                           &    \mco{964}        &    \mco{723}        &    \mco{964}        &    \mco{964}        &    \mco{964}        \\
% Number of women                                        &    \mco{241}        &    \mco{241}        &    \mco{241}        &    \mco{241}        &    \mco{241}        \\
% \addlinespace 
%                                                        & \multicolumn{5}{c}{Maximum education level in household} \\
% Crop loss \X no education                              &      -0.396\sym{***}&                     &       0.503         &       0.494         &       0.005         \\
%                                                        &     (0.026)         &                     &     (0.371)         &     (0.372)         &     (0.012)         \\
% Crop loss \X 1-6 years of education                    &      -0.016         &                     &       0.067         &       0.032         &       0.011         \\
%                                                        &     (0.065)         &                     &     (0.069)         &     (0.053)         &     (0.045)         \\
% Crop loss \X 7 plus years of education                 &      -0.114\sym{**} &                     &       0.064\sym{**} &       0.054\sym{**} &       0.000         \\
%                                                        &     (0.046)         &                     &     (0.032)         &     (0.026)         &     (0.014)         \\
% Crop loss (7-14 months)                                &                     &      -0.322\sym{*}  &                     &                     &                     \\
% \hs \X no education                                    &                     &     (0.176)         &                     &                     &                     \\
% Crop loss (7-14 months)                                &                     &      -0.118\sym{*}  &                     &                     &                     \\
% \hs \X 1-6 years of education                          &                     &     (0.071)         &                     &                     &                     \\
% Crop loss (7-14 months)                                &                     &      -0.170\sym{***}&                     &                     &                     \\
% \hs \X 7 plus years of education                       &                     &     (0.054)         &                     &                     &                     \\
\addlinespace 
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
Assets are per capita and measured in 10,000 TZS.
\end{tablenotes}
\end{threeparttable}
\end{small}
\end{center}
\end{table}

% Village crop loss

\begin{table}[htbp]
\begin{center}
\begin{small}
\begin{threeparttable}
\caption{The Effects of Individual and Village Crop Loss}
\label{tab:vill_croploss}
\begin{tabular}{@{} l D{.}{.}{1.5} D{.}{.}{1.5}  D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5}  @{}}
\toprule
                                                       & \mco{}              &\mco{}               &\multicolumn{3}{c}{Contraceptive Use}\\ \cmidrule(lr){4-6}
                                                       & \mco{Pregnant}      &\mco{Birth}          &\mco{Any}            &\mco{Traditional}    & \mco{Modern}        \\ \midrule
Crop loss - 1-7 months (200 TZS or above)              &      -0.105\sym{**} &                     &       0.071\sym{**} &       0.047\sym{**} &       0.006         \\
                                                       &     (0.049)         &                     &     (0.032)         &     (0.023)         &     (0.021)         \\
Fraction with crop loss in village                     &       0.059         &                     &      -0.003         &       0.030         &      -0.013         \\
                                                       &     (0.108)         &                     &     (0.070)         &     (0.056)         &     (0.039)         \\
Crop loss - 7-14 months (200 TZS or above)             &                     &      -0.152\sym{***}&                     &                     &                     \\
                                                       &                     &     (0.058)         &                     &                     &                     \\
Lagged fraction with crop loss in village              &                     &      -0.033         &                     &                     &                     \\
                                                       &                     &     (0.120)         &                     &                     &                     \\
\addlinespace 
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
Assets are per capita and measured in 10,000 TZS.
\end{tablenotes}
\end{threeparttable}
\end{small}
\end{center}
\end{table}

% Reverse regressions - predictive or pseudo

\begin{table}[htbp]
\begin{center}
\begin{small}
\begin{threeparttable}
\caption{The Association between Prior Fertility and Contraceptive Use and Current Crop Loss}
\label{tab:reverse}
\begin{tabular}{@{} l D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6} @{}}
\toprule
                                                       &\multicolumn{2}{c}{Crop loss - Dummy }& \multicolumn{2}{c}{} \\
                                                       &\multicolumn{2}{c}{200 TZS or above}       & \multicolumn{2}{c}{Log crop loss} \\ \midrule
                                                       & \multicolumn{4}{c}{Prior contraceptive use} \\
Used contraceptives prior survey                       &      -0.044         &      -0.034         &      -0.014         &      -0.013         \\
                                                       &     (0.049)         &     (0.054)         &     (0.012)         &     (0.012)         \\
Used contraceptives prior survey \X initial assets     &                     &      -0.001         &                     &      -0.000         \\
                                                       &                     &     (0.002)         &                     &     (0.000)         \\
\addlinespace 
                                                       & \multicolumn{4}{c}{Prior pregnancy} \\
Was pregnant prior survey                              &       0.017         &      -0.004         &       0.006         &       0.005         \\
                                                       &     (0.033)         &     (0.037)         &     (0.008)         &     (0.008)         \\
Pregnant prior survey \X initial assets                &                     &       0.003         &                     &       0.000         \\
                                                       &                     &     (0.002)         &                     &     (0.000)         \\
\addlinespace 
                                                       & \multicolumn{4}{c}{Prior birth} \\
Gave birth 7-14 months ago                             &      -0.034         &      -0.010         &      -0.002         &       0.003         \\
                                                       &     (0.029)         &     (0.031)         &     (0.007)         &     (0.008)         \\
Birth prior survey \X initial assets                   &                     &      -0.003\sym{**} &                     &      -0.001\sym{***}\\
                                                       &                     &     (0.001)         &                     &     (0.000)         \\
\addlinespace 
Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
Observations                                           &    \mco{741}        &    \mco{741}        &    \mco{741}        &    \mco{741}        \\
Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
\bottomrule
\end{tabular}
\begin{tablenotes} \footnotesize
\item \hspace*{-0.5em} \textbf{Note.}
Each panel a separate regression. 
Two left columns are linear probability models and two right columns 
are linear models.
Robust standard errors clustered at household level in parentheses; 
* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
Assets are per capita and measured in 10,000 TZS.
Log crop loss is log per capita crop loss plus 1.
Log of assets are taken off assets per capita in TZS.
\end{tablenotes}
\end{threeparttable}
\end{small}
\end{center}
\end{table}

\end{document}

### 2017-08-03 paper/incomeShocks-jde-r1.tex
Make minor edits
Fix order of packages
Include descriptive stats file for women

@@
 % Income Shocks, Contraceptive Use, and Timing of Fertility
-% 2016-08-18 - crop loss.tex
-% Begun.: 05/01/12
-% Edited: 2016-08-18
+% incomeShocks-jde-r1.tex
+% Begun.: 2017-07-13
+% Edited: 2017-07-13
 
@@
 \usepackage[title]{appendix}
+\usepackage{amsmath}
+
 \usepackage{fontspec}
@@
 \usepackage{multirow}
-\usepackage{amsmath}
 \usepackage[flushleft]{threeparttable}
@@
  \centering
- \includegraphics[width=0.6\textwidth]{../map/kagera.pdf}
+ \includegraphics[width=0.6\textwidth]{../staticPDF/kagera.pdf}
  \caption{Map of Kagera region}
@@
 % [Descriptive statistics and discussion here]
-
-\begin{table}
-\centering
-\footnotesize
-\begin{threeparttable}
-\caption{Wave 1 Descriptive Statistics for Women}
-\label{tab:desc_stat_women}
-\begin{tabular}{l  D{.}{.}{3,2} D{.}{.}{3,2} D{.}{.}{3,2} D{.}{.}{3,2}} \toprule
-                            	 & \mco{Mean} 	 & 	\mco{St Dev}	  \\ \midrule
-Age 18-22                                              &        0.23&        0.42\\
-Age 23-27                                              &        0.17&        0.38\\
-Age 28-32                                              &        0.23&        0.42\\
-Age 33-37                                              &        0.21&        0.41\\
-Age 38-45                                              &        0.16&        0.37\\ \addlinespace 
-No education                                           &        0.29&        0.45\\
-1 - 6 years of education                               &        0.26&        0.44\\
-7 plus years of education                              &        0.45&        0.50\\ \addlinespace 
-Assets per capita in wave 1 (10,000 TZS)\tnote{a}      &        7.67&       12.75\\ \addlinespace 
-Number of women                                        &     \mct{247}            \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} 
-\scriptsize
-\item[a] Assets capture self-reported values of land, livestock, business assets, durable
-goods, and savings.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
+\input{../tables/desstat1.tex}
 

### 2017-08-04 paper/incomeShocks-jde-r1.tex
Add file for crop loss and outcomes descriptive statistics

@@
 
-\begin{table}
-\centering
-\footnotesize
-\begin{threeparttable}
-\caption{Descriptive statistics for Crop loss and Outcomes}
-\label{tab:desc_stat_croploss}
-\begin{tabular}{l  D{.}{.}{2,3} D{.}{.}{2,3} D{.}{.}{2,3} D{.}{.}{2,3} D{.}{.}{2,3}} \toprule
-                                  & \multicolumn{4}{c}{Wave} & \mco{Average} \\ \cmidrule(lr){2-5}
-                                  &\multicolumn{1}{c}{1} &\multicolumn{1}{c}{2}   &\multicolumn{1}{c}{3} &\multicolumn{1}{c}{4} &   \\ \midrule
-Dummy crop loss (1-7 months) $\geq$ TZS 200            &        0.71&        0.26&        0.02&        0.03&        0.25\\
-                                                       &      (0.45)&      (0.44)&      (0.13)&      (0.18)&      (0.44)\\ 
-\addlinespace 
-Currently pregnant                                     &        0.17&        0.23&        0.15&        0.14&        0.17\\
-                                                       &      (0.37)&      (0.42)&      (0.36)&      (0.35)&      (0.38)\\
-Gave birth since last survey                           &        0.19&        0.16&        0.22&        0.12&        0.17\\
-                                                       &      (0.40)&      (0.37)&      (0.41)&      (0.32)&      (0.38)\\
-\addlinespace 
-Contraceptive use                                      &        0.15&        0.09&        0.09&        0.10&        0.11\\
-                                                       &      (0.36)&      (0.29)&      (0.29)&      (0.30)&      (0.31)\\
-Contraceptive use -- Traditional\tnote{a}              &        0.11&        0.06&        0.07&        0.04&        0.07\\
-                                                       &      (0.32)&      (0.24)&      (0.25)&      (0.21)&      (0.26)\\
-Contraceptive use -- Modern\tnote{b}                   &        0.04&        0.02&        0.02&        0.05&        0.03\\
-                                                       &      (0.20)&      (0.14)&      (0.14)&      (0.22)&      (0.18)\\ 
-\bottomrule
-\end{tabular}
-\begin{tablenotes} 
-\scriptsize
-\item[a] Traditional contraceptives include abstinence and rhythm method.
-\item[b] Modern contraceptives include condom, diaphragm, pill, IUD, injection, female and 
-male sterilization.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
+% Descriptive statistics by wave
+\input{../tables/desstat2.tex}
 

### 2017-08-06 paper/incomeShocks-jde-r1.tex
Add file with birth/pregnancy results

@@
 
-\begin{table}[htbp]
-\begin{center}
-\begin{threeparttable}
-\caption{The Effects of Crop Loss on Pregnancy and Births}
-\label{tab:birth}
-\begin{tabular}{@{} l D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5}  @{}}
-\toprule
-                                                       &\multicolumn{3}{c}{Currently pregnant}&\multicolumn{3}{c}{Birth since last survey}\\ \cmidrule(lr){2-4} \cmidrule(lr){5-7}
-                                                       &                     & \mct{Fixed Effects}                       &                     & \mct{Fixed Effects}                       \\ \cmidrule(lr){3-4} \cmidrule(lr){6-7}
-                                                       & \mco{OLS}           & \mco{Community}     & \mco{Woman}         & \mco{OLS}           & \mco{Community}     & \mco{Woman}         \\ \midrule
-Crop loss - 1-7 months                                 &      -0.053         &      -0.074\sym{*}  &      -0.093\sym{**} &                     &                     &                     \\
-\hs (200 TZS or above)                                 &     (0.038)         &     (0.039)         &     (0.043)         &                     &                     &                     \\
-Crop loss - 7-14 months                                &                     &                     &                     &      -0.103\sym{***}&      -0.115\sym{***}&      -0.159\sym{***}\\
-\hs (200 TZS or above)                                 &                     &                     &                     &     (0.040)         &     (0.042)         &     (0.050)         \\
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Community fixed effects                                &    \mco{No}         &    \mco{Yes}        &    \mco{No}         &    \mco{No}         &    \mco{Yes}        &    \mco{No}         \\
-Woman fixed effects                                    &    \mco{No}         &    \mco{No}         &    \mco{Yes}        &    \mco{No}         &    \mco{No}         &    \mco{Yes}        \\
-Observations                                           &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{741}        &    \mco{741}        &    \mco{741}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule    
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-All models are linear probability models.
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
-% Assets are per capita and measured in 10,000 TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{center}
-\end{table}
+% Table with pregnancy and birth results
+\input{../tables/main_pregnant_birth.tex}
 

### 2017-08-06 paper/incomeShocks-jde-r1.tex
Add contraceptive results table

@@
 
-\begin{table}[htbp]
-\begin{center}
-\begin{threeparttable}
-\caption{The Effects of Crop Loss on Contraceptive\\ Use Overall and By Type}
-\label{tab:contraceptive}
-\begin{tabular}{@{} l D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
-\toprule
-                                                       &                     & \mct{Fixed Effects}                       \\ \cmidrule(lr){3-4} 
-                                                       & \mco{OLS}           & \mco{Community}     & \mco{Woman}         \\ \midrule
-                                                       &\multicolumn{3}{c}{Any contraception}  \\ \cmidrule(lr){2-4}
-Crop loss - 1-7 months (200 TZS or above)              &       0.061\sym{**} &       0.059\sym{**} &       0.070\sym{**} \\
-                                                       &     (0.027)         &     (0.027)         &     (0.033)         \\
-\addlinespace 
-                                                       &\multicolumn{3}{c}{Traditional contraception}  \\ \cmidrule(lr){2-4}
-Crop loss - 1-7 months (200 TZS or above)              &       0.052\sym{**} &       0.048\sym{**} &       0.054\sym{**} \\
-                                                       &     (0.023)         &     (0.022)         &     (0.024)         \\
-\addlinespace 
-                                                       &\multicolumn{3}{c}{Modern contraception}  \\ \cmidrule(lr){2-4}
-Crop loss - 1-7 months (200 TZS or above)              &      -0.001         &       0.004         &       0.003         \\
-                                                       &     (0.014)         &     (0.013)         &     (0.020)         \\
-\addlinespace                                                        
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Community fixed effects                                &    \mco{No}         &    \mco{Yes}        &    \mco{No}         \\
-Woman fixed effects                                    &    \mco{No}         &    \mco{No}         &    \mco{Yes}        \\
-Observations                                           &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-All models are linear probability models.
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
-% Assets are per capita and measured in 10,000 TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{center}
-\end{table}
+% Table with contraceptive use results
+\input{../tables/main_contraceptives.tex}
 

### 2017-08-14 paper/incomeShocks-jde-r1.tex
Add health and marriage table

@@
 
-\begin{table}[htbp]
-\begin{center}
-\begin{threeparttable}
-\caption{The Effects of Crop Loss on Women's Health, Absence of Partner, and Marriage Dissolution}
-\label{tab:health}
-\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
-\toprule
-                                                       & \mct{Respondent} \\
-													   & \mco{BMI}           &  \mco{Illness}  	       \\ \midrule
-Crop loss - 1-7 months (200 TZS or above)              &      -0.014         &       0.023         \\
-                                                       &     (0.016)         &     (0.047)         \\
-Pregnant                                               &       0.083\sym{***}&                     \\
-                                                       &     (0.013)         &                     \\
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        \\
-Observations                                           &    \mco{988}        &    \mco{988}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        \\ 
-\addlinespace 
-                                                       &\mco{Absence/migration}&\mco{Dissolution of}\\
-                                                       &\mco{of partner}     &\mco{ of marriage}\\ \midrule
-Crop loss - 1-7 months (200 TZS or above)              &       0.011         &      -0.001         \\
-                                                       &     (0.011)         &     (0.012)         \\
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}       \\
-Observations                                           &    \mco{988}        &  \mco{1,293}        \\
-Number of women                                        &    \mco{247}        &    \mco{578}        \\
-\addlinespace 
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-The sample for the dissolution of marriage includes women who are dropped
-from the main sample because they do not show up in all four survey rounds
-or who are dropped before they are not partnered in all four survey rounds.
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-All values of crop lost and assets are per capita and measure in 1,000 TZS.
-Variables not shown: Age of mother and number of births.
-\end{tablenotes}
-\end{threeparttable}
-\end{center}
-\end{table}
+% Table with pregnancy and birth results
+\input{../tables/main_health_marriage.tex}
 

### 2017-08-17 paper/incomeShocks-jde-r1.tex
Add contraceptive's effect on pregnancy table

@@
 
-\begin{table}[htbp]
-\begin{center}
-\begin{threeparttable}
-\caption{Contraceptive Use and Pregnancy}
-\label{tab:effectiveness}
-\begin{tabular}{@{} l  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
-\toprule
-                                                       & \multicolumn{3}{c}{Currently Pregnant} \\ \cmidrule(lr){2-4}
-                                                       &\mco{Any}            & \mco{Traditional}   & \mco{Modern}        \\ \midrule
-Contraceptives (7 months)                              &       0.174\sym{**} &       0.161\sym{**} &       0.300\sym{*}  \\
-                                                       &     (0.067)         &     (0.066)         &     (0.167)         \\
-Crop loss (7-14 months) \X Any Contraceptive (7 months)&      -0.253\sym{**} &      -0.254\sym{**} &      -0.314         \\
-                                                       &     (0.110)         &     (0.113)         &     (0.235)         \\
-Crop loss - 7-14 months (200 TZS or above)             &       0.047         &       0.039         &       0.031         \\
-                                                       &     (0.051)         &     (0.052)         &     (0.049)         \\
-\addlinespace 
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Observations                                           &    \mco{741}        &    \mco{741}        &    \mco{741}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-% All values of crop lost and assets are per capita and measure in 1,000 TZS.
-% Variables not shown: Age of mother and number of births.
-\end{tablenotes}
-\end{threeparttable}
-\end{center}
-\end{table}
+% Table with contraceptive use effect on pregnancy
+\input{../tables/main_effectiveness.tex}
 

### 2017-08-18 paper/incomeShocks-jde-r1.tex
Add descriptive stats tables for women not in main sample

@@
 
-% Attrition descriptive statistics
-
-\begin{table}[htbp]
-\centering
-\footnotesize
-\begin{threeparttable}
-\caption{Descriptive statistics for Crop loss and Outcomes}
-\label{tab:attrition_desc_stat_croploss}
-\begin{tabular}{l  D{.}{.}{2,3} D{.}{.}{2,3} D{.}{.}{2,3} D{.}{.}{2,3} D{.}{.}{2,3}} \toprule
-                                  & \multicolumn{4}{c}{Wave} & \mco{Average} \\ \cmidrule(lr){2-5}
-                                  &\multicolumn{1}{c}{1} &\multicolumn{1}{c}{2}   &\multicolumn{1}{c}{3} &\multicolumn{1}{c}{4} &   \\ \midrule
-Dummy crop loss (1-7 months) $\geq$ TZS 200            &        0.58&        0.22&        0.01&        0.00&        0.34\\
-                                                       &      (0.49)&      (0.42)&      (0.12)&      (0.00)&      (0.48)\\ 
-\addlinespace 
-Currently pregnant                                     &        0.14&        0.10&        0.08&        0.00&        0.11\\
-                                                       &      (0.34)&      (0.30)&      (0.27)&      (0.00)&      (0.31)\\
-Gave birth since last survey                           &        0.19&        0.12&        0.12&        0.07&        0.15\\
-                                                       &      (0.39)&      (0.33)&      (0.33)&      (0.27)&      (0.36)\\
-\addlinespace 
-Contraceptive use                                      &        0.12&        0.17&        0.15&        0.19&        0.14\\
-                                                       &      (0.33)&      (0.37)&      (0.36)&      (0.40)&      (0.35)\\
-Contraceptive use -- Traditional\tnote{a}              &        0.08&        0.10&        0.05&        0.11&        0.08\\
-                                                       &      (0.27)&      (0.30)&      (0.23)&      (0.32)&      (0.28)\\
-Contraceptive use -- Modern\tnote{b}                   &        0.05&        0.07&        0.09&        0.07&        0.06\\
-                                                       &      (0.21)&      (0.26)&      (0.29)&      (0.27)&      (0.25)\\ 
-\bottomrule
-\end{tabular}
-\begin{tablenotes} 
-\scriptsize
-\item[a] Traditional contraceptives include abstinence and rhythm method.
-\item[b] Modern contraceptives include condom, diaphragm, pill, IUD, injection, female and 
-male sterilization.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\begin{table}[htbp]
-\centering
-\footnotesize
-\begin{threeparttable}
-\caption{Wave 1 Descriptive Statistics for Women}
-\label{tab:attrition_desc_stat_women}
-\begin{tabular}{l  D{.}{.}{3,2} D{.}{.}{3,2} D{.}{.}{3,2} D{.}{.}{3,2}} \toprule
-                            	 & \mco{Mean} 	 & 	\mco{St Dev}	  \\ \midrule
-Age 18-22                                              &        0.29&        0.45\\
-Age 23-27                                              &        0.17&        0.38\\
-Age 28-32                                              &        0.18&        0.38\\
-Age 33-37                                              &        0.14&        0.35\\
-Age 38-45                                              &        0.22&        0.41\\ 
-\addlinespace 
-No education                                           &        0.23&        0.42\\
-1 - 6 years of education                               &        0.24&        0.43\\
-7 plus years of education                              &        0.53&        0.50\\ 
-\addlinespace 
-Assets per capita in wave 1 (10,000 TZS)\tnote{a}      &        9.24&       21.44\\
-\addlinespace 
-Number of women                                        &     \mct{207}           \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} 
-\scriptsize
-\item[a] Assets capture self-reported values of land, livestock, business assets, durable
-goods, and savings.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
+% Attrition descriptive statistics across waves
+\input{../tables/appendix_desstat2.tex}
+
+\input{../tables/appendix_desstat1.tex}
 

### 2017-08-19 paper/incomeShocks-jde-r1.tex
Add full OLS/community fixed effects tables for Appendix

@@
 
-\begin{table}[htbp]
-\begin{center}
-\begin{small}
-\begin{threeparttable}
-\caption{Full OLS and Community Fixed Effects Results for Fertility Outcomes}
-\label{tab:ols_fertility}
-\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} @{}}
-\toprule
-                                                       &\multicolumn{2}{c}{Currently pregnant}&\multicolumn{2}{c}{Birth since last survey}\\ \cmidrule(lr){2-3} \cmidrule(lr){4-5}
-                                                       &                     & \mco{Community}     &                     & \mco{Community}      \\ 
-                                                       & \mco{OLS}           & \mco{Fixed Effects} & \mco{OLS}           & \mco{Fixed Effects}  \\ 
-\midrule
-Crop loss - 1-7 months (200 TZS or above)              &      -0.053         &      -0.074\sym{*}  &                     &                     \\
-                                                       &     (0.038)         &     (0.039)         &                     &                     \\
-Crop loss - 7-14 months (200 TZS or above)             &                     &                     &      -0.103\sym{***}&      -0.115\sym{***}\\
-                                                       &                     &                     &     (0.040)         &     (0.042)         \\
-Assets per capita in wave 1 (10,000 TZS)               &      -0.001         &      -0.002\sym{**} &       0.000         &       0.001         \\
-                                                       &     (0.001)         &     (0.001)         &     (0.001)         &     (0.001)         \\
-1 - 6 years of education                               &      -0.011         &      -0.005         &       0.006         &       0.006         \\
-                                                       &     (0.027)         &     (0.027)         &     (0.029)         &     (0.032)         \\
-7 plus years of education                              &      -0.016         &       0.002         &       0.026         &       0.042         \\
-                                                       &     (0.028)         &     (0.029)         &     (0.029)         &     (0.029)         \\
-Age 23-27                                              &       0.017         &       0.022         &      -0.036         &      -0.027         \\
-                                                       &     (0.033)         &     (0.036)         &     (0.033)         &     (0.032)         \\
-Age 28-32                                              &      -0.037         &      -0.033         &      -0.060\sym{*}  &      -0.059\sym{*}  \\
-                                                       &     (0.031)         &     (0.034)         &     (0.034)         &     (0.034)         \\
-Age 33-37                                              &      -0.079\sym{**} &      -0.068\sym{**} &      -0.055         &      -0.057         \\
-                                                       &     (0.032)         &     (0.034)         &     (0.035)         &     (0.036)         \\
-Age 38-45                                              &      -0.174\sym{***}&      -0.156\sym{***}&      -0.184\sym{***}&      -0.194\sym{***}\\
-                                                       &     (0.034)         &     (0.036)         &     (0.032)         &     (0.035)         \\
-Constant                                               &       0.270\sym{***}&       0.293\sym{***}&       0.278\sym{***}&       0.237\sym{**} \\
-                                                       &     (0.048)         &     (0.091)         &     (0.048)         &     (0.103)         \\
-\addlinespace 
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Community fixed effects                                &    \mco{No}         &    \mco{Yes}        &    \mco{No}         &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{No }        &    \mco{No }        &    \mco{No }        &    \mco{No }        \\
-Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
-Assets are per capita and measured in 10,000 TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{small}
-\end{center}
-\end{table}
-
-\begin{table}[htbp]
-\begin{center}
-\begin{footnotesize}
-\begin{threeparttable}
-\caption{Full OLS and Community Fixed Effects Results for Contraceptive Use}
-\label{tab:ols_contraceptive}
-\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
-\toprule
-                                                       &                    \mct{Any}              &  \mct{Traditional}                        & \mct{Modern}        \\  \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
-                                                       &                     & \mco{Community}     &                     & \mco{Community}     &                     & \mco{Community}      \\ 
-                                                       & \mco{OLS}           & \mco{Fixed Effects} & \mco{OLS}           & \mco{Fixed Effects} & \mco{OLS}           & \mco{Fixed Effects}  \\
-
-\midrule
-Crop loss - 1-7 months                                 &       0.061\sym{**} &       0.059\sym{**} &       0.052\sym{**} &       0.048\sym{**} &      -0.001         &       0.004         \\
-\hs (200 TZS or above)                                 &     (0.027)         &     (0.027)         &     (0.023)         &     (0.022)         &     (0.014)         &     (0.013)         \\
-Assets per capita in wave 1                            &       0.001         &       0.001         &       0.001         &       0.001         &       0.001         &       0.000         \\
-\hs (10,000 TZS)                                       &     (0.001)         &     (0.001)         &     (0.001)         &     (0.001)         &     (0.001)         &     (0.001)         \\
-1 - 6 years of education                               &       0.025         &      -0.015         &       0.012         &      -0.020         &       0.020         &       0.003         \\
-                                                       &     (0.022)         &     (0.023)         &     (0.019)         &     (0.020)         &     (0.013)         &     (0.015)         \\
-7 plus years of education                              &       0.123\sym{***}&       0.095\sym{***}&       0.076\sym{**} &       0.056\sym{*}  &       0.058\sym{***}&       0.035\sym{**} \\
-                                                       &     (0.035)         &     (0.036)         &     (0.030)         &     (0.030)         &     (0.019)         &     (0.016)         \\
-Age 23-27                                              &       0.013         &       0.016         &       0.005         &       0.024         &       0.021         &      -0.000         \\
-                                                       &     (0.037)         &     (0.038)         &     (0.024)         &     (0.028)         &     (0.030)         &     (0.023)         \\
-Age 28-32                                              &       0.074\sym{*}  &       0.060         &       0.059\sym{*}  &       0.066\sym{*}  &       0.032         &      -0.012         \\
-                                                       &     (0.039)         &     (0.044)         &     (0.033)         &     (0.039)         &     (0.021)         &     (0.019)         \\
-Age 33-37                                              &       0.072\sym{*}  &       0.075\sym{*}  &       0.061\sym{*}  &       0.084\sym{**} &       0.048\sym{*}  &       0.018         \\
-                                                       &     (0.040)         &     (0.043)         &     (0.035)         &     (0.036)         &     (0.029)         &     (0.023)         \\
-Age 38-45                                              &       0.045         &       0.044         &       0.033         &       0.028         &       0.024         &       0.010         \\
-                                                       &     (0.047)         &     (0.061)         &     (0.038)         &     (0.047)         &     (0.021)         &     (0.022)         \\
-Constant                                               &      -0.005         &      -0.014         &       0.007         &      -0.002         &      -0.021         &      -0.017         \\
-                                                       &     (0.040)         &     (0.065)         &     (0.035)         &     (0.056)         &     (0.018)         &     (0.024)         \\
-\addlinespace 
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Community dummies                                      &    \mco{No}         &    \mco{Yes}        &    \mco{No}         &    \mco{Yes}        &    \mco{No}         &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{No }        &    \mco{No }        &    \mco{No }        &    \mco{No }        &    \mco{No }        &    \mco{No }        \\
-Observations                                           &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \tiny
-\item \hspace*{-0.5em} \textbf{Note.}
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
-Assets are per capita and measured in 10,000 TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{footnotesize}
-\end{center}
-\end{table}
+% Pregnancy and birth results
+\input{../tables/appendix_pregnant_birth.tex}
+
+% Contraceptives results
+\input{../tables/appendix_contraceptives.tex}
 

### 2017-08-19 paper/incomeShocks-jde-r1.tex
Add community level clustering results to Appendix

@@
 Y_{i,t} &=  \beta_1 S_{i,t}  +\boldsymbol{D_{i}}'\boldsymbol{\alpha} 
-+ \mu_i + \varepsilon_{i,t},   \label{eq:fixed_effects} 
++ \mu_i + \varepsilon_{i,t},   \label{eq:fixed_effects1} 
 \end{align}
@@
 \boldsymbol{D_{i}}'\boldsymbol{\alpha} + \mu_i + \varepsilon_{i,t},   
-\label{eq:fixed_effects} 
+\label{eq:fixed_effects2} 
 \end{align}
@@
 
-% OLS estimates
-
-% Pregnancy and birth results
+% Full OLS and community fixed effects results for Pregnancy and birth
 \input{../tables/appendix_pregnant_birth.tex}
 
-% Contraceptives results
+% Community clustering for pregnancy and birth results
+\input{../tables/appendix_cluster_pregnant_birth.tex}
+
+% Full OLS and community fixed effects results for Contraceptives results
 \input{../tables/appendix_contraceptives.tex}
 
-% Community level clustering for two main tables
-
-\begin{table}[htbp]
-\begin{center}
-\begin{threeparttable}
-\caption{The Effects of Crop Loss on Pregnancy and Births}
-\label{tab:com_cluster_birth}
-\begin{tabular}{@{} l D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5}  @{}}
-\toprule
-                                                       &\multicolumn{3}{c}{Currently pregnant}&\multicolumn{3}{c}{Birth since last survey}\\ \cmidrule(lr){2-4} \cmidrule(lr){5-7}
-                                                       &                     & \mct{Fixed Effects}                       &                     & \mct{Fixed Effects}                       \\ \cmidrule(lr){3-4} \cmidrule(lr){6-7}
-                                                       & \mco{OLS}           & \mco{Community}     & \mco{Woman}         & \mco{OLS}           & \mco{Community}     & \mco{Woman}         \\ \midrule
-Crop loss - 1-7 months                                 &      -0.053         &      -0.074         &      -0.093\sym{*}  &                     &                     &                     \\
-\hs (200 TZS or above)                                 &     (0.041)         &     (0.045)         &     (0.049)         &                     &                     &                     \\
-Crop loss - 7-14 months                                &                     &                     &                     &      -0.103\sym{***}&      -0.115\sym{***}&      -0.159\sym{***}\\
-\hs (200 TZS or above)                                 &                     &                     &                     &     (0.036)         &     (0.039)         &     (0.046)         \\
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Community fixed effects                                &    \mco{No}         &    \mco{Yes}        &    \mco{No}         &    \mco{No}         &    \mco{Yes}        &    \mco{No}         \\
-Woman fixed effects                                    &    \mco{No}         &    \mco{No}         &    \mco{Yes}        &    \mco{No}         &    \mco{No}         &    \mco{Yes}        \\
-Observations                                           &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{741}        &    \mco{741}        &    \mco{741}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule    
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-All models are linear probability models.
-Robust standard errors clustered at community level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
-% Assets are per capita and measured in 10,000 TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{center}
-\end{table}
-
-\begin{table}[htbp]
-\begin{center}
-\begin{threeparttable}
-\caption{The Effects of Crop Loss on Contraceptive\\ Use Overall and By Type}
-\label{tab:com_cluster_contraceptive}
-\begin{tabular}{@{} l D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
-\toprule
-                                                       &                     & \mct{Fixed Effects}                       \\ \cmidrule(lr){3-4} 
-                                                       & \mco{OLS}           & \mco{Community}     & \mco{Woman}         \\ \midrule
-                                                       &\multicolumn{3}{c}{Any contraception}  \\ \cmidrule(lr){2-4}
-Crop loss - 1-7 months (200 TZS or above)              &       0.061\sym{***}&       0.059\sym{**} &       0.070\sym{**} \\
-                                                       &     (0.022)         &     (0.024)         &     (0.032)         \\
-\addlinespace 
-                                                       &\multicolumn{3}{c}{Traditional contraception}  \\ \cmidrule(lr){2-4}
-Crop loss - 1-7 months (200 TZS or above)              &       0.052\sym{**} &       0.048\sym{**} &       0.054\sym{**} \\
-                                                       &     (0.021)         &     (0.022)         &     (0.025)         \\
-\addlinespace 
-                                                       &\multicolumn{3}{c}{Modern contraception}  \\ \cmidrule(lr){2-4}
-Crop loss - 1-7 months (200 TZS or above)              &      -0.001         &       0.004         &       0.003         \\
-                                                       &     (0.013)         &     (0.013)         &     (0.021)         \\
-\addlinespace                                                        
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Community fixed effects                                &    \mco{No}         &    \mco{Yes}        &    \mco{No}         \\
-Woman fixed effects                                    &    \mco{No}         &    \mco{No}         &    \mco{Yes}        \\
-Observations                                           &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-All models are linear probability models.
-Robust standard errors clustered at community level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
-% Assets are per capita and measured in 10,000 TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{center}
-\end{table}
+% Community clustering for for contraceptives results
+\input{../tables/appendix_cluster_contraceptives.tex}
 

### 2017-08-19 paper/incomeShocks-jde-r1.tex
Add Logit results to Appendix

@@
 % Logit versions of main dummy set-up
-
-\begin{table}[htbp]
-\begin{center}
-\begin{small}
-\begin{threeparttable}
-\caption{The Effects of Crop Loss -- Logit Fixed Effects Estimates}
-\label{tab:logit}
-\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
-\toprule
-                                                       & \mco{}              &\mco{}               &\multicolumn{3}{c}{Contraceptive Use}\\ \cmidrule(lr){4-6}
-                                                       & \mco{Pregnant}      &\mco{Birth}          &\mco{Any}            &\mco{Traditional}    & \mco{Modern}        \\ \midrule
- & \multicolumn{5}{l}{Without interaction with assets} \\
-Crop loss - 1-7 months                                 &      -0.607\sym{**} &                     &       0.804\sym{**} &       0.764\sym{*}  &       0.299         \\
-                                                       &     (0.274)         &                     &     (0.390)         &     (0.444)         &     (0.691)         \\
-Crop loss - 7-14 months                                &                     &      -0.946\sym{***}&                     &                     &                     \\
-                                                       &                     &     (0.294)         &                     &                     &                     \\
-\addlinespace 
- & \multicolumn{5}{l}{With interaction with assets} \\
-Crop loss - 1-7 months                                 &      -0.570\sym{*}  &                     &       1.107\sym{**} &       0.934\sym{*}  &       1.379         \\
-                                                       &     (0.335)         &                     &     (0.449)         &     (0.500)         &     (0.945)         \\
-Crop loss - 1-7 months                                 &      -0.006         &                     &      -0.027         &      -0.013         &      -0.070         \\
-\hs \X initial assets (10,000 TZS)                                                       &     (0.032)         &                     &     (0.021)         &     (0.019)         &     (0.045)         \\
-Crop loss - 7-14 months                                &                     &      -0.952\sym{***}&                     &                     &                     \\
-                                                       &                     &     (0.337)         &                     &                     &                     \\
-Crop loss - 7-14 months                                &                     &       0.001         &                     &                     &                     \\
-\hs \X initial assets (10,000 TZS)                     &                     &     (0.020)         &                     &                     &                     \\
-\addlinespace 
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Observations                                           &    \mco{560}        &    \mco{360}        &    \mco{300}        &    \mco{228}        &     \mco{88}        \\
-Number of women                                        &    \mco{140}        &    \mco{120}        &     \mco{75}        &     \mco{57}        &     \mco{22}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
-Assets are per capita and measured in 10,000 TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{small}
-\end{center}
-\end{table}
+\input{../tables/appendix_logit.tex}
 

### 2017-08-21 paper/incomeShocks-jde-r1.tex
Add log croploss results to Appendix
Add continuous loss measure table to appendix

@@
 % Log of crop loss
-
-\begin{table}[htbp]
-\begin{center}
-\begin{small}
-\begin{threeparttable}
-\caption{The Effects of Log Crop Loss}
-\label{tab:ln_croploss}
-\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
-\toprule
-                                                       & \mco{}              &\mco{}               &\multicolumn{3}{c}{Contraceptive Use}\\ \cmidrule(lr){4-6}
-                                                       & \mco{Pregnant}      &\mco{Birth}          &\mco{Any}            &\mco{Traditional}    & \mco{Modern}        \\ \midrule
- & \multicolumn{5}{c}{Without interaction with assets} \\
-Log crop loss - 1-7 months                             &      -0.102\sym{*}  &                     &       0.322\sym{***}&       0.333\sym{***}&      -0.013         \\
-                                                       &     (0.060)         &                     &     (0.062)         &     (0.060)         &     (0.025)         \\
-Log crop loss - 7-14 months                            &                     &      -0.119\sym{*}  &                     &                     &                     \\
-                                                       &                     &     (0.061)         &                     &                     &                     \\
-\addlinespace 
- & \multicolumn{5}{c}{With interaction with assets} \\
-Log crop loss - 1-7 months                             &      -0.135\sym{*}  &                     &       0.360\sym{***}&       0.353\sym{***}&       0.012         \\
-                                                       &     (0.070)         &                     &     (0.086)         &     (0.083)         &     (0.023)         \\
-Log crop loss - 1-7 months                             &       0.002\sym{*}  &                     &      -0.002         &      -0.001         &      -0.002         \\
-\hs \X initial assets (10,000 TZS)                     &     (0.001)         &                     &     (0.005)         &     (0.004)         &     (0.001)         \\
-Log crop loss - 7-14 months                            &                     &      -0.149\sym{**} &                     &                     &                     \\
-                                                       &                     &     (0.075)         &                     &                     &                     \\
-Log crop loss - 7-14 months                            &                     &       0.002         &                     &                     &                     \\
-\hs \X initial assets (10,000 TZS)                     &                     &     (0.002)         &                     &                     &                     \\
-\addlinespace 
- & \multicolumn{5}{c}{With interaction with log assets} \\
-Log crop loss - 1-7 months                             &      -0.401         &                     &       1.684\sym{***}&       1.518\sym{**} &       0.283\sym{*}  \\
-                                                       &     (0.278)         &                     &     (0.601)         &     (0.597)         &     (0.145)         \\
-Log crop loss - 1-7 months                             &       0.027         &                     &      -0.122\sym{**} &      -0.106\sym{*}  &      -0.027\sym{*}  \\
-\hs \X log initial assets (10,000 TZS)                 &     (0.025)         &                     &     (0.056)         &     (0.056)         &     (0.014)         \\
-Log crop loss - 7-14 months                            &                     &      -0.469         &                     &                     &                     \\
-                                                       &                     &     (0.309)         &                     &                     &                     \\
-Log crop loss - 7-14 months                            &                     &       0.031         &                     &                     &                     \\
-\hs \X log initial assets (10,000 TZS)                 &                     &     (0.026)         &                     &                     &                     \\
-\addlinespace 
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is log per capita crop loss plus 1.
-Assets are assets per capita in round 1 of the survey and are measured in 10,000 TZS.
-Log of assets are taken off assets per capita in TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{small}
-\end{center}
-\end{table}
+\input{../tables/appendix_log_croploss.tex}
 
 % Linear crop loss
-
-\begin{table}[htbp]
-\begin{center}
-\begin{small}
-\begin{threeparttable}
-\caption{The Effects of Crop Loss Using Continuous Measure}
-\label{tab:continuous}
-\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
-\toprule
-                                                       & \mco{}              &\mco{}               &\multicolumn{3}{c}{Contraceptive Use}\\ \cmidrule(lr){4-6}
-                                                       & \mco{Pregnant}      &\mco{Birth}          &\mco{Any}            &\mco{Traditional}    & \mco{Modern}        \\ \midrule
- & \multicolumn{5}{c}{Without interaction with assets} \\
-Crop loss - 1-7 months (10,000 TZS)                    &      -0.022         &                     &       0.102\sym{***}&       0.114\sym{***}&      -0.013         \\
-                                                       &     (0.018)         &                     &     (0.030)         &     (0.023)         &     (0.008)         \\
-Crop loss - 7-14 months (10,000 TZS)                   &                     &      -0.023         &                     &                     &                     \\
-                                                       &                     &     (0.017)         &                     &                     &                     \\
-\addlinespace 
- & \multicolumn{5}{c}{With interaction with assets} \\
-Crop loss - 1-7 months (10,000 TZS)                    &      -0.041\sym{*}  &                     &       0.124\sym{***}&       0.127\sym{***}&      -0.003         \\
-                                                       &     (0.024)         &                     &     (0.038)         &     (0.033)         &     (0.008)         \\
-Crop loss - 1-7 months                                 &       0.001         &                     &      -0.001         &      -0.001         &      -0.001         \\
-\hs \X initial assets (10,000 TZS)                     &     (0.001)         &                     &     (0.002)         &     (0.002)         &     (0.000)         \\
-Crop loss - 7-14 months (10,000 TZS)                   &                     &      -0.039         &                     &                     &                     \\
-                                                       &                     &     (0.025)         &                     &                     &                     \\
-Crop loss - 7-14 months                                &                     &       0.001         &                     &                     &                     \\
-\hs \X initial assets (10,000 TZS)                     &                     &     (0.001)         &                     &                     &                     \\
-\addlinespace 
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a per capita crop loss and measured in 10,000 TZS.
-Assets are assets per capita in round 1 of the survey and are measured in 10,000 TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{small}
-\end{center}
-\end{table}
+\input{../tables/appendix_cont_croploss.tex}
 

### 2017-08-22 paper/incomeShocks-jde-r1.tex
Add age group results to appendix
Add effects on different contraceptive to appendix

@@
 % Age groups and crop loss effects
-
-\begin{table}[htbp]
-\begin{center}
-\begin{small}
-\begin{threeparttable}
-\caption{The Effects of Crop Loss by Age}
-\label{tab:agegroups}
-\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}  @{}}
-\toprule
-                                                       & \mco{}              &\mco{}               &\multicolumn{3}{c}{Contraceptive Use}\\ \cmidrule(lr){4-6}
-                                                       & \mco{Pregnant}      &\mco{Birth}          &\mco{Any}            &\mco{Traditional}    & \mco{Modern}        \\ \midrule
-Crop loss - 1-7 months \X age 18-22                    &      -0.160\sym{**} &                     &       0.096\sym{*}  &       0.048         &      -0.003         \\
-                                                       &     (0.075)         &                     &     (0.052)         &     (0.044)         &     (0.016)         \\
-Crop loss - 1-7 months \X age 23-27                    &      -0.170\sym{**} &                     &       0.004         &      -0.017         &       0.009         \\
-                                                       &     (0.072)         &                     &     (0.066)         &     (0.063)         &     (0.019)         \\
-Crop loss - 1-7 months \X age 28-32                    &      -0.153\sym{**} &                     &       0.149\sym{**} &       0.126\sym{**} &       0.029         \\
-                                                       &     (0.065)         &                     &     (0.058)         &     (0.055)         &     (0.031)         \\
-Crop loss - 1-7 months \X age 33-37                    &       0.031         &                     &       0.048         &       0.038         &       0.001         \\
-                                                       &     (0.072)         &                     &     (0.070)         &     (0.055)         &     (0.044)         \\
-Crop loss - 1-7 months \X age 38-45                    &      -0.021         &                     &       0.020         &       0.050         &      -0.034\sym{*}  \\
-                                                       &     (0.043)         &                     &     (0.054)         &     (0.048)         &     (0.020)         \\
-Crop loss - 7-14 months \X age 18-22                   &                     &      -0.235\sym{***}&                     &                     &                     \\
-                                                       &                     &     (0.087)         &                     &                     &                     \\
-Crop loss - 7-14 months \X age 23-27                   &                     &      -0.129         &                     &                     &                     \\
-                                                       &                     &     (0.094)         &                     &                     &                     \\
-Crop loss - 7-14 months \X age 28-32                   &                     &      -0.225\sym{***}&                     &                     &                     \\
-                                                       &                     &     (0.073)         &                     &                     &                     \\
-Crop loss - 7-14 months \X age 33-37                   &                     &      -0.127         &                     &                     &                     \\
-                                                       &                     &     (0.082)         &                     &                     &                     \\
-Crop loss - 7-14 months \X age 38-45                   &                     &      -0.042         &                     &                     &                     \\
-                                                       &                     &     (0.050)         &                     &                     &                     \\
-\addlinespace 
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
-Age groups is based on age at the first round of the survey as described in the text.
-\end{tablenotes}
-\end{threeparttable}
-\end{small}
-\end{center}
-\end{table}
+\input{../tables/appendix_age_groups.tex}
 
@@
 % Excluding abstinence
-
-\begin{table}[htbp]
-\begin{center}
-\begin{small}
-\begin{threeparttable}
-\caption{Crop loss and Abstinence as Contraceptives}
-\label{tab:excludeAbs}
-\begin{tabular}{@{} l D{.}{.}{2.6} D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}   @{}}
-\toprule
-                                                       & \multicolumn{4}{c}{Contraceptive Use}\\ \cmidrule(lr){2-5}
-                                                       & \mct{Any, except abstinence}              & \mct{Rhythm and withdrawal}               & \mct{Only abstinence}                     \\ \midrule
-Crop loss - 1-7 months                                 &       0.036         &       0.039         &       0.018         &       0.002         &       0.021         &       0.025         \\
-\hs (200 TZS or above)                                 &     (0.030)         &     (0.031)         &     (0.021)         &     (0.026)         &     (0.019)         &     (0.027)         \\
-Crop loss - 1-7 months                                 &                     &      -0.000         &                     &       0.002         &                     &      -0.000         \\
-\hs  \X initial assets (10,000 TZS)                    &                     &     (0.003)         &                     &     (0.002)         &                     &     (0.003)         \\
-\addlinespace 
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Observations                                           &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
-Assets are per capita and measured in 10,000 TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{small}
-\end{center}
-\end{table}
+\input{../tables/appendix_exclude_abstinence.tex}
 

### 2017-08-22 paper/incomeShocks-jde-r1.tex
Add results with interaction with assets to appendix

@@
 
-\begin{table}[htbp]
-\begin{center}
-\begin{small}
-\begin{threeparttable}
-\caption{The Effects of Crop Loss on Contraceptives Use by Assets}
-\label{tab:assets}
-\begin{tabular}{@{} l D{.}{.}{1.5} D{.}{.}{1.5}  D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5}  @{}}
-\toprule
-                                                       & \mco{}              &\mco{}               &\multicolumn{3}{c}{Contraceptive Use}\\ \cmidrule(lr){4-6}
-                                                       & \mco{Pregnant}      &\mco{Childbirth}     &\mco{Any}            &\mco{Traditional}    & \mco{Modern}        \\ \midrule
-Crop loss - 1-7 months (200 TZS or above)              &      -0.095\sym{**} &                     &       0.093\sym{**} &       0.065\sym{**} &       0.021         \\
-                                                       &     (0.045)         &                     &     (0.037)         &     (0.032)         &     (0.015)         \\
-Crop loss - 1-7 months                                 &       0.000         &                     &      -0.003         &      -0.001         &      -0.002         \\
-\hs \X initial assets (10,000 TZS)                     &     (0.001)         &                     &     (0.003)         &     (0.003)         &     (0.002)         \\
-Crop loss - 7-14 months (200 TZS or above)             &                     &      -0.164\sym{***}&                     &                     &                     \\
-                                                       &                     &     (0.053)         &                     &                     &                     \\
-Crop loss - 7-14 months                                &                     &       0.001         &                     &                     &                     \\
-\hs \X initial assets (10,000 TZS)                     &                     &     (0.002)         &                     &                     &                     \\
-\addlinespace 
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
-Assets are per capita and measured in 10,000 TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{small}
-\end{center}
-\end{table}
+% Effects of Crop Loss on Contraceptives Use by Assets
+\input{../tables/appendix_asset_interaction.tex}
 

### 2017-08-24 paper/incomeShocks-jde-r1.tex
Add education results to appendix

@@
 
-\begin{table}[htbp]
-\begin{center}
-\begin{small}
-\begin{threeparttable}
-\caption{The Effects of Crop Loss on Contraceptives Use by Education Level}
-\label{tab:education}
-\begin{tabular}{@{} l D{.}{.}{1.5} D{.}{.}{1.5}  D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5}  @{}}
-\toprule
-                                                       & \mco{}              &\mco{}               &\multicolumn{3}{c}{Contraceptive Use}\\ \cmidrule(lr){4-6}
-                                                       & \mco{Pregnant}      &\mco{Birth}          &\mco{Any}            &\mco{Traditional}    & \mco{Modern}        \\ \midrule
-                                                       & \multicolumn{5}{c}{Respondent's education level} \\
-Crop loss \X no education                              &      -0.069         &                     &       0.080         &       0.041         &       0.032         \\
-                                                       &     (0.062)         &                     &     (0.051)         &     (0.041)         &     (0.029)         \\
-Crop loss \X 1-6 years of education                    &      -0.006         &                     &       0.019         &       0.036         &      -0.033         \\
-                                                       &     (0.063)         &                     &     (0.051)         &     (0.040)         &     (0.031)         \\
-Crop loss \X 7 plus years of education                 &      -0.167\sym{***}&                     &       0.098\sym{**} &       0.073\sym{*}  &       0.009         \\
-                                                       &     (0.050)         &                     &     (0.043)         &     (0.039)         &     (0.018)         \\
-Crop loss (7-14 months)                                &                     &      -0.165\sym{**} &                     &                     &                     \\
-\hs \X no education                                    &                     &     (0.064)         &                     &                     &                     \\
-Crop loss (7-14 months)                                &                     &      -0.115         &                     &                     &                     \\
-\hs \X 1-6 years of education                          &                     &     (0.072)         &                     &                     &                     \\
-Crop loss (7-14 months)                                &                     &      -0.185\sym{***}&                     &                     &                     \\
-\hs \X 7 plus years of education                       &                     &     (0.062)         &                     &                     &                     \\
-% \addlinespace 
-% Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
-% Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\addlinespace 
-                                                       & \multicolumn{5}{c}{Head's education level} \\
-Crop loss \X no education                              &      -0.071         &                     &       0.168\sym{**} &       0.149\sym{**} &       0.017         \\
-                                                       &     (0.074)         &                     &     (0.080)         &     (0.069)         &     (0.040)         \\
-Crop loss \X 1-6 years of education                    &      -0.024         &                     &       0.017         &       0.006         &      -0.015         \\
-                                                       &     (0.061)         &                     &     (0.049)         &     (0.039)         &     (0.030)         \\
-Crop loss \X 7 plus years of education                 &      -0.147\sym{***}&                     &       0.083\sym{**} &       0.062\sym{*}  &       0.013         \\
-                                                       &     (0.048)         &                     &     (0.040)         &     (0.036)         &     (0.017)         \\
-Crop loss (7-14 months)                                &                     &      -0.083         &                     &                     &                     \\
-\hs \X no education                                    &                     &     (0.095)         &                     &                     &                     \\
-Crop loss (7-14 months)                                &                     &      -0.137\sym{**} &                     &                     &                     \\
-\hs \X 1-6 years of education                          &                     &     (0.066)         &                     &                     &                     \\
-Crop loss (7-14 months)                                &                     &      -0.194\sym{***}&                     &                     &                     \\
-\hs \X 7 plus years of education                       &                     &     (0.057)         &                     &                     &                     \\
-% \addlinespace 
-% Observations                                           &    \mco{964}        &    \mco{723}        &    \mco{964}        &    \mco{964}        &    \mco{964}        \\
-% Number of women                                        &    \mco{241}        &    \mco{241}        &    \mco{241}        &    \mco{241}        &    \mco{241}        \\
-% \addlinespace 
-%                                                        & \multicolumn{5}{c}{Maximum education level in household} \\
-% Crop loss \X no education                              &      -0.396\sym{***}&                     &       0.503         &       0.494         &       0.005         \\
-%                                                        &     (0.026)         &                     &     (0.371)         &     (0.372)         &     (0.012)         \\
-% Crop loss \X 1-6 years of education                    &      -0.016         &                     &       0.067         &       0.032         &       0.011         \\
-%                                                        &     (0.065)         &                     &     (0.069)         &     (0.053)         &     (0.045)         \\
-% Crop loss \X 7 plus years of education                 &      -0.114\sym{**} &                     &       0.064\sym{**} &       0.054\sym{**} &       0.000         \\
-%                                                        &     (0.046)         &                     &     (0.032)         &     (0.026)         &     (0.014)         \\
-% Crop loss (7-14 months)                                &                     &      -0.322\sym{*}  &                     &                     &                     \\
-% \hs \X no education                                    &                     &     (0.176)         &                     &                     &                     \\
-% Crop loss (7-14 months)                                &                     &      -0.118\sym{*}  &                     &                     &                     \\
-% \hs \X 1-6 years of education                          &                     &     (0.071)         &                     &                     &                     \\
-% Crop loss (7-14 months)                                &                     &      -0.170\sym{***}&                     &                     &                     \\
-% \hs \X 7 plus years of education                       &                     &     (0.054)         &                     &                     &                     \\
-\addlinespace 
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
-Assets are per capita and measured in 10,000 TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{small}
-\end{center}
-\end{table}
+% Effects of Crop Loss on Contraceptives Use by Education Level
+\input{../tables/appendix_education.tex}
 

### 2017-08-28 paper/incomeShocks-jde-r1.tex
Fix variable labels and change table notes

@@
 % Hours worked and crop loss
-
-\begin{table}[htbp]
-\begin{center}
-\begin{small}
-\begin{threeparttable}
-\caption{Crop loss and Hours Worked per Week for Women}
-\label{tab:hours}
-\begin{tabular}{@{} l D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6}   @{}}
-\toprule
-                                                       &\multicolumn{5}{c}{Hours worked by woman} \\ \midrule
-Crop loss - 1-7 months                                 &       1.391         &       1.097         &                     &                     &                     \\
-\hs (200 TZS or above)                                 &     (2.319)         &     (2.543)         &                     &                     &                     \\
-Crop loss - 1-7 months                                 &                     &       0.037         &                     &                     &                     \\
-\hs  \X initial assets (10,000 TZS)                    &                     &     (0.103)         &                     &                     &                     \\
-\addlinespace 
-Log crop loss - 1-7 months                             &                     &                     &       8.013\sym{**} &      10.073\sym{**} &      15.213         \\
-                                                       &                     &                     &     (4.041)         &     (4.526)         &    (25.934)         \\
-Log crop loss - 1-7 months                             &                     &                     &                     &      -0.132         &                     \\
-\hs \X initial assets (10,000 TZS)                     &                     &                     &                     &     (0.113)         &                     \\
-Log crop loss - 1-7 months                             &                     &                     &                     &                     &      -0.647         \\
-\hs \X log initial assets                              &                     &                     &                     &                     &     (2.371)         \\
-\addlinespace 
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Observations                                           &    \mco{982}        &    \mco{982}        &    \mco{982}        &    \mco{982}        &    \mco{982}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
-Assets are per capita and measured in 10,000 TZS.
-Log crop loss is log per capita crop loss plus 1.
-Log of assets are taken off assets per capita in TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{small}
-\end{center}
-\end{table}
+\input{../tables/appendix_hours.tex}
 

### 2017-08-28 paper/incomeShocks-jde-r1.tex
Update variable labels and table notes

@@
 % Village crop loss 
-
-\begin{table}[htbp]
-\begin{center}
-\begin{small}
-\begin{threeparttable}
-\caption{The Effects of Individual and Village Crop Loss}
-\label{tab:vill_croploss}
-\begin{tabular}{@{} l D{.}{.}{1.5} D{.}{.}{1.5}  D{.}{.}{1.5} D{.}{.}{1.5} D{.}{.}{1.5}  @{}}
-\toprule
-                                                       & \mco{}              &\mco{}               &\multicolumn{3}{c}{Contraceptive Use}\\ \cmidrule(lr){4-6}
-                                                       & \mco{Pregnant}      &\mco{Birth}          &\mco{Any}            &\mco{Traditional}    & \mco{Modern}        \\ \midrule
-Crop loss - 1-7 months (200 TZS or above)              &      -0.105\sym{**} &                     &       0.071\sym{**} &       0.047\sym{**} &       0.006         \\
-                                                       &     (0.049)         &                     &     (0.032)         &     (0.023)         &     (0.021)         \\
-Fraction with crop loss in village                     &       0.059         &                     &      -0.003         &       0.030         &      -0.013         \\
-                                                       &     (0.108)         &                     &     (0.070)         &     (0.056)         &     (0.039)         \\
-Crop loss - 7-14 months (200 TZS or above)             &                     &      -0.152\sym{***}&                     &                     &                     \\
-                                                       &                     &     (0.058)         &                     &                     &                     \\
-Lagged fraction with crop loss in village              &                     &      -0.033         &                     &                     &                     \\
-                                                       &                     &     (0.120)         &                     &                     &                     \\
-\addlinespace 
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Observations                                           &    \mco{988}        &    \mco{741}        &    \mco{988}        &    \mco{988}        &    \mco{988}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
-Assets are per capita and measured in 10,000 TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{small}
-\end{center}
-\end{table}
+\input{../tables/appendix_village.tex}
 

### 2017-08-28 paper/incomeShocks-jde-r1.tex
Add table on whether prior behavior affects future shocks

@@
 % Reverse regressions - predictive or pseudo
-
-\begin{table}[htbp]
-\begin{center}
-\begin{small}
-\begin{threeparttable}
-\caption{The Association between Prior Fertility and Contraceptive Use and Current Crop Loss}
-\label{tab:reverse}
-\begin{tabular}{@{} l D{.}{.}{2.6}  D{.}{.}{2.6} D{.}{.}{2.6} D{.}{.}{2.6} @{}}
-\toprule
-                                                       &\multicolumn{2}{c}{Crop loss - Dummy }& \multicolumn{2}{c}{} \\
-                                                       &\multicolumn{2}{c}{200 TZS or above}       & \multicolumn{2}{c}{Log crop loss} \\ \midrule
-                                                       & \multicolumn{4}{c}{Prior contraceptive use} \\
-Used contraceptives prior survey                       &      -0.044         &      -0.034         &      -0.014         &      -0.013         \\
-                                                       &     (0.049)         &     (0.054)         &     (0.012)         &     (0.012)         \\
-Used contraceptives prior survey \X initial assets     &                     &      -0.001         &                     &      -0.000         \\
-                                                       &                     &     (0.002)         &                     &     (0.000)         \\
-\addlinespace 
-                                                       & \multicolumn{4}{c}{Prior pregnancy} \\
-Was pregnant prior survey                              &       0.017         &      -0.004         &       0.006         &       0.005         \\
-                                                       &     (0.033)         &     (0.037)         &     (0.008)         &     (0.008)         \\
-Pregnant prior survey \X initial assets                &                     &       0.003         &                     &       0.000         \\
-                                                       &                     &     (0.002)         &                     &     (0.000)         \\
-\addlinespace 
-                                                       & \multicolumn{4}{c}{Prior birth} \\
-Gave birth 7-14 months ago                             &      -0.034         &      -0.010         &      -0.002         &       0.003         \\
-                                                       &     (0.029)         &     (0.031)         &     (0.007)         &     (0.008)         \\
-Birth prior survey \X initial assets                   &                     &      -0.003\sym{**} &                     &      -0.001\sym{***}\\
-                                                       &                     &     (0.001)         &                     &     (0.000)         \\
-\addlinespace 
-Wave dummies                                           &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Woman fixed effects                                    &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        &    \mco{Yes}        \\
-Observations                                           &    \mco{741}        &    \mco{741}        &    \mco{741}        &    \mco{741}        \\
-Number of women                                        &    \mco{247}        &    \mco{247}        &    \mco{247}        &    \mco{247}        \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes} \footnotesize
-\item \hspace*{-0.5em} \textbf{Note.}
-Each panel a separate regression. 
-Two left columns are linear probability models and two right columns 
-are linear models.
-Robust standard errors clustered at household level in parentheses; 
-* significant at 10\%; ** significant at 5\%; *** significant at 1\%.
-Crop loss is a dummy for a per capita crop loss of 200 TZS or above.
-Assets are per capita and measured in 10,000 TZS.
-Log crop loss is log per capita crop loss plus 1.
-Log of assets are taken off assets per capita in TZS.
-\end{tablenotes}
-\end{threeparttable}
-\end{small}
-\end{center}
-\end{table}
+\input{../tables/appendix_reverse.tex}
 

### 2017-08-30 response/response-jde-1.tex
Add discussion of code rewrite and sample size

@@
 [Additional discussion of overall things we have changed---if any]
+
+We rewrote the code for cleaning and analyzing the data from the ground up 
+to make it easier to follow and eventually publish on GitHub.
+This rewrite slightly lowered the sample used by three observations.
+Including all married/partnered women up to age 45, rather than exclude
+women younger than 18, the final sample size is, however, the same as 
+before and the results are essentially unchanged.
 

### 2017-08-30 paper/incomeShocks-jde-r1.tex
Change wave dummy indicator for regression equation

@@
 \begin{align}
-Y_{i,t} &=  \beta_1 S_{i,t}  +\boldsymbol{D_{i}}'\boldsymbol{\alpha} 
+Y_{i,t} &=  \beta_1 S_{i,t}  + \alpha_{t} 
 + \mu_i + \varepsilon_{i,t},   \label{eq:fixed_effects1} 
@@
 $S_{i,t}$ is whether the household experienced a shock in the relevant period,
-$\boldsymbol{D_{i}}$ is a vector of survey wave dummies, 
+$\alpha_{t}$ is a vector of survey wave dummies, 
 and $\mu_i$ is time-invariant individual specific characteristics.
@@
 Y_{i,t} &=  \beta_1 S_{i,t}  +  \beta_2 S_{i,t} \times X_{i,1} +
-\boldsymbol{D_{i}}'\boldsymbol{\alpha} + \mu_i + \varepsilon_{i,t},   
+\alpha_{t} + \mu_i + \varepsilon_{i,t},   
 \label{eq:fixed_effects2} 

### 2017-08-31 response/response-jde-1.tex
Change wave dummy indicator for regression equation
Allow easy changing of asset and crop loss divisor
Prior assets and crop loss were always divided by the same number
(implemented to ease readability of regression results), but it
was spread over two different files. Moved code from womenCommon.do
to crBase.do, to make it possible to change only one place instead.

Also split so that assets and crop loss are now divided by different
numbers (1,000,000 for assets and 10,000 for crop loss).

All of this is in response to R1's comment on coefficient and standard
errors that showed as 0.000.

On

@@
 
-[RESPONSE:]
+[RESPONSE:] We have changed the equation and the associated text.
 
@@
 
-[RESPONSE:]
+[RESPONSE:] The 0.000 coefficients were predominantly for variables
+involving initial assets. 
+We have changed the variable definition for assets, so that it is
+now measured in 1,000,000 TZS.
+
+Note that there are still a few places---mainly for estimations
+involving the use of modern contraceptives---where the estimated 
+coefficients are so small that presenting additional digit does not make
+much economic or practical sense.
+In those cases the standard errors are always equal to or larger than
+the coefficients.
+See, for example, Table A-8 on the effects of log crop loss.
 

### 2017-08-31 response/response-jde-1.tex
Minor edits

@@
 involving the use of modern contraceptives---where the estimated 
-coefficients are so small that presenting additional digit does not make
-much economic or practical sense.
+coefficients are so small that presenting additional digits does not make
+much economic sense.
 In those cases the standard errors are always equal to or larger than

### 2017-09-13 response/response-jde-1.tex
Draft response to R1's questions on Table 6 and exclude wave 1
Table 6 shows how contraceptive use only leads to reduction in
pregnancy if exposed to a shock.

@@
 
-[RESPONSE:]
+[RESPONSE:] Although we appreciate the concern about endogeneity, 
+we believe the table is worth keeping. 
+In the demography literature there is a substantial disagreement on whether
+sustained fertility reduction requires access to modern contraceptives.
+Part of the problem in this literature is that the ``intensity'' with
+which a couple uses contraceptives---either traditional or modern---is
+difficult to measure correctly.
+[something on the panel nature of the data and that that is important
+for getting at the incentives and the use over time; not perfect but
+at least is for a shorter period than other (cross-sectional) surveys]
+The table is important because it provides us a way to get at this
+intensity by showing that only couples that have a strong incentive to
+prevent pregnancies appear to be successful users of contraceptives.
+Furthermore, this shows that it is possible to postpone pregnancies with
+just traditional contraceptives because traditional contraceptives account 
+for the vast proportion of contraception used.
+
+[also speaks to whether this is a conscious decision or not] 
+The table also supports our conjecture that the estimated effects 
+[might need a less causal word] 
+of crop loss capture a conscious decision. 
+If shocks lead to lower probability of pregnancy and birth because
+of other factors than the decision to use contraceptive we should see
+a negative and statistically significant coefficient for the crop loss
+dummy by itself.
+We do not.
+Instead, what we get is a substantial negative coefficient for the interaction
+between contraceptive use and crop loss and a coefficient for crop loss 
+on its own that is positive and far from statistically significant.
 
@@
 
-[RESPONSE:]
+[RESPONSE:] We have added the results where round 1 has been excluded
+to the relevant tables.
+The results are broadly consistent with what we originally found, 
+although substantially less likely to be statistically significant 
+because of the lower variation in crop loss and the smaller sample
+size.
+
+[Already ran these; might be worthwhile having the results as an extra
+panel in the relevant tables]
 

### 2017-09-13 response/response-jde-1.tex
Add notes on work to be done based on 2017-09-13 phone conference

@@
 
-Specific comments:
+\noindent \textbf{Specific comments:}
 
@@
 [RESPONSE:]
+
+[difficult because the short period of time the data cover; only
+really cover 32 periods]
+
+Claus: run main regression with longer lags to check if any effects
+on pregnancy and birth
+
+Claus: write something on Guatemala paper and how to think about how
+long it takes to find an effect / recover from a shock
 
@@
 [RESPONSE:]
+
+[Matches with ADF's question on rainfall and with R2's first question 
+on what drives crop losses]
+
+Shamma: look for data on rainfall and papers that have used rainfall
+for Kagera data
+
+- problem is that we do not know when the crop loss occurred within
+the survey period (which is 12 months for the first round and therefore
+very long).
 
@@
 
+[see response to R1's q5 on causal (or something like that)]
+
 \item Trend
@@
 [RESPONSE:]
+
+Claus: 
+- run regression with time trend
+- run regression with time trend interacted with wave 1 household characteristics
 

### 2017-09-16 response/response-jde-1.tex
Add discussion of competing factors when successive shocks
Also minor edits and additional reference on historical demography

@@
 
+[Also include Cinnirella since that covers a similar topic]
+
 \noindent \textbf{Specific comments:}
@@
 [difficult because the short period of time the data cover; only
-really cover 32 periods]
+really cover 32 months]
 
@@
 long it takes to find an effect / recover from a shock
+
+There are two competing factors here.
+First, a negative shocks will make people more likely to postpone 
+their pregnancy/birth because of the factors discussed in the
+paper.
+Second, longer spacing between births is costly because economies
+of scale are harder to realize and because it means that the mother
+will be out of the labor force for longer.
+
+This might explain why there are negative effects for both shock
+and lagged shock, but the interaction between the two is positive.
+
+One problem with the regression is that we lose 1/4 of observations
+each time we include an additional lag.
 

### 2017-09-17 response/response-jde-1.tex
Add discussion of long-term risk to postponement response

@@
 each time we include an additional lag.
+
+Important distinction is that long term fertility is likely
+affected by the overall \emph{risk}, whereas the exact timing of
+fertility is driven by a combination of risk and shocks (probably
+mostly shocks).
+
+I have also added an interaction between crop loss for different periods
+with the idea that there is a cost to continually postponing births. If
+that holds, the effect of both shocks on their own should be same and
+the interaction the opposite. Broadly speaking that appears to hold. The
+upshot–as far as I can tell–is that what we observe is postponement and
+not an overall limitation in fertility. This makes sense given my
+results for Guatemala, where long-term risk affects the desire fertility
+*level* whereas shocks experienced affect the *timing* of fertility. I
+will try to write something more up on this.
 

### 2017-09-17 response/response-jde-1.tex
Add initial discussion of time trend results

@@
 
-Claus: 
-- run regression with time trend
-- run regression with time trend interacted with wave 1 household characteristics
+Claus: run regression with time trend interacted with wave 1 household characteristics
+
+[need something on weather here?]
+
+It is, indeed, concerning that there are so substantial
+differences in the proportion of households that experienced a 
+shocks over the different rounds, although it is possible that
+this simply is a stark reminder of the volatile nature of life
+in a developing country when dependent on rain-fed agriculture.
+We have added an additional table to the Appendix to address this
+concern. In addition to your suggestion of interacting a linear
+time trend with wave 1 characteristics (assets per capita,
+age group dummies, and education group dummies), we also present
+a separate model where wave dummies are interacted with the
+same set of wave 1 characteristics.
+The small differences in point estimate and level of significance 
+indicate that our results are not driven by differential trends
+in which types of households that experience crop loss.
+
+[Add something on this in paper?]
 

### 2017-09-19 response/response-jde-1.tex
Expand discussion of postponement vs limitation of fertility

@@
 
-[difficult because the short period of time the data cover; only
-really cover 32 months]
+It is correct that the current fixed effects model cannot tell
+us whether we observe postponement of fertility or a permanent
+reduction.
+We focus on postponement for two reasons.
+
+First, active postponement is a necessary---but not sufficient---condition
+for permanent fertility reduction.
+If we cannot show a conscious decision to delay fertility, it is 
+hard to argue that we can ever find evidence of conscious permanent 
+reduction in fertility.
+
+Second, it is not clear that shocks---in and off themselves---should
+substantially affect completed fertility. 
+We discuss evidence from Guatemala that the overall \emph{risk} of 
+experiencing a shock is likely to affect fertility, but that only
+shocks that occur late in a woman's reproductive life cycle is
+likely to affect completed fertility (because the opportunities to
+make up for the ``missed'' birth are fewer and because the marginal
+benefit of a child is smaller the more children you already have and
+the older you are).
+Hence, an important distinction is that long term fertility is likely
+affected by the overall \emph{risk}, whereas the exact timing of
+fertility is driven by a combination of risk and shocks (probably
+mostly shocks).
+
+The longitudinal nature of the data is important for our ability
+to control for unobservable village, household, and individual 
+characteristics, but the trade-off is that the data only cover 
+32 months and that is too short to analyze long-term fertility
+outcomes.
 
@@
 each time we include an additional lag.
-
-Important distinction is that long term fertility is likely
-affected by the overall \emph{risk}, whereas the exact timing of
-fertility is driven by a combination of risk and shocks (probably
-mostly shocks).
 

### 2017-09-22 response/response-jde-1.tex
Add information on rainfall and discuss timing of marriage

@@
 
-[RESPONSE:]
+[RESPONSE:] That is, indeed, a very interesting question but we have 
+decided not to pursue it at the moment for three reasons.
+First, there are already two papers that examine this question, 
+albeit for other countries \citet{mbiti2008,Hoogeveen2011}.
+Second, as those two papers show, the length of our paper would have
+to increase substantially to do this question justice, and the Editor
+has asked us to not increase the length of the paper.
+Finally, the sample size is, indeed, on the small size to really get 
+at this question.
+
+We may return to this question in a separate paper.
+A problem with the current literature on timing of marriage
+is that it tend to focus on only the female side of the marriage market.
+A more satisfactory model will have to consistently explain both sides 
+of the market.
+Hence, results that show, for example, that women are less likely to
+marry after a crop loss would have to show the same for men.
 
@@
 
+- there is no 30 year average to compare the observed district level
+rainfall data to.
+
 \item In the regression specifications, the authors express the survey
@@
 
-[RESPONSE:]
+[RESPONSE:] [Rainfall data analysis]
+
+- no live stock shock question
+- household member death
+- price indices
+- illness of mother and other household members
 
@@
 
-[RESPONSE:] Although we appreciate the concern about endogeneity, 
+[RESPONSE:] Although we understand the concern about endogeneity, 
 we believe the table is worth keeping. 
@@
 
+Weather is likely an important factor, but unfortunately rainfall data
+are only available at district level (there are five districts in the
+data) and even that has missing information.
+
 \item Trend

### 2017-09-26 response/response-jde-1.tex
Add response to R1's question on abstinence
Question was whether abstinence could be recorded as a contraceptive
strategy even if it was caused by other factors, such as stress.
Added table to contraceptive effectiveness file that subdivides
contraceptive use so abstinence only is a separate category,
and updated Makefile and main paper with this new appendix table.

Fixed the definitions in exclude abstinence file.

Wrote initial response based on new results to argue that if
stress could cause this effect we should see something similar
with illness but there is no evidence of that.

@@
 
-\begin{description}
+\begin{enumerate}
 
@@
 
+[Need to rewrite the section to cover both work and illness]
+[A-12 for any, except abstinence only, is very close to statistically significant]
+
+We assume that the reference to ``modern'' is a typo and that the
+first sentence is meant to refer to traditional contraceptives.
+To be clear, we argue that both modern and traditional contraceptive
+use is intentional.
+It is correct that we cannot directly rule out that some women will 
+report abstinence as a contraceptive strategy even if the abstinence 
+arise from something other than a desire to limit the risk of pregnancy.
+It is important, however, to note that the question was specifically
+about contraception.
+
+\begin{quote}
+Some couples use contraception methods to avoid pregnancy or to
+space births. Are you currently using a method of contraception?
+(For example, the pill, the IUD, condom, withdrawal, rhythm,
+abstinence.)
+\end{quote}
+
+Although we cannot measure directly measure stress, we do have information 
+on illnesses and their durations.
+If stress is an important factor in reporting abstinence, we should
+expect the same to be the case with illness, especially those that 
+prevent a woman from working.
+Hence, although we did not explicitly cover this possibility in the original 
+version, coital frequency may also decrease because of illness.
+% We would expect illness to go up both because of the direct effect
+% on the household and from increased stress.
+We show the effect of crop loss on illness in Table 5.
+Crop loss is  not associated with substantial enough changes in BMI or 
+self-reported illness for the women in the sample to explain our fertility results.
+Furthermore, new regression we ran show that there is only a very weak 
+positive association between having been ill for two or more days in the last 
+6 months and the likelihood of reporting using abstinence as a contraceptive 
+strategy.
+This relationship becomes only marginally stronger, and never anywhere near
+statistically significant, if defining illness as not being able to
+perform usual activities for 7 days or more.
+Going above 7 days makes the relationship negative, most likely because
+very few women report being unable to perform regular activities for that
+long.
+
+To further investigate the role of abstinence we added a new
+table to the Appendix, which further subdivides the results in Table 6.
+Table A-13 [TK] show, not surprisingly, that abstinence is the most 
+effective traditional contraception, although only in combination 
+with a crop loss as in the table in the main text.
+A major caveat is that only 19 women report using abstinence only
+across the three waves used for these regressions.
+Modern contraceptive also are mainly effective in combination with
+a crop loss, although the direct effect of using modern contraceptives
+is not statistically significant.
+
+[TK summing up]
+
+% Reduction in pregnancy is 0.3 for abstinence with crop loss
+% Reduction in pregnancy is 0.2 for modern contraceptives with crop loss
+
 \item The authors analyze the fertility of women who were married at the
@@
 
-[RESPONSE:] That is, indeed, a very interesting question but we have 
+[RESPONSE:] That is, indeed, a very interesting question, but we have 
 decided not to pursue it at the moment for three reasons.
@@
 rainfall data to.
+
+Crop loss question:
+
+QB9: ``Have you lost any part of the harvested crop to insects,
+rodents, fire, rotting, etc.?
 
@@
 
-\end{description}
+\end{enumerate}
 

### 2017-09-26 response/response-jde-1.tex
Work on response on cut-off points for crop loss
Added new file anCut_offs.do that show the five main outcomes for
a variety of cut-offs for crop loss.

Also minor edits of bib file

@@
 
+[run regressions with 0, 100, 200, 300, 400, 500, 1,000]
+
+[In many ways, the results look even better if we use the log version
+instead of the dummy approach, although it makes it more difficult
+to interpret.]
+
+[We presented this paper a number of times and whether we led with
+our dummy results or with the log results the discussant invariably 
+asked to show the other set of results.
+We settled on the dummy results before submission because they are
+easier to interpret for most people.
+The downside is, of course, that the cut-off used is subject to
+discussion and that there is no one right number.]
+
 \item The authors use causal language throughout the paper, describing

### 2017-09-26 response/response-jde-1.tex
Part response to dummy cut-off question

--- 2017-09-26 response/response-jde-1.tex
@@
 
+We presented this paper a number of times before submission and whether 
+we led with our dummy results or with the log results the discussant 
+invariably asked us to focus on the other set of results.
+We settled on the dummy results because they are easier to interpret 
+for most people, but the downside is, as you point out, that the 
+cut-off used is subject to discussion and that there is not one right level.
+
+In many ways, the log specification in Table A-8 TK is more theoretically 
+appealing because it captures that we should expect shocks to have
+a diminishing impact on fertility related outcome as the size of the
+crop loss increase (which is also why the linear version is not 
+very appealing; it assumes that a one TZS increase in crop loss has
+the same effect whether it is at 0 or 50,000 TZS). 
+Furthermore, the conclusion one would draw from the log results
+is exactly the same as what we draw from the dummy approach.
+Crop losses lead to statistically significant reductions in the likelihoods 
+of pregnancy and birth with crop losses, and statistically significant 
+increases in the likelihood of using contraceptives overall and
+for the use of traditional contraceptives.
+The difference from the dummy result is that the pregnancy and
+birth results are slightly less statistically significant, whereas
+the contraceptives results are slightly more statistically significant.
+The main downside is that the log version is more difficult to interpret.
+
+To show that the cut-off chosen does not have substantial effects on
+the conclusions drawn we have added a table that shows the main results
+for different cut-offs and the associated percentages of observations
+that would be counted as having a crop loss.
+
+[should we drop those between 0 and the crop loss cut-off?]
+
 \item The authors use causal language throughout the paper, describing

### 2017-09-27 response/response-jde-1.tex
Finish up code and discussion of cut-offs

--- 2017-09-26 response/response-jde-1.tex
@@
 
+We presented this paper a number of times before submission and whether 
+we led with our dummy results or with the log results the discussant 
+invariably asked us to focus on the other set of results.
+We settled on the dummy results because they are easier to interpret 
+for most people, but the downside is, as you point out, that the 
+cut-off used is subject to discussion and that there is not one right level.
+
+In many ways, the log specification in Table A-8 TK is more theoretically 
+appealing because it captures that we should expect a given increase
+in crop loss to have a diminishing impact on fertility related outcomes 
+as the size of the crop loss increase (which is also why the linear version is 
+not appealing; it assumes that a one TZS increase in crop loss has
+the same effect whether it is at 0 or 50,000 TZS). 
+Furthermore, the conclusion one would draw from the log results
+is exactly the same as what we draw from the dummy approach.
+Crop losses lead to statistically significant reductions in the likelihoods 
+of pregnancy and birth with crop losses, and statistically significant 
+increases in the likelihood of using contraceptives overall and
+for the use of traditional contraceptives.
+The differences from the dummy results are that the pregnancy and
+birth results are slightly less statistically significant, whereas
+the contraceptives results are slightly more statistically significant.
+The main downside is that the log version is more difficult to interpret.
+
+To show that the specific cut-off chosen does not have substantial effects on
+the conclusions drawn we have added Table A-10 TK, which shows the main results
+for different cut-offs together with the associated percentages of observations
+that would be counted as having a crop loss.
+Except for any crop loss (cut-off of above zero) the results
+are consistent across all other specifications.
+As the cut-off becomes very large statistically significance is lower
+for pregnancy and births and eventually also for contraceptive use.
+This is not surprising.
+As we increase the cut-off there are more and more households 
+that have experienced a crop loss but are not counted as having a
+crop loss, which will tend to bias results towards zero if there,
+indeed, is an effect of crop loss (this is confirmed by regressions---not
+shown---where we drop all observations with crop loss larger than zero but
+smaller or equal to the cut-off and find substantially large coefficients).
+
 \item The authors use causal language throughout the paper, describing

### 2017-10-03 response/response-jde-1.tex
Response to R2 on weather and crop loss
Added references on growing season in Kagera to bib file.

Made table for crop loss and average monthly rainfall and added
table to paper. Wrote response to R2.

@@
 
-Shamma: look for data on rainfall and papers that have used rainfall
-for Kagera data
-
-- problem is that we do not know when the crop loss occurred within
-the survey period (which is 12 months for the first round and therefore
-very long).
-
-- there is no 30 year average to compare the observed district level
-rainfall data to.
-
-Crop loss question:
-
-QB9: ``Have you lost any part of the harvested crop to insects,
-rodents, fire, rotting, etc.?
-
 \item In the regression specifications, the authors express the survey
@@
 
-- no live stock shock question
+- live stock shock question
 - household member death
@@
 - illness of mother and other household members
+- rainfall does not do anything
 
@@
 
-Weather is likely an important factor, but unfortunately rainfall data
-are only available at district level (there are five districts in the
-data) and even that has missing information.
+[we need to tone down the causality part; explain the growing season and
+the crop loss question]
+
+The question ask (Question 11.B.9): ``Have you lost any part of the 
+harvested crop to insects, rodents, fire, rotting, etc.?''
+It is correct that the crop loss is subjectively measured and for
+most of the suggested reasons it is not possible to verify what
+role they played in the crop loss.
+
+As you suggest, weather is a potential factor.
+Unfortunately the only available weather data consist of monthly rainfall 
+at district level (there are five districts in the data) with missing 
+information for some months.
+Furthermore, there is no 30 year average to compare the observed 
+district level rainfall data to.
+This means that it is not possible to reliably separate out positive 
+(above mean) and negative (below mean) rainfall shocks.
+Although community or women fixed effects will partly ameliorate the
+problem of missing average data, we cannot rule out the possibility
+that all observed periods deviated substantially from the average in 
+the same direction in which case we will fail to capture that the observed
+rainfalls were all ``shocks''.
+
+A number of different specifications of rainfall's association with 
+crop loss is possible.
+Kagera has two growing seasons, February through May and
+October through December \citet[p 6]{Tanzania2007}.
+We tried two different strategies: 
+one where rainfall is measured as the average monthly rainfall 
+over a set period prior to the survey month; 
+and one based on average rainfall for growing season(s) preceding 
+the survey month 
+(that is, not including the current growing season if the survey
+took place during a growing season).
+Despite that rainfall is measured at district level either strategy
+lead to variation across communities because of differences in 
+when each community was surveyed.
+
+The first strategy did not provide any clear results for a variety
+of period lengths and we therefore focus on the growing season approach.
+Table TK shows individual level fixed effects estimates of the 
+effects of average monthly rainfall in centimeters on the likelihood
+of experiencing a crop loss over the last using prior growing season 
+in one specification and one specification with prior growing for
+waves 2 through 4 and the last two prior growing seasons for wave 1.
+Higher rainfall is associated with a lower likelihood of a crop loss. 
+A one standard deviation change in rainfall is associated with an
+approximately 6 percentage point change in the probability of 
+experiencing a crop loss.
+We also tried adding square terms, but the estimated coefficients
+were small and far from statistically significant.
+The same is the case for using crop loss per capita as the dependent
+variable.
+
+It is worthwhile noting the relatively small difference in the 
+estimated effect of rainfall between the two models, 
+although the coefficient on rainfall in Model I is just outside
+the standard levels of statistically significance.
+This suggests that the high level of crop loss observed in
+wave 1 (see discussion below as well) is not mainly due to 
+rainfall shocks that occurred longer occur than for the other waves.
+
+In both models we include a dummy for missing information on rainfall 
+for any month in the period / district combination.
+If the raindata has one or more missing values for the months
+covered in the growing season there is a large and statistically 
+significant increase in the probability of reporting a crop loss
+(dropping observations with missing information does little
+to change the coefficients on rainfall, but do make them
+statistically insignificant).
+It is, however, very difficult to establish why this is because
+no additional information in provided with the data.
+Furthermore, the missing information problem only affects
+observations in waves 2 and 3.
+
+We also tried including rainfall---as defined above---as 
+additional variables in the main regressions.
+Consistent with the small effects of rainfall on crop loss
+there are no apparent effects of the rainfall measures on the
+outcomes and little to no change in the coefficients for crop
+loss on our five main outcomes.
+
+We realize that this answer does fully explain where the
+crop losses come from but unfortunately there is not much
+more relevant information that we can find beside the
+rainfall data.
+[Expand discussion of these issues in paper]
+
+% - problem is that we do not know when the crop loss occurred within
+% the survey period (which is 12 months for the first round and therefore
+% very long).
 

### 2017-10-04 response/response-jde-1.tex
Response to R1 on time varying variables
Tried price index and livestock loss. Neither changed the results.

Added user guide to bib file to use for response.

@@
 
-[RESPONSE:] [Rainfall data analysis]
+[RESPONSE:]
 
-- live stock shock question
-- household member death
-- price indices
-- illness of mother and other household members
-- rainfall does not do anything
+There are three time-varying covariates in the survey that could 
+potentially affect crop loss or the outcome variables directly:
+rainfall, prices, and loss of livestock.
+Other time-varying variables are illness and health variables and 
+we already show that these are unlikely to explain the results we find.
+
+As we discuss in detail below in response to Reviewer 2, higher 
+rainfall is associated with lower risk of crop loss, although there
+are a number of caveats to that result.
+Furthermore, including rainfall as an independent covariate does 
+not change the relationship between crop loss and the five main outcomes.
+
+In Table TK we show the main results when including a dummy for livestock
+loss and the price index included with the Kagera data.
+Although both new variables generally have the expected sign, neither are 
+close to statistically significant and there is only minimal change in the 
+estimated coefficients for crop loss.
+There are two caveats to the results.
+First, the loss of livestock question is not as ``clean'' as the
+crop loss question in that the livestock question also includes 
+any livestock given away by the household and that is not indication which
+proportion falls into actual loss and which proportion is gifts.
+Second, the prices used for the price index are measured at the time
+of the survey.
+We tried lagging the price index, but it does not change the findings.
+
+% - rainfall does not do anything
+% - illness of mother and other household members
 
@@
 It is correct that the crop loss is subjectively measured and for
-most of the suggested reasons it is not possible to verify what
-role they played in the crop loss.
+most of the listed reasons for crop loss it is not possible to verify 
+what role they played.
 
@@
 Kagera has two growing seasons, February through May and
-October through December \citet[p 6]{Tanzania2007}.
+October through December \citep[p 6]{Tanzania2007}.
 We tried two different strategies: 

### 2017-10-04 paper/incomeShocks-jde-r1.tex
Update month on front page
Add postponement results to appendix
Add postponement and time trend results to Appendix
Add response to R1's question on abstinence
Question was whether abstinence could be recorded as a contraceptive
strategy even if it was caused by other factors, such as stress.
Added table to contraceptive effectiveness file that subdivides
contraceptive use so abstinence only is a separate category,
and updated Makefile and main paper with this new appendix table.

Fixed the definitions in exclude abstinence file.

Wrote initial response based on new results to argue that if
st

@@
 \usepackage{xfrac}
+
+% Checking for unused references/labels - comment out before final run
+\usepackage{refcheck}
 
@@
 
-\date{July 2017}
+\date{September 2017}
 
@@
 Furthermore, our results are also qualitatively similar if we use a log of the amount of 
-crop lost or a linear specification of crop loss instead of the dummy variable.
+crop lost or a linear specification of crop loss instead of the dummy variable.%
 \footnote{
@@
 
+% Different cut-offs 
+\input{../tables/appendix_cut_offs.tex}
+
 % Age groups and crop loss effects
@@
 
+% Effectiveness of contraceptives by method
+\input{../tables/appendix_abstinence.tex}
+
 % Effects of Crop Loss on Contraceptives Use by Assets
@@
 
+% Postponement or limiting of fertility
+\input{../tables/appendix_postpone_birth.tex}
+
+\input{../tables/appendix_postpone_contraceptives.tex}
+
+% Time trend and wave dummies interacted with wave 1 characteristics
+\input{../tables/appendix_timetrend.tex}
+
+% Does rainfall explain crop losses?
+\input{../tables/appendix_rainfall.tex}
+
+% Time-varying variables
+\input{../tables/appendix_timevarying.tex}
+
 \end{document}

### 2017-10-05 response/response-jde-1.tex
Minor edits

@@
 rainfall, prices, and loss of livestock.
-Other time-varying variables are illness and health variables and 
+Other time-varying variables are illness and health variables, which 
 we already show that these are unlikely to explain the results we find.
@@
 The results are broadly consistent with what we originally found, 
-although substantially less likely to be statistically significant 
+although less likely to be statistically significant 
 because of the lower variation in crop loss and the smaller sample
@@
 
+We already include a version of this in the paper.
+Table A-17 TK include the fraction of households in the community 
+who have experienced a crop (excluding the household itself).
+The sign of the village level crop loss are generally the same as
+those for the household crop loss, although---with the exception
+of traditional contraceptives---the village level variable is not
+statistically significant.
+Furthermore, the crop loss coefficients decline only slightly
+when including village level crop loss indicating that there is
+only relatively mild spill-over between households for this
+particular type of shock.
+
 \item Births analysis

### 2017-10-07 response/response-jde-1.tex
Respond to R1 on historical fertility and traditional contraceptives

@@
 \usepackage{rotating}  
-\usepackage{amsmath}
-\usepackage{caption} 
 \usepackage{dcolumn} 
@@
 to make it easier to follow and eventually publish on GitHub.
-This rewrite slightly lowered the sample used by three observations.
+This rewrite lowered the sample by three observations.
 Including all married/partnered women up to age 45, rather than exclude
@@
 
-[Also include Cinnirella since that covers a similar topic]
+We have edited the introduction to highlight the high use of traditional 
+contraception and that most of the increase in contraception after a 
+shock comes from traditional methods.
+We have also added an additional point to the discussion of the
+main contributions of the paper.
+In addition to the \citet{Guinnane2011} paper, we also cite a just
+published paper on spacing in pre-fertility transition England
+\citep{Cinnirella2017} and \citet{Michael1976}, \citet{David1986},
+and \citet{Santow1995}
+on the role of different types of traditional contraception in
+controlling fertility before the emergence of modern contraception.
+
+In the Conclusion we already discussed that our results were
+especially of interest because of the composition of contraceptive
+methods used.
+We have edited the discussion to further clarify the major role
+of traditional contraceptives.
+In addition, we have recast the second to last paragraph to
+highlight the connection to long-term fertility and the literature
+on fertility transition in Europe and the US.
+
+% [should the footnote on historical control of fertility be moved up
+% to why this is important?]
+
+% [few data sets with panel data on contraceptive use and economic
+% circumstances]
 
@@
 First, there are already two papers that examine this question, 
-albeit for other countries \citet{mbiti2008,Hoogeveen2011}.
+albeit for other countries \citet{mbiti08,Hoogeveen2011}.
 Second, as those two papers show, the length of our paper would have

### 2017-10-07 paper/incomeShocks-jde-r1.tex
Remove map of Kagera
Respond to R1 on historical fertility and traditional contraceptives

@@
 received little attention in the literature.
-There is evidence that shocks can affect the timing of births, but it is unclear
-whether the cause is households' intentional planning or unintended consequences 
+A correlation between shocks and the timing of births is well-established, but it is 
+unclear whether the cause is households' intentional planning or unintended consequences 
 of the shocks \citep{pitt98b,lindstrom99,Evans2010,Portner2014}.%
 \footnote{
-There is also a related literature on the effects of weather, especially temperature,
+There is a related literature on the effects of weather, especially temperature,
 on timing of fertility \citep[see, for example,][]{Barreca2015}.
@@
 }
-Furthermore, standard models of fertility focus mainly on the effects of time
-invariant household and individual characteristics on fertility outcomes, and 
-tend to ignore the dynamic aspects of fertility decisions.%
-\footnote{
-For an early exception see \cite{Newman1984}.
-}
+There is, in general, little evidence on what determines the timing 
+of fertility in developing countries, with most of the literature instead 
+focusing on the number of children born \citep{Portner2018}.
+% [Little evidence in general on timing of fertility in developing countries;
+% most of the research has focused on completed fertility / children-ever-born] 
+% Furthermore, standard models of fertility focus mainly on the effects of time
+% invariant household and individual characteristics on fertility outcomes, and 
+% tend to ignore the dynamic aspects of fertility decisions.%
+% \footnote{
+% For an early exception see \cite{Newman1984}.
+% }
 
 We seek to answer two key questions in this paper.
-First, do income shocks, measured by accidental crop loss in Tanzania, affect timing of 
-fertility?
+First, do income shocks, measured by accidental crop loss in Tanzania, affect the
+timing of fertility?
 Second, do changes in fertility timing, if any, arise from an intentional decision
@@
 
-There are three reasons why a household may want to delay fertility in the
+There are three reasons a household may want to delay fertility in the
 event of a shock.
@@
 of resources is almost certainly negative \citep{Portner2014}.
-Secondly, diverting time away from market work to child care will be even more costly
-if households need to respond to a shock by working more \citep{kochar99};
-although it is also possible that a shock will temporarily lower women's cost 
-of time, making it more attractive to have a birth now.
+Secondly, diverting time away from market work to child care will be even more 
+costly if the household needs to respond to a shock by working more \citep{kochar99}.%
+\footnote{
+It is, however, also possible that a shock could temporarily lower women's 
+cost of time, making it more attractive to have a birth now.
+}
 Finally, the household may realize that children born following a shock 
@@
 not evidence of a conscious decision to change the timing of fertility.
-A shock may lead one or more members of the household to migrate 
-in search of better economic opportunities.
-If either the husband or the wife is gone for extended periods, this 
-will lower the probability of conception.
-Furthermore, a shock may reduce intercourse frequency---either
-from psychological effects or from increased hours of work.
-We would also observe a decline in births if people delay marriage or if there
-is an increase in the dissolution of marriage.
+A shock may, for example, lead one or more members of the household to 
+migrate in search of better economic opportunities, and if either the husband 
+or the wife are gone for extended periods, this will lower the probability of 
+conception.
+Furthermore, a shock may reduce intercourse frequency---either for
+physiological or psychological reasons or from increased hours of work.
+We would also observe a decline in births if people delay marriage or 
+if there is an increase in the dissolution of marriage.
 Finally, severe shocks can lead to starvation and health problems \citep{lindstrom99}.
-Malnourished pregnant mothers are more likely to have stillbirths,%
+Malnourished pregnant mothers are more likely to have stillbirths,
+and non-pregnant women may see an increased incidence of secondary sterility
+from a reduction in age of menopause or famine induced amenorrhea.%
 \footnote{
@@
 }
-and non-pregnant women may see an increased incidence of secondary sterility
-from a reduction in age of menopause or famine induced amenorrhea.
 
@@
 % This could be better explained
-Finally, parents do not directly control fertility and its timing, but rather 
-control type and intensiveness of contraceptive use and the frequency and timing of 
-sexual activity.
-Hence, it helps us understand if, and how, households regulate their fertility.
+Finally, it helps us understand if, and how, households regulate their fertility.
+% parents do not directly control fertility and its timing, but rather 
+% control type and intensiveness of contraceptive use and the frequency and timing of 
+% sexual activity.
 This, in turn, can inform policies on availability and targeting
@@
 
-Most of the prior literature on the impact of economic and other shocks on fertility has
-used historic data on what are now developed countries.
+Most of the prior literature on the impact of economic and other shocks on fertility 
+has used historic data on what are now developed countries.
 Data from Rouen, France, over the period 1681-1787, show that increases in the
 price of wheat led to a dramatic fertility decline for the urban poor, 
-but fertility of the urban wealthy was unaffected \citep{Galloway1987}.
+whereas fertility of the urban wealthy was unaffected \citep{Galloway1987}.
 Similarly, English data from 1542 to 1800 show that mortality shocks led to short-term
@@
 }
-
-In pre-industrial Sweden, a positive crop shock, positive weather shock, and positive wage 
-shock increased fertility for a five- to ten-year period following the shocks \citep{Eckstein1985}.
-However, the increase in birth rate changes only the timing of births and has no cumulative 
-effect on long-term fertility rate.
-Similarly, the increasing price of rye in Southern Sweden led landless and semi-landless 
-families to reduce their fertility, whereas landed farmers showed no effect 
-\citep{bengtsson06}.
-
-Using present-day data, there is evidence that famine and domestic/regional military
-attacks in Ethiopia led to short-term decreases in the likelihood of conception
-\citep{lindstrom99}.
+In pre-industrial Sweden, positive crop shock, positive weather shock, and 
+positive wage shock all increased fertility for a five- to ten-year period following 
+the shocks \citep{Eckstein1985}.
+However, the increase in birth rate changed only the timing of births with no 
+cumulative effect on the long-term fertility rate.
+Similarly, the increasing price of rye in Southern Sweden led landless and 
+semi-landless families to reduce their fertility, whereas landed farmers 
+showed no change \citep{bengtsson06}.
+
+Using present-day data, there is evidence that famine and domestic/regional 
+military attacks in Ethiopia led to short-term decreases in the likelihood of 
+conception \citep{lindstrom99}.
 Closer to our paper, positive rainfall shocks in Senegal increased 
@@
 
-Many studies examine how a variety of factors affect contraception use, such 
+Many studies show how a variety of factors affect contraception use, such 
 as, 
@@
 contraception use \citep{Hernandez-Correa2010}. 
-Cross-sectional data from Madagascar show that households that experience adverse events 
-are more likely to use contraception.\footnote{%
+Cross-sectional data from Madagascar show that households that experience adverse 
+events are more likely to use contraception.\footnote{%
 Adverse events include economic and environmental aspect, such as: rise in input 
@@
 } 
-However, households may be able to anticipate many of these events before they 
-occur and adjust their behavior accordingly.
-Combined with possible correlation between the probability of
-experiencing an adverse event and unobserved individual and household
-characteristics, this may bias the estimates, and the cross-sectional
-data allows for neither household nor individual level fixed effects
-estimates.
-
-We use longitudinal data from the Kagera region in Tanzania with information on accidental 
-crop loss, fertility, and contraceptive use to estimate individual level fixed 
-effects models of the effects of income shocks on both timing of fertility and 
-contraceptive use.
+However, 
+% households may be able to anticipate many of these events before 
+% they occur and adjust their behavior accordingly.
+% Furthermore, 
+the possible correlation between the probability of experiencing 
+an adverse event and unobserved individual and household characteristics may 
+bias the estimates, and the cross-sectional data allows for neither household 
+nor individual level fixed effects estimates.
+
+We use longitudinal data from the Kagera region in Tanzania with information on 
+accidental crop loss, fertility, and contraceptive use to estimate individual 
+level fixed effects models of the effects of income shocks on both timing of 
+fertility and contraceptive use.
 We show that income shocks decrease the likelihood of pregnancies and births.
-At the same time, shocks significantly increase use of contraceptives, 
-especially traditional contraceptives.
+At the same time, shocks significantly increase use of contraceptives.
+This increase mainly occurs through the use of traditional contraceptives, 
+which are the dominant form of contraception in the area.
 We argue that the postponement of fertility in households experiencing income 
@@
 
-The paper makes two main contributions to the literature. 
+The paper makes three main contributions to the literature. 
 First, this is the first paper to show that households respond to income shocks 
-through family planning, which, in turn, affects the timing of child bearing. 
-Our study provides evidence that the postponement of fertility occurs through a 
-planned decision process rather than as an unplanned consequence.
-Secondly, the paper contributes to the consumption smoothing literature as households 
+through contraceptive use, which, in turn, affects the timing of child bearing. 
+% Our study provides evidence that the postponement of fertility occurs through a 
+% planned decision process rather than as an unplanned consequence.
+Second, with low rates of modern contraceptive use in the area, the paper contributes 
+to the literature on how populations control fertility before widespread adoption of
+modern contraception.%
+\footnote{
+There has long been a discussion of when and how families 
+undertook fertility control before the emergence of modern 
+contraceptives \citep{Guinnane2011,Cinnirella2017}.
+Simulations show that it clearly is possible to substantially limit fertility
+even with only access to traditional contraceptives, but there is only 
+limited evidence available on exactly which methods were used 
+\citep{Michael1976,David1986,Santow1995}.
+}
+Finally, the paper contributes to the consumption smoothing literature, as households 
 affected by income shock use family planning as a mechanism to smooth their consumption. 
@@
 Finally, households were randomly sampled from the groups.
-For further details on the sample selection, please see World Bank (2004).
+For further details on the sample selection, see World Bank (2004).
 }
@@
 The Kagera region is in the northwestern corner of Tanzania, west of Lake Victoria, and 
-borders Uganda, Rwanda, and Burundi as shown in Figure \ref{fig:kagera_map}. 
+borders Uganda, Rwanda, and Burundi. 
 The population of the area---1.3 million in 1988---is 
@@
 
-\begin{figure}[ht] 
- \centering
- \includegraphics[width=0.6\textwidth]{../staticPDF/kagera.pdf}
- \caption{Map of Kagera region}
-  \label{fig:kagera_map}
-\end{figure}
+% \begin{figure}[ht] 
+%  \centering
+%  \includegraphics[width=0.6\textwidth]{../staticPDF/kagera.pdf}
+%  \caption{Map of Kagera region}
+%   \label{fig:kagera_map}
+% \end{figure}
 
@@
 The TZS 200 cut-off is equivalent to approximately 5th 
-percentile of all household savings.%
+percentile of all household savings.
 % That period is approximately 6 to 7 months for the last three surveys, but covers 
@@
 response to a crop loss and that contraceptive use, especially use of traditional 
-methods like abstinence and rhythm, significantly increases.
+methods, such as abstinence and rhythm, significantly increases.
 
@@
 
-Based on this, our main conclusion is that households consciously delay 
-fertility through adjustment in contraceptive use to cope 
-with income shocks. 
-This is especially of interest since there is little use of modern
-contraceptives in this setting.
+Our main conclusion is that households consciously delay 
+fertility through contraceptive use to cope with income shocks. 
+This is especially of interest since this is accomplished 
+predominately through the use of traditional contraceptives.
 If traditional contraceptives have low efficacy, a natural question is 
@@
 
-The data unfortunately do not allow us to establish whether there 
-are long-term impacts on fertility outcomes.
+The data unfortunately do not allow us to establish whether shocks
+have long-term impacts on fertility outcomes.
 Results from Guatemala indicate that only shocks that occur 
 toward the end of a woman's reproductive life is likely to have an 
-impact on final fertility \citep{Portner2014}.
+impact on final fertility because the ability to ``make up'' 
+postponed births are smaller \citep{Portner2014}.
 There is, therefore, little reason to believe that temporary shocks,
-like the crop losses studied here, lead to over-all changes in fertility.
-
-Hence, an important unanswered question is what the implications
-are for our understanding of people's ability to control long-term
-fertility when there is limited access to modern contraceptives.
-As we have shown, it is possible for households to, at least temporarily, 
-limit fertility using traditional methods once the incentives are
-strong enough.
-What we do not know is whether the psychic cost of using traditional 
-contraceptives is so high that although households are willing to use 
-them to deal with short-term shocks, they become difficult to use for 
-long-run moderation of fertility.
-That studies of the introduction of modern family planning, using individual
-level data, rarely find reductions in fertility larger than one child
-is an indication that households were already able to limit long-term 
-fertility using traditional contraceptives.%
+like the crop losses studied here, lead to over-all changes in fertility,
+although the underlying \emph{risk} of shocks may.
+
+[Lead with this or the next sentence?]
+Our result that people are able to postpone fertility using mainly 
+traditional contraceptive do, however, have implications for our 
+understanding of long-term fertility.
+[Alternative lead]
+What implications do our results that people are able to postpone
+fertility using mainly traditional contraceptive have for our
+understanding of long-term fertility?
+% What we do not know is whether the psychic cost of using traditional 
+% contraceptives is so high that although households are willing to use 
+% them to deal with short-term shocks, they become difficult to use for 
+% long-run moderation of fertility.
+The economic literature on fertility transition in Europe and the US 
+show a substantial decline in fertility and control of the timing of 
+fertility before the introduction of modern contraceptives, although 
+the methods used are generally not well documented.%
+\footnote{
+See, for example, \citet{Guinnane2011} on fertility reduction and 
+\citet{Cinnirella2017} on spacing of birth.
+On contraceptive methods used see, for example, \citet{Michael1976},
+\citet{David1986}, and \citet{Santow1995}.
+}
+Furthermore, studies of the introduction of modern family planning, 
+using individual level data, rarely find reductions in fertility larger 
+than one child.
+Both set of results indicate that households are able to 
+limit long-term fertility using traditional contraceptives.%
 \footnote{
@@
 }
-
-This does not mean that the ability to postpone birth is unimportant
-or that there is not an important role for modern contraceptives.
+Our individual level data allow us to directly link the use of
+traditional contraceptives and fertility outcomes.
+We argue that, contrary to common perception, traditional
+contraceptive methods have a high degree of efficacy when the
+incentives to prevent pregnancy or birth are sufficiently high.
+Because of the small data set we cannot, however, differentiate between
+the different traditional methods, although---contrary to Europe 
+and the US---withdrawal clearly plays little or no role in this area.
+
+Finally, our focus on traditional contraceptives does
+not indicate that access to modern contraceptives is unimportant.
+Although use of modern contraceptives is low in the area we show
+they are also effective in preventing pregnancies and births after a shock.
 Children born after shocks are likely to fare worse than children born 
@@
 to recover and care for the new child.
-Our results imply that providing better access to contraceptives 
-during times of economic shocks could improve the ability of households 
+Our results imply that providing better access to modern contraceptives 
+during times of economic shocks may improve the ability of households 
 to postpone births.

### 2017-10-09 response/response-jde-1.tex
Edits of paper and reviewer response

@@
 [Additional discussion of overall things we have changed---if any]
+- move discussion of education, wealth, and education results to appendix
+to save space
 
@@
 We list below our responses to the individual comments and suggestions.
+
+[Possible changes]
+If the paper is too long, one possibility is to move the discussion of
+age, education, and wealth descriptive statistics to the Appendix.
+We only need this for the OLS and for the discussion of those factors,
+which is now in the Appendix.
 
@@
 outcomes.
+We followed your suggestion of adding an additional lag to 
+the model and discuss the results in a new section in the Appendix.
+There are, however, a number of issues with this approach.
+First, it is possible that the effect of a crop loss is felt
+in subsequent period and not only in the period in which it
+was observed.
+Second, we lose 1/4 of observations each time we include an 
+additional lag, which leaves us with less variation because
+the first round had a higher proportion of crop losses than
+the subsequent rounds
 
-Claus: run main regression with longer lags to check if any effects
-on pregnancy and birth
-
-Claus: write something on Guatemala paper and how to think about how
-long it takes to find an effect / recover from a shock
-
-There are two competing factors here.
+Furthermore, even if shocks only had an impact in one period, 
+there are two competing factors at work here.
 First, a negative shocks will make people more likely to postpone 
@@
 paper.
-Second, longer spacing between births is costly because economies
-of scale are harder to realize and because it means that the mother
-will be out of the labor force for longer.
+Second, longer spacing between births can be costly because economies
+of scale are harder to realize and because the mother may have to 
+be out of the labor force for longer if she already have young
+children to care for.
+We have therefore also added an interaction between crop loss for 
+different periods with the idea that there is a cost to continually 
+postponing births.
 
-This might explain why there are negative effects for both shock
-and lagged shock, but the interaction between the two is positive.
-
-One problem with the regression is that we lose 1/4 of observations
-each time we include an additional lag.
-
-I have also added an interaction between crop loss for different periods
-with the idea that there is a cost to continually postponing births. If
-that holds, the effect of both shocks on their own should be same and
-the interaction the opposite. Broadly speaking that appears to hold. The
-upshot–as far as I can tell–is that what we observe is postponement and
-not an overall limitation in fertility. This makes sense given my
-results for Guatemala, where long-term risk affects the desire fertility
-*level* whereas shocks experienced affect the *timing* of fertility. I
-will try to write something more up on this.
+If crop losses last more than one period and there is a cost 
+associated with further postponement, the effect of both shocks 
+on their own should be same and the interaction the opposite. 
+Broadly speaking that appears to hold. 
+The upshot---as far as we can tell---is that what we observe is 
+postponement and not an overall limitation in fertility. 
+This makes sense given results for Guatemala, where long-term risk 
+affects the desire fertility \emph{level}, whereas shocks experienced 
+affect the \emph{timing} of fertility. 
+We have added a discussion of this in the Conclusion.
 
@@
 [RESPONSE:]
-
-[Need to rewrite the section to cover both work and illness]
-[A-12 for any, except abstinence only, is very close to statistically significant]
 
@@
 version, coital frequency may also decrease because of illness.
+We have added that possibility to the text.
+
 % We would expect illness to go up both because of the direct effect
 % on the household and from increased stress.
+
 We show the effect of crop loss on illness in Table 5.
@@
 self-reported illness for the women in the sample to explain our fertility results.
-Furthermore, new regression we ran show that there is only a very weak 
-positive association between having been ill for two or more days in the last 
-6 months and the likelihood of reporting using abstinence as a contraceptive 
-strategy.
+Furthermore, we ran regressions (not included in the paper) that show that there 
+is only a very weak positive association between having been ill for two or more 
+days in the last 6 months and the likelihood of reporting using abstinence 
+as a contraceptive strategy.
 This relationship becomes only marginally stronger, and never anywhere near
@@
 Modern contraceptive also are mainly effective in combination with
-a crop loss, although the direct effect of using modern contraceptives
+a crop loss, and the direct effect of using modern contraceptives
 is not statistically significant.
-
-[TK summing up]
+We discuss these results at the end of the section on 
+``Conscious Decision or Unintended Consequences?''.
 
@@
 
+Our original discussion of this issue was scattered throughout the
+original version of the paper.
+We have now combined and expanded the discussion in the
+Robustness section.
+Before we respond, it is worth pointing out why we present the dummy
+cut-off results in the paper.
 We presented this paper a number of times before submission and whether 
@@
 
-In many ways, the log specification in Table A-8 TK is more theoretically 
+In many ways, the log specification in Table A- TK is more theoretically 
 appealing because it captures that we should expect a given increase
@@
 To show that the specific cut-off chosen does not have substantial effects on
-the conclusions drawn we have added Table A-10 TK, which shows the main results
+the conclusions drawn we have added Table A-TK, which shows the main results
 for different cut-offs together with the associated percentages of observations
@@
 that have experienced a crop loss but are not counted as having a
-crop loss, which will tend to bias results towards zero if there,
-indeed, is an effect of crop loss (this is confirmed by regressions---not
+crop loss, which will tend to bias results towards zero if there
+is an effect of crop loss for levels lower than the cut-off 
+(this is confirmed by regressions---not
 shown---where we drop all observations with crop loss larger than zero but
-smaller or equal to the cut-off and find substantially large coefficients).
+smaller or equal to the cut-off and find substantially larger coefficients).
 
@@
 
-[see response to R1's q5 on causal (or something like that)]
+[see response to R1's q7 on time-varying covariates]
 
-[we need to tone down the causality part; explain the growing season and
+[TK we need to tone down the causality part; explain the growing season and
 the crop loss question]
 
-The question ask (Question 11.B.9): ``Have you lost any part of the 
-harvested crop to insects, rodents, fire, rotting, etc.?''
-It is correct that the crop loss is subjectively measured and for
-most of the listed reasons for crop loss it is not possible to verify 
-what role they played.
-
+% The question ask (Question 11.B.9): ``Have you lost any part of the 
+% harvested crop to insects, rodents, fire, rotting, etc.?''
+It is correct that the crop loss is subjectively measured and that
+it is not possible to verify what factors played a role in the crop loss.
 As you suggest, weather is a potential factor.
@@
 effects of average monthly rainfall in centimeters on the likelihood
-of experiencing a crop loss over the last using prior growing season 
-in one specification and one specification with prior growing for
-waves 2 through 4 and the last two prior growing seasons for wave 1.
+of experiencing a crop loss.
+We use two specification, one with the last prior growing season and 
+one with prior growing for waves 2 through 4 and the last two prior 
+growing seasons for wave 1.
 Higher rainfall is associated with a lower likelihood of a crop loss. 
@@
 wave 1 (see discussion below as well) is not mainly due to 
-rainfall shocks that occurred longer occur than for the other waves.
+rainfall shocks that occurred longer ago than for the other waves.
 
@@
 for any month in the period / district combination.
-If the raindata has one or more missing values for the months
+If the rain data has one or more missing values for the months
 covered in the growing season there is a large and statistically 
@@
 more relevant information that we can find beside the
-rainfall data.
-[Expand discussion of these issues in paper]
+rainfall data and the variables discussed in our response
+to Reviewer 1.
 
@@
 
-Claus: run regression with time trend interacted with wave 1 household characteristics
-
-[need something on weather here?]
-
-It is, indeed, concerning that there are so substantial
+It is, indeed, potentially concerning that there are so substantial
 differences in the proportion of households that experienced a 
@@
 this simply is a stark reminder of the volatile nature of life
-in a developing country when dependent on rain-fed agriculture.
+in a developing country.
+
 We have added an additional table to the Appendix to address this
-concern. In addition to your suggestion of interacting a linear
+concern and discuss the results in the Robustness section. 
+In addition to your suggestion of interacting a linear
 time trend with wave 1 characteristics (assets per capita,
@@
 in which types of households that experience crop loss.
-
-[Add something on this in paper?]
 
@@
 Table A-17 TK include the fraction of households in the community 
-who have experienced a crop (excluding the household itself).
+who have experienced a crop loss (excluding the household itself).
 The sign of the village level crop loss are generally the same as
@@
 
+The discussion of the longer lag was hidden in the middle of the 
+Estimation Strategy section and the motivation was not clear.
+We have rewritten that part of the Estimation Strategy to better
+explain why the lag is there.
+
 \end{enumerate}

### 2017-10-09 paper/incomeShocks-jde-r1.tex
Edits of paper and reviewer response

@@
 \usepackage[xetex,colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black]{hyperref}
-\usepackage{tikz}
-\usetikzlibrary{decorations}
 \usepackage{xfrac}
@@
 % Checking for unused references/labels - comment out before final run
-\usepackage{refcheck}
+% \usepackage{refcheck}
 
@@
 Lawrence Wu, 
+two anonymous reviewers, 
 and seminar participants at 
@@
 The increase in contraceptive use comes almost entirely from traditional 
-contraceptive methods, such as abstinence, withdrawal, and the rhythm method.
+contraceptive methods, such as abstinence and the rhythm method.
 We argue that these changes in behavior are the result of deliberate decisions of
@@
 
-How shocks affect household decisions on fertility in developing countries has 
+How shocks affect fertility decisions in developing countries has 
 received little attention in the literature.
@@
 }
-There is, in general, little evidence on what determines the timing 
+Furthermore, there is, in general, little evidence on what determines timing 
 of fertility in developing countries, with most of the literature instead 
@@
 showed no change \citep{bengtsson06}.
-
 Using present-day data, there is evidence that famine and domestic/regional 
@@
 even with only access to traditional contraceptives, but there is only 
-limited evidence available on exactly which methods were used 
+limited evidence available on exactly which methods were actually used 
 \citep{Michael1976,David1986,Santow1995}.
@@
 married women, regardless of age, and unmarried women 14 years and older. 
-We restrict our sample to married or partnered women 18-45 years of age who we 
+We restrict our sample to married or partnered women 17-45 years of age who we 
 observe in all 4 rounds of the survey.
@@
 prior survey round (or prior 1 year for the first round).
-% We create a dummy for accidental crop losses greater than 200 TZS,
-% which is our preferred measure of income shock.
+There is substantial variation in the number of households that experience a 
+crop loss between the 4 survey rounds.
+In the 12 months prior to the first survey just below 70\% of households 
+lost crops, whereas 25\% lost crops between the 1st and 2nd surveys.
+That falls even further for the 3rd and 4th rounds, with less than 3\% 
+reporting crop loss of more than TZS 200 for both.
 The TZS 200 cut-off is equivalent to approximately 5th 
 percentile of all household savings.
-% That period is approximately 6 to 7 months for the last three surveys, but covers 
-% the previous 12 months for the first survey. 
-% [MAY HAVE TO EXPLAIN MORE ABOUT WHAT THIS MEANS FOR OUR ESTIMATES]
-There is substantial variation in the number of households that experience a 
-crop loss between the 4 survey rounds.
-In the 12 months prior to the first survey round just over 70\% of households 
-lost crops, whereas just over 25\% lost crops between the 1st and 2nd survey round.
-That falls even further for the 3rd and 4th rounds, with less than 5\% 
-reporting crop loss of more than TZS 200 for both.
-Our results are robust to defining crop loss as any crop loss, with the 
-coefficients slightly smaller and slightly less statistically 
-significant as expected. 
-% Crop losses between the valuation of 0-200 TZS are not coded as an income shock,
-% because of the small magnitude (equal to  of savings). 
-% and unlikely to affect fertility decisions of households. 
-Furthermore, our results are also qualitatively similar if we use a log of the amount of 
-crop lost or a linear specification of crop loss instead of the dummy variable.%
-\footnote{
-Results using these alternative specifications of crop loss are available in the Appendix.
-}
-
-On average across the four survey rounds, 17\% of women report being pregnant 
-at the time of the interviews and 17\% of women report giving birth since the 
+Our results are robust to alternative definitions of crop loss as we
+discuss in Section \ref{sec:robustness}.
+
+On average across the four surveys, 17\% of women report being pregnant 
+at the time of the interviews and 16\% of women report giving birth since the 
 prior survey round.
 In total, just below 70\% of women in the sample report giving birth in
-the covered period (3\sfrac{1}{2} years).
-Similarly, 70\% report that they are currently pregnant at one point during 
-the 4 rounds.
+the covered period (approximately 3\sfrac{1}{2} years).
+Similarly, 70\% report a pregnancy at one point during the 4 rounds.
 
@@
 \footnote{
-The relevant questions are "Some couples use contraception methods to avoid
-pregnancy or to space births. Are you currently using a method of contraception?".
+The relevant question is "Some couples use contraception methods to avoid
+pregnancy or to space births. Are you currently using a method of contraception?"
 If the woman answers yes, she is then asked "What contraceptive method are you
@@
 
-Contraceptive prevalence is low; at the time of the first survey round, 15\% 
+Contraceptive prevalence is low; at the time of the first survey round, 14\% 
 of women report using any type of contraceptive---split between 11 percent that use 
 traditional contraceptives and 4\% that use modern contraceptives.
-In subsequent survey rounds, the contraceptive use rate hovers around 10\%, 
-with 4 to 7\% using traditional contraceptives and 2 to 5\% 
+In subsequent survey rounds, the contraceptive use rate is just below 10\%, 
+with 4 to 6\% using traditional contraceptives and 2 to 5\% 
 using modern contraceptives.
@@
 demographic and household characteristics at wave 1.
-% This survey suffers from two data quality issues: misreporting of 
-% age and misreporting of education level.
 As the time span of the surveys is too short to confidently identify age effects,
@@
 }
-% If her reported age vary more than the length of interval between each survey 
-% round, we use her median reported age rounded down.
 We use respondent's median reported age, rounded down, and 5-year age groups 
@@
 \footnote{
-For many women, the reported education level actually decreases in certain survey 
+For many women, the reported education level decreases in certain survey 
 rounds, or increases by 2 years or more in consecutive survey rounds, which 
@@
 or above (7 or more years of education).
-Only 29\% of the women received no education, 26\% received some 
-primary education, and 45\% completed at least primary education.
+Only 29\% of the women received no education, 27\% received some 
+primary education, and 44\% completed at least primary education.
 The high reported education level is most likely the result of the 1974 
@@
 which increased accessibility of primary education and enrollment rates, 
-even though there are reports that the quality of education was very low.
-In addition, the crisis Tanzania experienced in the 1980s further lowered the 
-quality and enrollments declined significantly.%
-\footnote{See \cite{Galabawa2001} and \cite{Wedgwood2005} for detailed discussions 
+even though there are reports that the quality of education was very low.%
+% In addition, the crisis Tanzania experienced in the 1980s further lowered the 
+% quality and enrollments declined significantly.%
+\footnote{
+See \cite{Galabawa2001} and \cite{Wedgwood2005} for detailed discussions 
 of the development of education in Tanzania.
@@
 
-To capture wealth status of the household, we use total reported assets of the 
+Finally, to capture wealth status of the household, we use total reported assets of the 
 household.
@@
 goods, and savings.
-We measure assets per capita in 10,000 Tanzanian Shilling (TZS).
-
-Finally, the average level of assets in wave 1 was just shy of TZS 80,000.%
-\footnote{%
+The average level of assets in wave 1 was TZS 70,000.%
+\footnote{
 In 1991 the official exchange rate was TZS 200 to the dollar, while it
@@
 There was, however, wide dispersion in asset levels, with the minimum reported 
-wave 1 asset level around TZS 510 and the maximum almost TZS 19,000,000.
-The 25-percentile asset level is around TZS 19,000, median is approximately 
-TZS 35,000, and the 75-percentile asset level is only slightly higher than the 
-mean at TZS 86,000.
-In other words, the asset levels for a majority of households in the area is very low,
-with a few relatively wealthy households.
+wave 1 asset level around TZS 500 and the maximum almost TZS 1,100,000.
+The 25-percentile asset level is below TZS 19,000, median is approximately 
+TZS 34,000, and the 75-percentile asset level is only slightly higher than the 
+mean at TZS 71,500.
+In other words, the asset levels for the majority of households is 
+very low, with a few relatively wealthy households.
 
@@
 non-random attrition.
-In Appendix Tables \ref{tab:attrition_desc_stat_croploss} and \ref{tab:attrition_desc_stat_women}
+In Appendix Tables \ref{tab:attrition_desc_stat_croploss} and 
+\ref{tab:attrition_desc_stat_women}
 we show descriptive statistics for women who fulfilled the requirements for being in the 
 sample in wave 1, but were not included in the main sample.
-A total of 207 women were dropped because they were not observed in all four waves 
+A total of 164 women were dropped because they were not observed in all four waves 
 and/or failed the criteria in at least a one of the later waves, for example 
@@
 \footnote{
-Divorce, separation, and widowhood accounted for 28 women, whereas not observed in all
-4 waves for accounted for 188 women. 
-The two numbers do not add to 207 because some women experienced both a change
+Divorce, separation, and widowhood accounted for 22 women, whereas not observed in all
+4 waves for accounted for 147 women. 
+The two numbers do not add to 169 because some women experienced both a change
 in marital status and were not observed in all 4 waves.
 }
-In each period, women who were dropped from the sample were less likely to have been 
-affected by a shock.
-Furthermore, they were less likely to be pregnant and to have given birth, and,
-correspondingly, a larger percentage use contraceptives.
-Compared to our main sample the women dropped were more likely to be at the ends of
-the included age samples, with 6 percentage points more women in the very youngest
-and the very oldest age groups.
+In each period, women who were dropped from the sample were less likely to have 
+been affected by a shock.
+Furthermore, they were slightly less likely to be pregnant and to have given 
+birth, and, correspondingly, a larger percentage use contraceptives.
+Compared to our main sample the women dropped were more likely to be young,
+with 10 percentage points more in the very youngest age group and 3 
+percentage points at the second youngest age group.
 They were also more educated and had a slightly higher level of asset per capita.
 Hence, it appears that we are not dropping women who were more likely to 
-experience shocks and less likely to cope with them, which in turn suggests that 
-non-random attrition is not a substantial problem.
-However, to further investigate whether there is non-random attrition, we 
+experience shocks and less likely to cope with them.
+However, to further investigate possible non-random attrition, we 
 estimate the association between shocks and partner migration and between
-shocks and dissolution of marriage as described below.
+shocks and dissolution of marriage below.
 
@@
 
-We begin by estimating OLS and community and individual level fixed effects 
-versions of the following equation:
+We estimate OLS and community and individual level fixed effects 
+versions of equation (\ref{eq:fixed_effects1}),
 \begin{align}
@@
 are linear probability models.
+
 For the outcome variables, currently pregnant and contraceptive use, we use 
-the dummy variable for crop loss that occurred in the period immediately 
+a dummy variable for crop loss that occurred in the period immediately 
 prior to the interview, which is 6 to 7 months long, as our main explanatory 
 variable of interest. 
-When we use birth as the outcome variable, 
-we use the dummy variable for crop loss that occurred in the period between 
+When birth is the outcome variable, 
+we use a dummy variable for crop loss that occurred in the period between 
 the prior two survey rounds.%
@@
 }
-
-To further investigate whether the response to shock vary by household
-characteristics, we also estimate individual level fixed effects versions 
-of the following equation:
-\begin{align}
-Y_{i,t} &=  \beta_1 S_{i,t}  +  \beta_2 S_{i,t} \times X_{i,1} +
-\alpha_{t} + \mu_i + \varepsilon_{i,t},   
-\label{eq:fixed_effects2} 
-\end{align}
-where $X_{i,1}$ are either assets holdings at the time of the first
-survey or education levels of the respondent or household head.
+The reason for the longer lag is that with an average gestation period
+of 9 months only crop losses that occurred longer ago are relevant.%
+\footnote{
+It is possible that the most recent crop loss could impact the likelihood
+of carrying a pregnancy to term, although this would require severe
+food deprivation \citep{Hernandez-Julian2014}.
+We investigate this possibility indirectly when examining the effects
+of crop loss on health.
+}
 
@@
 Our preferred estimates---woman level fixed effects---show that experiencing 
-crop loss decreases the likelihood of being pregnant by 9 percentage points 
-and the likelihood of a birth by 16 percentage points.
+crop loss decreases the likelihood of being pregnant by 10 percentage points 
+and the likelihood of a birth by 17 percentage points.
 These effects are substantial; on average across the four survey rounds, approximately 
@@
 to account for potential spacial correlation in shocks.
-Clustering at community level leads to slightly less statistically significance for 
-pregnancies and slightly higher statistically significance for births.
+Clustering at community level leads to slightly lower statistically significance 
+for pregnancies and slightly higher statistically significance for births.
 }
@@
 Our preferred model shows that crop loss is associated with
-a 7 percentage points increase in contraceptive use.%
+an almost 8 percentage points increase in contraceptive use.%
 \footnote{
@@
 The effect on overall contraceptive use is substantial considering that the 
-average use rates across the four rounds vary between 9 and 15\%.
+average use rates across the four rounds vary between 8 and 14\%.
 The statistically significant increase in contraceptive use following
@@
 crop loss leads to a statistically significant increase in
-traditional contraceptive use by 5.4 percentage points, whereas
-the effect on use of modern contraception is close to zero
+traditional contraceptive use by 6.2 percentage points, whereas
+the effect on use of modern contraception is 2 percentage points
 and not statistically significant. 
@@
 These results are consistent with the idea that households 
-deliberately postpone births in response to adverse shocks.%
-\footnote{
-To examine whether the effects of crop loss are an artifact of our
-chosen specifications we also tried alternative specifications
-as shown in Appendix Tables \ref{tab:logit}, \ref{tab:ln_croploss}, 
-and \ref{tab:continuous}.
-The results are qualitatively the same, especially for the Logit model
-and for the log of crop loss.
-Furthermore, Appendix Table \ref{tab:agegroups} shows the effects
-of crop losses by age groups.
-The effect of crop losses are concentrated for women younger than 32
-years of age, although as discussed above there appears to be 
-substantial measurement error in self-reported age.
-}
+deliberately postpone births in response to adverse shocks.
 
@@
 are we picking up behavior changes brought on by
-crop loss, such as decrease in coital frequency from substantial
-increase in hours worked or poorer health, that is later 
+crop loss---such as decrease in coital frequency from poor
+health or a substantial increase in hours worked---that is later 
 rationalized or reported as increased use of traditional 
@@
 strong if traditional contraceptives have low efficacy?
-We attempt to answer these questions in the remainder of the paper.
+We attempt to answer these questions in this section.
 
@@
 Indeed, we cannot reject that the absolute values of the point 
-estimates for births and contraceptive use and pregnancy and contraceptive use are
-the same, although we can reject at the 5\% significance level that the
-point estimates for births and pregnancy are the same.%
+estimates for pregnancy and contraceptive use are the same, 
+although we can reject that the point estimates are the same
+at the 5\% significance level for births and pregnancy and at
+the 10\% level for births and contraceptive use.%
 \footnote{
@@
 or the absence of the husband or the wife.
-Worsening health has two effects.
+Worsening health has three effects.
 First, more women will be unable to conceive, which would make
@@
 larger than the effect on pregnancy.
+Third, worse health could lead to lower coital frequency,
+which is also consistent with larger effects on births and
+pregnancy than on contraceptive use.
 Similarly, the absence of a partner will lower the risk of
 pregnancy and birth for a given level of contraceptive use.
+
+% Table with pregnancy and birth results
+\input{../tables/main_health_marriage.tex}
+
 Table \ref{tab:health} shows the impacts of crop loss on body mass index 
@@
 outmigration of partner, and risk of marriage dissolution. 
-
-% Table with pregnancy and birth results
-\input{../tables/main_health_marriage.tex}
-
 Both health outcomes have the expected signs, but the effects are small
-and not statistically significant.
+and not statistically significant, making it unlikely that it 
+would be associated even with a sufficient reduction in 
+coital frequency to explain the results.
 Similarly, Table \ref{tab:health} shows that shocks do not lead to a
@@
 % - hours worked and abstinence  / also shows that the shocks are "real"
-The second question is whether the amount worked increases so substantially 
+In addition to health related reductions in coital frequency, 
+it also possible that the amount worked increases so substantially 
 that coital frequency declines and the respondent reports this as abstinence.%
@@
 }
-We examine this possibility in two ways.
-First, for this explanation to hold the amount worked must increase 
+% We examine this possibility in two ways.
+% First, 
+For this explanation to hold the amount worked must increase 
 dramatically.%
@@
 and that effect is not statistically significant.
-Using log of crop loss, the effect is statistically significant
-although the effect is still relatively small; 
-a 1\% increase in the size of a crop loss leads to a 0.08 hours
+Using log of crop loss, the effect is statistically insignificant
+and the effect is still relatively small; 
+a 1\% increase in the size of a crop loss leads to a 0.05 hours
 increase in amount worked per week.
-
-Second, we can disentangle the different types of contraceptives.
-Given the small sample size this is, however, at best suggestive.
-Excluding abstinence, the effects of crop loss on contraceptive use
-are smaller and no longer statistically significant.%
-\footnote{
-Appendix Table \ref{tab:excludeAbs} shows the results.
-}
-With the relatively low use rates of modern contraceptives, this
-is not surprising given that abstinence is clearly the only 100\%
-effective traditional contraceptive.
 In sum, combined with an average work week of just below 44 hours a week,
@@
 
+% Second, we can disentangle the different types of contraceptives.
+% Given the small sample size this is, however, at best suggestive.
+% Excluding abstinence, the effects of crop loss on contraceptive use
+% are smaller and just above the 10\% significance level.
+% The effect on abstinence only is almost the same size, but that
+% effect is statistically significant at the 5\% level.%
+% \footnote{
+% Appendix Table \ref{tab:excludeAbs} shows the results.
+% }
+% With the relatively low use rates of modern contraceptives, this
+% is not surprising given that abstinence is clearly the only 100\%
+% effective traditional contraceptive.
+
 % Third, 
@@
 The final question is how the effect on postponement of fertility 
-is so strong when there is almost no use of modern contraceptives.
+is so strong when there is little use of modern contraceptives.
 In Table \ref{tab:effectiveness}, we examine to what extent shocks cause
@@
 that these women are also the most likely to be sexually active.
+Appendix Table \ref{tab:abstinence} shows the association 
+between crop loss, contraceptive use, and current pregnancy with 
+contraceptive use split by abstinence only, other traditional
+contraceptive use, and modern contraceptives.
+With the caveat that the low number of observations for each use 
+makes it difficult to draw strong conclusions, it appear that
+abstinence is the most ``effective'', followed by modern
+contraceptives, whereas other traditional contraceptive interacted
+with crop loss do not have a statistically significant association.
 }
@@
 
-\subsection{Robustness of Results}
-
-Finally, we present a number of robustness checks for our results.
-We first turn to household wealth.
+\subsection{Robustness of Results\label{sec:robustness}}
+
+Finally, we discuss a number of robustness checks for our results.
+The most important concern is our definition of crop loss.
+Above we define crop loss as a loss of more than 200 TZS.
+To check how sensitive our results are to this cut-off,
+we re-ran the regressions with varying cut-offs from 
+0 to 5,000 TZS.%
+\footnote{
+The results are shown in Appendix Table \ref{tab:cut_offs}.
+}
+The signs of the coefficients are the same across all cut-offs.
+Between 100 and 1,000 TZS, crop loss has a statistically significant
+effect on same outcome as above.
+Above 1,000 TZS the effects on pregnancy and birth become smaller and 
+eventually statistically significant insignificant.
+This is not surprising.
+With a higher cut-off, more and more households will have experienced 
+a crop loss but are not counted as having one, which will tend to 
+bias results towards zero if crop loss have an effect at levels 
+lower than the cut-off.%
+\footnote{
+This is supported by regressions---not shown---where we drop all observations 
+with crop loss larger than zero but smaller or equal to the cut-off and find 
+substantially larger coefficients.
+} 
+Using the log of crop loss instead of a cut-off does
+not change our conclusion that crop loss significantly reduces
+the likelihood of a pregnancy and birth and significantly
+increases the use of contraception.%
+\footnote{
+See Appendix Table \ref{tab:ln_croploss}.
+Using continuous crop loss leads to similar signed coefficients,
+although neither pregnancy or birth are statistically significant
+as shown in Appendix Table \ref{tab:continuous}.
+Given the unattractive assumption that the effect of an unit
+increase in crop loss is the same at all levels inherent in
+this specification, the lack of statistically significance is
+not surprising. 
+}
+The same is the case if we use a Logit model instead of the
+linear probability model.%
+\footnote{
+See Appendix Table \ref{tab:logit}.
+}
+
+% Trend (based on R2 question)
+A potential concern with the data is the large differences in
+the number of households that experience a crop loss across
+the waves.
+One possibility is that our results are driven by a time
+trend that affect households that are more likely to
+experience crop losses.
+We use two models to investigate this possibility.
+First, we interact a linear time trend with wave 1 characteristics 
+(assets per capita, age group dummies, and education group dummies).
+Second, we interact wave dummies with the same set of wave 1 
+characteristics.%
+\footnote{
+The results are shown in Appendix Table \ref{tab:timetrend_birth}.
+}
+The differences from the original results in point estimates 
+and levels of significance are minimal, indicating that 
+our results are not driven by differential trends in which 
+types of households that experience crop loss.
+
+% combine these two questions
+% Omitted variables (based on R2 question)
+% Time varying results (based on R1 question)
+Furthermore, our definition of crop loss is based on the 
+household's assessment and cannot be directly verified.
+It is therefore possible that omitted, time-varying 
+factors are responsible for the results. 
+We approach this issue in two ways.
+First, we included two other time-varying factors that
+may affect fertility, loss of livestock and a price
+index, in the main regressions.%
+\footnote{
+See results in Appendix Table \ref{tab:timevarying}.
+Note, that the loss of livestock unfortunately also
+includes livestock given away by the household.
+}
+Neither of these variables are statistically significant
+and the coefficient on crop loss does not change.
+[TK bla, bla - next sentences need a rewrite]
+Although this does not completely rule out the possibility
+that other time-varying factors, such as an epidemic or
+local conflict, could affect fertility decisions, it does
+imply that unobserved time-varying factors were not a 
+major influence.
+Conflict and epidemics would most likely have had 
+substantial effects on either or both of the new
+time-varying variables included.
+
+Second, we examined if rainfall predicts crop loss.%
+\footnote{
+See Appendix Table \ref{tab:rainfall}.
+}
+Higher rainfall is significantly associated with a 
+lower risk of crop loss,
+% , a dummy for missing
+% information on rainfall is also substantially 
+% associated with crop loss.
+% Furthermore, 
+although including rainfall as an additional
+variable in the main regressions did little to
+change the estimated effects.
+
+% - Something else changes in the village
+It is also possible that crop loss results in spill-overs to other
+households.
+We therefore added the fraction of households that experienced a crop
+loss in the community---over both the last 7 months and the period 
+between the two prior surveys---to the specifications.%
+\footnote{
+This fraction is calculated on the complete sample of households
+rather than the sample of women used here.
+Results are shown in Appendix Table \ref{tab:vill_croploss}.
+}
+The effects of crop loss on pregnancy and birth are slightly smaller,
+although the levels of statistical significance remain the same.
+The effects of any contraception use and use of traditional
+contraception are both lower, with the effect on traditional
+contraception no longer statistically significant.
+Only for traditional contraception use is the village level 
+coefficient statistically significant.
+Another way to look at this is that the spillovers from other
+households' exposure to shocks are generally not large enough to 
+force not exposed households to change their fertility and 
+contraceptive behavior.
+
+% - preemptive behavior (pregnant / contraceptive use predict shocks)
+
+Households may also anticipate shocks and
+change their behavior in response to the expected shocks.
+One way to examine this is to reverse the estimations.
+We therefore estimate whether contraceptive use at the time of the last 
+survey was associated with observing a crop loss between the last survey
+and the current one and similarly for pregnancy at prior survey and 
+birth in the period before the prior survey.%
+\footnote{
+Appendix Table \ref{tab:reverse} shows the results.
+We examine both the dummy for crop loss and the log of crop loss.
+}
+There are no statistically significant direct effects of 
+prior behavior on current crop loss, measured as either a
+dummy or log of amount of crop loss.
+The small effects and lack of statistical significance 
+indicate that crop losses are, indeed, unanticipated 
+shocks.
+
+% Other results
+
+[TK This last part is crap!]
+
+Finally, the Appendix discuss in detail whether the results 
+vary by age group, wealth, and education levels.
+None of these lead to strong conclusions on how crop loss 
+affect fertility decision by household and individual
+characteristics.
+The Appendix also discuss how including additional lags
+of crop loss affects the results.
+In principle, if shocks lead to a postponement of fertility we 
+should see longer lags be associated with higher fertility, 
+whereas if shocks lead to a permanent reduction in fertility we
+should not see any compensatory effect.
+There are, however, a number of issues with this approach.
+First, it is possible that the effect of a crop loss is felt
+in subsequent period and not only in the period in which it
+was observed.
+Second, longer spacing between births can be costly because economies
+of scale are harder to realize and because the mother may have to 
+be out of the labor force for longer if she already have young
+children to care for.
+This implies that a household that are hit twice in a row by
+a crop loss is likely to behave differently from a household that
+is hit once.
+Both of these factors make it difficult to draw strong 
+predictions on what the effects should be (except that the
+initial impact should lead to postponement of fertility).
+
+\section{Conclusion}
+
+We estimate the relationship between household income shocks and decisions
+on timing of fertility. 
+To account for potential correlation between unobservable household
+characteristics and both shocks and fertility decisions, we employ a fixed effect model.
+Our estimates show that both the likelihood of pregnancy and birth decrease in
+response to a crop loss and that contraceptive use, especially use of traditional 
+methods, such as abstinence and rhythm, significantly increases.
+
+We examine a number of causes that may explain the reduction 
+in fertility following a crop loss.
+Crop loss shocks do not worsen women's health enough to 
+reduce pregnancies and childbirths to the extent that we find.
+Furthermore, with no significant change in the likelihood of migration or 
+dissolution of partnership in response to crop loss, there is no evidence 
+that the reduction is due to physical separation of spouses or partner.
+As expected, the number of hours worked increases for women after a shock, 
+but the effect is unlikely to be large enough to affect fertility 
+through a change in coital frequency.
+
+Our main conclusion is that households consciously delay 
+fertility through contraceptive use to cope with income shocks. 
+This is especially of interest since this is accomplished 
+predominately through the use of traditional contraceptives.
+If traditional contraceptives have low efficacy, a natural question is 
+how the substantial reductions in likelihood of births and pregnancy 
+come about.
+We present suggestive evidence that show that only those households
+with a strong incentive, i.e. those exposed to crop loss, are 
+effective users of traditional contraceptives.%
+\footnote{
+This is in line with the results in \cite{Rosenzweig1989},
+where women with the highest opportunity cost of children
+are also the most effective user of traditional contraceptives,
+although the setting is very different.
+}
+
+There are two potential caveats to our results.
+First, the sample size is small and the survey rounds
+could have been further apart, but the longitudinal nature
+of the data means that we can control for unobservable 
+time-invariant individual and household characteristics.
+Second, not surprisingly there appears to be measurement 
+errors.
+% Reported age and education levels are not always consistent
+% across rounds, which is an indication that we probably also
+% have measurement errors in the other variables, although it is
+% harder to verify.
+To the extent that these measurement errors are classical 
+measurement errors, this biases our results towards zero,
+making it harder to show an effect of crop loss.%
+\footnote{
+The problems with measurement error, combined with the small
+sample size, may also explain why we do not find clear
+differences in the response to crop loss across age groups,
+education, and wealth levels.
+}
+
+The data unfortunately do not allow us to establish whether shocks
+have long-term impacts on fertility outcomes.
+Results from Guatemala indicate that only shocks that occur 
+toward the end of a woman's reproductive life is likely to have an 
+impact on final fertility because the ability to ``make up'' 
+postponed births are smaller \citep{Portner2014}.
+There is, therefore, little reason to believe that temporary shocks,
+like the crop losses studied here, lead to over-all changes in fertility,
+although the underlying \emph{risk} of shocks may.
+
+% [Lead with this or the next sentence?]
+% Our result that people are able to postpone fertility using mainly 
+% traditional contraceptive do, however, have implications for our 
+% understanding of long-term fertility.
+% [Alternative lead]
+What implications do our results have for our understanding of 
+long-term fertility?
+% What we do not know is whether the psychic cost of using traditional 
+% contraceptives is so high that although households are willing to use 
+% them to deal with short-term shocks, they become difficult to use for 
+% long-run moderation of fertility.
+The economic literature on fertility transition in Europe and the US 
+show a substantial decline in fertility and control of the timing of 
+fertility before the introduction of modern contraceptives, although 
+the methods used are generally not well documented.%
+\footnote{
+See, for example, \citet{Guinnane2011} on fertility reduction and 
+\citet{Cinnirella2017} on spacing of birth.
+On contraceptive methods used see, for example, \citet{Michael1976},
+\citet{David1986}, and \citet{Santow1995}.
+}
+Furthermore, studies of the introduction of modern family planning, 
+using individual level data, rarely find reductions in fertility larger 
+than one child.
+Both set of results indicate that households are able to 
+limit long-term fertility using traditional contraceptives.%
+\footnote{
+See, for example, \cite{Portner2014a} and the literature surveyed therein.
+}
+Our individual level data allow us to directly link the use of
+traditional contraceptives and fertility outcomes.
+We argue that, contrary to common perception, traditional
+contraceptive methods have a high degree of efficacy when the
+incentives to prevent pregnancy or birth are sufficiently high.
+Because of the small data set we cannot, however, differentiate between
+the different traditional methods, although---contrary to Europe 
+and the US---withdrawal plays little or no role in this area.
+
+Finally, our focus on traditional contraceptives does
+not indicate that access to modern contraceptives is unimportant.
+Although use of modern contraceptives is low in the area we show
+they are also effective in preventing pregnancies and births after a shock.
+Children born after shocks are likely to fare worse than children born 
+during better times.
+The mother is also likely to fare better if she does not give birth 
+immediately following an economic shock that may impact her ability 
+to recover and care for the new child.
+Our results imply that providing better access to modern contraceptives 
+during times of economic shocks may improve the ability of households 
+to postpone births.
+In turn, this could lead to better mother and child health outcomes.
+
+% \clearpage
+% 
+% \newpage
+
+\bibliographystyle{econometrica}
+\bibliography{incomeShocks-jde-r1}
+\addcontentsline{toc}{section}{References}
+
+% \newpage
+
+\clearpage
+
+\appendix
+%\renewcommand\thesection{Appendix \Alph{section}:}
+
+\renewcommand\thetable{A-\arabic{table}}
+\renewcommand\thefigure{A-\arabic{figure}}
+\setcounter{table}{0}
+\setcounter{figure}{0}
+
+\section{Appendix}
+
+% Attrition descriptive statistics across waves
+\input{../tables/appendix_desstat2.tex}
+
+\input{../tables/appendix_desstat1.tex}
+
+% Full OLS and community fixed effects results for Pregnancy and birth
+\input{../tables/appendix_pregnant_birth.tex}
+
+% Community clustering for pregnancy and birth results
+\input{../tables/appendix_cluster_pregnant_birth.tex}
+
+% Full OLS and community fixed effects results for Contraceptives results
+\input{../tables/appendix_contraceptives.tex}
+
+% Community clustering for for contraceptives results
+\input{../tables/appendix_cluster_contraceptives.tex}
+
+% Conscious Decision section
+
+% Hours worked and crop loss
+\input{../tables/appendix_hours.tex}
+
+% % Excluding abstinence
+% \input{../tables/appendix_exclude_abstinence.tex}
+
+% Effectiveness of contraceptives by method
+\input{../tables/appendix_abstinence.tex}
+
+% Robustness section tables
+
+% Different cut-offs 
+\input{../tables/appendix_cut_offs.tex}
+
+% Log of crop loss
+\input{../tables/appendix_log_croploss.tex}
+
+% Linear crop loss
+\input{../tables/appendix_cont_croploss.tex}
+
+% Logit versions of main dummy set-up
+\input{../tables/appendix_logit.tex}
+
+% Time trend and wave dummies interacted with wave 1 characteristics
+\input{../tables/appendix_timetrend.tex}
+
+% Time-varying variables
+\input{../tables/appendix_timevarying.tex}
+
+% Does rainfall explain crop losses?
+\input{../tables/appendix_rainfall.tex}
+
+% Village crop loss 
+\input{../tables/appendix_village.tex}
+
+% Reverse regressions - predictive or pseudo
+\input{../tables/appendix_reverse.tex}
+
+\clearpage
+\section{Age, Wealth, and Education}
+
+To further investigate whether the response to shock vary by household
+characteristics, we also estimated individual level fixed effects versions 
+of equation (\ref{eq:fixed_effects2}),
+\begin{align}
+Y_{i,t} &=  \beta_1 S_{i,t}  +  \beta_2 S_{i,t} \times X_{i,1} +
+\alpha_{t} + \mu_i + \varepsilon_{i,t},   
+\label{eq:fixed_effects2} 
+\end{align}
+where $X_{i,1}$ are either age group, assets holdings at the time of 
+the first survey or education levels of the respondent or household head.
+Table \ref{tab:agegroups} shows the effects of crop losses by age groups.
+The effect of crop losses are concentrated for women younger than 32
+years of age, although as discussed above there appears to be 
+substantial measurement error in self-reported age.
+
+% Age groups and crop loss effects
+\input{../tables/appendix_age_groups.tex}
+
+We next turn to household wealth.
 There are three, potentially opposite, effects of wealth.
@@
 We therefore use assets owned in the first round, even though 
-shocks that occurred before the first round likely affect these.%
-\footnote{
+shocks that occurred before the first round likely affect these.
 Appendix Table \ref{tab:assets} expands the main results above by adding 
 the interaction of crop loss and assets.
-}
 In all cases, the coefficients for the interaction of assets in round 1 and 
@@
 
+% Effects of Crop Loss on Contraceptives Use by Assets
+\input{../tables/appendix_asset_interaction.tex}
+
 We next turn to education.
@@
 Second, education, especially for women, might affect the efficacy of 
-traditional contraceptives \citep{Rosenzweig1989}.%
-\footnote{
-Appendix Table \ref{tab:education} presents the effect of crop loss on
+traditional contraceptives \citep{Rosenzweig1989}.
+Table \ref{tab:education} presents the effect of crop loss on
 likelihood of pregnancy, birth, and contraceptive use for women
-by the three levels of education discussed above 
-}
+by the three levels of education discussed above.
+
+% Effects of Crop Loss on Contraceptives Use by Education Level
+\input{../tables/appendix_education.tex}
 
@@
 
-% [robustness checks]
-
-% - Something else changes in the village
-It is also possible that households respond to changed circumstances
-in their local area.
-We therefore used the fraction of households that experienced a crop
-loss in the community---over both the last 7 months and the period 
-between the two prior surveys.%
-\footnote{
-This fraction is calculated on the complete sample of households
-rather than the sample of women used here.
-Results are shown in Appendix Table \ref{tab:vill_croploss}.
-}
-The effect on pregnancy remains the same, whereas the effect on
-births is slightly larger, although not statistically significantly so.
-There is no change in the effect on the different measures of
-contraceptive use.
-Furthermore, none of the village level coefficients are statistically 
-significant.
-Another way to look at this is that the spillovers from other
-households' exposure to shocks are not large enough to force 
-not exposed households to change their fertility and contraceptive
-behavior.
-
-% - preemptive behavior (pregnant / contraceptive use predict shocks)
-
-Finally, households may anticipate shocks and
-change their behavior in response to the expected shocks.
-One way to examine this is to reverse the estimations.
-We therefore estimate whether contraceptive use at the time of the last 
-survey was associated with observing a crop loss between the last survey
-and the current one and similarly for pregnancy at prior survey and 
-birth in the period before the prior survey.%
-\footnote{
-Appendix Table \ref{tab:reverse} shows the results.
-We examine both the dummy for crop loss and the log of crop loss.
-}
-There are no statistically significant direct effects of 
-prior behavior on current crop loss, measured as either a
-dummy or log of amount of crop loss.
-The small effects and lack of statistical significance 
-indicate that crop losses are, indeed, unanticipated 
-shocks.
-
-\section{Conclusion}
-
-We estimate the relationship between household income shocks and decisions
-on timing of fertility. 
-To account for potential correlation between unobservable household
-characteristics and both shocks and fertility decisions, we employ a fixed effect model.
-Our estimates show that both the likelihood of pregnancy and birth decrease in
-response to a crop loss and that contraceptive use, especially use of traditional 
-methods, such as abstinence and rhythm, significantly increases.
-
-We examine a number of potential explanations for these results.
-Crop loss shocks do not worsen women's health enough to 
-reduce pregnancies and childbirths to the extent that we find.
-Furthermore, with no significant change in the likelihood of migration or 
-dissolution of partnership in response to crop loss, there is no evidence 
-that the reduction is due to physical separation of spouses or partner.
-As expected, the number of hours worked increases for women after a shock, 
-but the effect is unlikely to be large enough to affect fertility 
-through a change in coital frequency.
-
-Our main conclusion is that households consciously delay 
-fertility through contraceptive use to cope with income shocks. 
-This is especially of interest since this is accomplished 
-predominately through the use of traditional contraceptives.
-If traditional contraceptives have low efficacy, a natural question is 
-how the substantial reductions in likelihood of births and pregnancy 
-come about.
-We present suggestive evidence that show that only those households
-with a strong incentive, i.e. those exposed to crop loss, are 
-effective users of traditional contraceptives.%
-\footnote{
-This is in line with the results in \cite{Rosenzweig1989},
-where women with the highest opportunity cost of children
-are also the most effective user of traditional contraceptives,
-although the setting is very different.
-}
-
-There are two potential caveats to our results.
-First, the sample size is small and the survey rounds
-could have been further apart, but the longitudinal nature
-of the data means that we can control for unobservable 
-time-invariant individual and household characteristics.
-Second, not surprisingly there appears to be measurement 
-errors.
-Reported age and education levels are not always consistent
-across rounds, which is an indication that we probably also
-have measurement errors in the other variables, although it is
-harder to verify.
-To the extent that these measurement errors are classical 
-measurement errors, this biases our results towards zero,
-making it harder to show an effect of crop loss.%
-\footnote{
-The problems with measurement error, combined with the small
-sample size, may also explain why we do not find clear
-differences in the response to crop loss across education
-and wealth levels.
-}
-
-The data unfortunately do not allow us to establish whether shocks
-have long-term impacts on fertility outcomes.
-Results from Guatemala indicate that only shocks that occur 
-toward the end of a woman's reproductive life is likely to have an 
-impact on final fertility because the ability to ``make up'' 
-postponed births are smaller \citep{Portner2014}.
-There is, therefore, little reason to believe that temporary shocks,
-like the crop losses studied here, lead to over-all changes in fertility,
-although the underlying \emph{risk} of shocks may.
-
-[Lead with this or the next sentence?]
-Our result that people are able to postpone fertility using mainly 
-traditional contraceptive do, however, have implications for our 
-understanding of long-term fertility.
-[Alternative lead]
-What implications do our results that people are able to postpone
-fertility using mainly traditional contraceptive have for our
-understanding of long-term fertility?
-% What we do not know is whether the psychic cost of using traditional 
-% contraceptives is so high that although households are willing to use 
-% them to deal with short-term shocks, they become difficult to use for 
-% long-run moderation of fertility.
-The economic literature on fertility transition in Europe and the US 
-show a substantial decline in fertility and control of the timing of 
-fertility before the introduction of modern contraceptives, although 
-the methods used are generally not well documented.%
-\footnote{
-See, for example, \citet{Guinnane2011} on fertility reduction and 
-\citet{Cinnirella2017} on spacing of birth.
-On contraceptive methods used see, for example, \citet{Michael1976},
-\citet{David1986}, and \citet{Santow1995}.
-}
-Furthermore, studies of the introduction of modern family planning, 
-using individual level data, rarely find reductions in fertility larger 
-than one child.
-Both set of results indicate that households are able to 
-limit long-term fertility using traditional contraceptives.%
-\footnote{
-See, for example, \cite{Portner2014a} and the literature surveyed therein.
-}
-Our individual level data allow us to directly link the use of
-traditional contraceptives and fertility outcomes.
-We argue that, contrary to common perception, traditional
-contraceptive methods have a high degree of efficacy when the
-incentives to prevent pregnancy or birth are sufficiently high.
-Because of the small data set we cannot, however, differentiate between
-the different traditional methods, although---contrary to Europe 
-and the US---withdrawal clearly plays little or no role in this area.
-
-Finally, our focus on traditional contraceptives does
-not indicate that access to modern contraceptives is unimportant.
-Although use of modern contraceptives is low in the area we show
-they are also effective in preventing pregnancies and births after a shock.
-Children born after shocks are likely to fare worse than children born 
-during better times.
-The mother is also likely to fare better if she does not give birth 
-immediately following an economic shock that may impact her ability 
-to recover and care for the new child.
-Our results imply that providing better access to modern contraceptives 
-during times of economic shocks may improve the ability of households 
-to postpone births.
-In turn, this could lead to better mother and child health outcomes.
-
-\clearpage
-
-\newpage
-\bibliographystyle{econometrica}
-\bibliography{incomeShocks-jde-r1}
-\addcontentsline{toc}{section}{References}
-
-% \newpage
-
-\clearpage
-
-\appendix
-%\renewcommand\thesection{Appendix \Alph{section}:}
-
-\renewcommand\thetable{A-\arabic{table}}
-\renewcommand\thefigure{A-\arabic{figure}}
-\setcounter{table}{0}
-\setcounter{figure}{0}
-
-\section{Appendix}
-
-% Attrition descriptive statistics across waves
-\input{../tables/appendix_desstat2.tex}
-
-\input{../tables/appendix_desstat1.tex}
-
-% Full OLS and community fixed effects results for Pregnancy and birth
-\input{../tables/appendix_pregnant_birth.tex}
-
-% Community clustering for pregnancy and birth results
-\input{../tables/appendix_cluster_pregnant_birth.tex}
-
-% Full OLS and community fixed effects results for Contraceptives results
-\input{../tables/appendix_contraceptives.tex}
-
-% Community clustering for for contraceptives results
-\input{../tables/appendix_cluster_contraceptives.tex}
-
-% Logit versions of main dummy set-up
-\input{../tables/appendix_logit.tex}
-
-% Log of crop loss
-\input{../tables/appendix_log_croploss.tex}
-
-% Linear crop loss
-\input{../tables/appendix_cont_croploss.tex}
-
-% Different cut-offs 
-\input{../tables/appendix_cut_offs.tex}
-
-% Age groups and crop loss effects
-\input{../tables/appendix_age_groups.tex}
-
-% Hours worked and crop loss
-\input{../tables/appendix_hours.tex}
-
-% Excluding abstinence
-\input{../tables/appendix_exclude_abstinence.tex}
-
-% Effectiveness of contraceptives by method
-\input{../tables/appendix_abstinence.tex}
-
-% Effects of Crop Loss on Contraceptives Use by Assets
-\input{../tables/appendix_asset_interaction.tex}
-
-% Effects of Crop Loss on Contraceptives Use by Education Level
-\input{../tables/appendix_education.tex}
-
-% Village crop loss 
-\input{../tables/appendix_village.tex}
-
-% Reverse regressions - predictive or pseudo
-\input{../tables/appendix_reverse.tex}
+\section{The Effect of Crop Loss in Prior Periods}
+
+In this section we discuss how including additional lags
+of crop loss affects the results.
+In principle, if shocks lead to a postponement of fertility we 
+should see longer lags be associated with higher fertility, 
+whereas if shocks lead to a permanent reduction in fertility we
+should not see any compensatory effect.
+There are, however, a number of issues with this approach.
+First, it is possible that the effect of a crop loss is felt
+in subsequent periods and not only in the period in which it
+was observed.
+Second, we lose 1/4 of observations each time we include an 
+additional lag, which leaves us with less variation because
+the first round had a higher proportion of crop losses than
+the subsequent rounds.
+This is especially a concern for contraception use because
+of the higher likelihood of measurement error discussed in the
+paper.
+
+Furthermore, even if shocks only had an impact in one period, 
+there are two competing factors at work here.
+First, a negative shocks will make people more likely to postpone 
+their pregnancy/birth because of the factors discussed in the
+paper.
+Second, longer spacing between births can be costly because economies
+of scale are harder to realize and because the mother may have to 
+be out of the labor force for longer if she already have young
+children to care for.
+This implies that a household that are hit twice in a row by
+a crop loss is likely to behave differently from a household that
+is hit once.
+Both of these factors make it difficult to draw strong 
+predictions on what the effects should be (except that the
+initial impact should lead to postponement of fertility).
+We have therefore also added an interaction between crop loss for 
+different periods with the idea that there is a cost to continually 
+postponing births.
 
@@
 
-% Time trend and wave dummies interacted with wave 1 characteristics
-\input{../tables/appendix_timetrend.tex}
-
-% Does rainfall explain crop losses?
-\input{../tables/appendix_rainfall.tex}
-
-% Time-varying variables
-\input{../tables/appendix_timevarying.tex}
+Tables \ref{tab:postpone_birth} and \ref{tab:postpone_contraceptive}
+show three models, the main results presented in the paper, one
+where an additional lag is included, and one where the interaction
+between crop loss and the additional lag in crop loss are included.
+Including only the additional lag of crop loss it appears that 
+there is a weak positive association between crop loss in the prior
+period and the likelihood of pregnancy and birth, although the 
+coefficients are small and statistically insignificant.
+There is still a negative effect of the immediate crop loss
+variable, although these are also not statistically significant.
+If we include the interaction between crop losses we instead
+find negative association for both immediate and lagged crop
+loss with a positive interaction coefficient.
+For contraceptive use we see a similar pattern (of course,
+with the signs reversed).
+
+Although these results should be interpreted with caution the
+results appear most in line with the following conclusions.
+First, crop losses tend to affect not only this period but also 
+(weakly) the period after as indicated by the negative coefficient 
+on the lagged crop loss.
+Second, experiencing two crop losses in a row makes it \emph{less}
+likely for people to postpone fertility (although the effect 
+of the crop losses is still negative), most likely because of
+the cost associated with longer spacing.
 

### 2017-10-09 paper/incomeShocks-jde-r1.tex
Respond to R1's question on wave 1 assets
The discussion of assets interacted with crop loss was made a separate
section prior. This revision adds log of crop loss interacted with
log of assets and results where the first wave crop loss is dropped
because it might affect wave 1 assets.

The new section is edited to reflect changes in results.

All results are combined into one table, so three original analysis
files are deleted because their results show up elsewhere.

@@
 Table \ref{tab:agegroups} shows the effects of crop losses by age groups.
-The effect of crop losses are concentrated for women younger than 32
-years of age, although as discussed above there appears to be 
-substantial measurement error in self-reported age.
+The effect of crop losses appear to be mainly concentrated among women 
+younger than 32 years of age, although as discussed above there appears 
+to be substantial measurement error in self-reported age.
 
@@
 
+Appendix Table \ref{tab:assets} expands the main results above by adding 
+the interaction of crop loss and assets, both for the main specification
+and using log of crop loss and log of assets.
 As crop loss may affect asset levels, we should ideally use assets owned 
@@
 survey round.
-We therefore use assets owned in the first round, even though 
-shocks that occurred before the first round likely affect these.
-Appendix Table \ref{tab:assets} expands the main results above by adding 
-the interaction of crop loss and assets.
-In all cases, the coefficients for the interaction of assets in round 1 and 
-crop loss are small and not statistically significant,
-suggesting that wealth level may not be a determining
-factor regarding the contraceptive use decision following shocks. 
+We therefore present two different specifications.
+The first specification includes all waves, even though shocks that 
+occurred before the first round potentially affect the asset level.
+The second specification excludes wave 1 shocks, which results in
+substantial loss of sample size and variation in crop loss.
+
+For pregnancy and births, the coefficients for the interaction of assets 
+in round 1 and crop loss are small and not statistically significant,
+suggesting that wealth level may have a limited impact on fertility
+decisions following shocks. 
+The interactions are more likely to have a statistically significant
+effect for contraceptive use, although the coefficients are still
+small (recall that assets are measured in 1,000,000 TZS).
+
 There are three potential reasons why we do not find an effect of wealth.
@@
 The reductions in likelihoods of being pregnant or having a birth
-following crop loss are now close to each other at 17 and 19 percentage
-points, and the increase in use of any contraception is 10 percentage
+following crop loss are now close to each other at 16 and 19 percentage
+points, and the increase in use of any contraception is 8 percentage
 points.
@@
 A crop loss leads to reductions in pregnancy and birth by
-15 and 19 percentage points and an increase of 8 percentage points
+14 and 19 percentage points and an increase of 7 percentage points
 in the use of any contraception.
@@
 effect of contraceptive use following a crop loss, but the
-effects on pregnancy and birth are about half the size and
-not statistically significant.
+effects on pregnancy and birth---while close to the 7 or more 
+years of education group---are not statistically significant.
 

### 2017-10-09 response/response-jde-1.tex
Respond to R1's question on assets

@@
 
-[RESPONSE:] We have added the results where round 1 has been excluded
-to the relevant tables.
-The results are broadly consistent with what we originally found, 
-although less likely to be statistically significant 
-because of the lower variation in crop loss and the smaller sample
-size.
-
-[Already ran these; might be worthwhile having the results as an extra
-panel in the relevant tables]
+[RESPONSE:] We have added the results where round 1 shocks has been 
+excluded to the relevant table in the Appendix section on age
+groups, wealth, and education.
+The results are broadly consistent with what we originally found
+for pregnancy and birth, 
+The contraceptive use results have the correct sign but the direct
+effect is not statistically significant, presumably because of the 
+lower variation in crop loss and the smaller sample size.
 

### 2017-10-09 response/response-jde-1.tex
Add list of major changes to paper in response

@@
 
-[Additional discussion of overall things we have changed---if any]
-- move discussion of education, wealth, and education results to appendix
-to save space
-
 We rewrote the code for cleaning and analyzing the data from the ground up 
@@
 
+To remain within the page limit we had to substantially reorganize
+part of the paper.
+The following is a list of the more substantial changes made:
+\begin{enumerate}
+\item We removed the map of Kagera.
+\item In response to the reviewers' comments we have almost completely
+rewritten the Robustness section as detailed below.
+\item As part of the rewrite of the Robustness section we moved 
+the discussion of age, wealth, and education results to a new Appendix
+section since these results do little to clarify the mechanisms 
+behind the main results.
+We still address and incorporate the comments from reviewers that
+relate to these outcomes.
+\item We moved the discussion of estimation strategy for age,
+wealth, and education estimation to the new Appendix section.
+\item We have added another new Appendix section that examine results 
+with including an additional lag for crop loss (see response to
+Reviewer 1 below).
+\item A number of new Appendix tables have been added in response
+to various comments and other edited.
+\end{enumerate}
+
+We have tried to highlight the factors that can affect our identification 
+strategy, but if we need to tone down the causality claims further
+we are happy to do so.
 We list below our responses to the individual comments and suggestions.
-
-[Possible changes]
-If the paper is too long, one possibility is to move the discussion of
-age, education, and wealth descriptive statistics to the Appendix.
-We only need this for the OLS and for the discussion of those factors,
-which is now in the Appendix.
 

### 2017-10-09 response/response-jde-1.tex
Respond to R1 on the role of traditional contraceptives

@@
 
-We have edited the introduction to highlight the high use of traditional 
+We have edited the Introduction to highlight the role of traditional 
 contraception and that most of the increase in contraception after a 
 shock comes from traditional methods.
-We have also added an additional point to the discussion of the
-main contributions of the paper.
+We have also added an additional point to our list of the main 
+contributions of the paper reflecting your suggestion.
 In addition to the \citet{Guinnane2011} paper, we also cite a just

### 2017-10-09 paper/incomeShocks-jde-r1.tex
Respond to R1 on the role of traditional contraceptives

@@
 using individual level data, rarely find reductions in fertility larger 
-than one child.
+than one child.%
+\footnote{
+See, for example, \citet{Portner2014a} 
+and \citet{Portner2018} and the literature surveyed therein.
+}
 Both set of results indicate that households are able to 
-limit long-term fertility using traditional contraceptives.%
-\footnote{
-See, for example, \cite{Portner2014a} and the literature surveyed therein.
-}
+limit long-term fertility using traditional contraceptives.
 Our individual level data allow us to directly link the use of

### 2017-10-10 response/response-jde-1.tex
Respond to R1 on traditional contraceptives

@@
 It is correct that the current fixed effects model cannot tell
-us whether we observe postponement of fertility or a permanent
-reduction.
+us whether we observe postponement of fertility or a permanent reduction.
 We focus on postponement for two reasons.
 
-First, active postponement is a necessary---but not sufficient---condition
-for permanent fertility reduction.
-If we cannot show a conscious decision to delay fertility, it is 
-hard to argue that we can ever find evidence of conscious permanent 
-reduction in fertility.
+First, we consider the ability to actively postpone fertility as a 
+necessary---but not sufficient---condition for permanent fertility reduction.
+In other words, we see postponement as a weaker claim than permanent 
+fertility reduction.
+If we cannot show a conscious decision to delay fertility under the
+right circumstances, it is hard to argue that we can convincingly
+show find evidence of conscious, permanent reduction in fertility.
+Note that this does not follow the original demography
+literature, but is in line with the more recent research
+\citep[See discussion in][]{Cinnirella2017}.
 
-Second, it is not clear that shocks---in and off themselves---should
-substantially affect completed fertility. 
-We discuss evidence from Guatemala that the overall \emph{risk} of 
+Second, from a theoretical standpoint it is not clear that 
+shocks---in and off themselves---should substantially affect completed 
+fertility. 
+We now discuss evidence from Guatemala that the overall \emph{risk} of 
 experiencing a shock is likely to affect fertility, but that only
@@
 affected by the overall \emph{risk}, whereas the exact timing of
-fertility is driven by a combination of risk and shocks (probably
-mostly shocks).
+fertility is driven by a combination of risk and shocks, but probably
+mostly shocks.
 
-The longitudinal nature of the data is important for our ability
-to control for unobservable village, household, and individual 
-characteristics, but the trade-off is that the data only cover 
-32 months and that is too short to analyze long-term fertility
-outcomes.
-We followed your suggestion of adding an additional lag to 
-the model and discuss the results in a new section in the Appendix.
-There are, however, a number of issues with this approach.
-First, it is possible that the effect of a crop loss is felt
-in subsequent period and not only in the period in which it
-was observed.
-Second, we lose 1/4 of observations each time we include an 
-additional lag, which leaves us with less variation because
-the first round had a higher proportion of crop losses than
-the subsequent rounds
+% The longitudinal nature of the data is important for our ability
+% to control for unobservable village, household, and individual 
+% characteristics, but the trade-off is that the data only cover 
+% 32 months and that is too short to analyze long-term fertility
+% outcomes.
+That said, we did follow your suggestion of adding an additional 
+lag to the model and discuss the results in a new Appendix section.
+In principle, if shocks lead to a postponement of fertility,
+crop losses that occurred further in the past should be
+associated with an \emph{increased} likelihood of current pregnancy 
+or birth (an initial reduction in fertility should be followed
+by a compensatory increase later on), 
+whereas if shocks lead to a permanent reduction in fertility we
+should not see any compensatory effect.
 
-Furthermore, even if shocks only had an impact in one period, 
-there are two competing factors at work here.
-First, a negative shocks will make people more likely to postpone 
-their pregnancy/birth because of the factors discussed in the
-paper.
-Second, longer spacing between births can be costly because economies
-of scale are harder to realize and because the mother may have to 
-be out of the labor force for longer if she already have young
-children to care for.
-We have therefore also added an interaction between crop loss for 
-different periods with the idea that there is a cost to continually 
-postponing births.
+Although the results are potentially interesting in themselves,
+because of data limitation the approach unfortunately does not  
+provide a clean test for whether households postpone or reduce their 
+fertility in response to shocks.
+First, given the short duration between surveys and the lack of 
+information about when the shocks occurred, it is possible and
+even likely that the effects of a crop loss last for more than one period.
+Hence, if we find a negative effect of more distant crop losses
+this may be the result of the negative effects of the crop losses
+still affecting the household.
+Second, we lose 1/4 of the original number of observations each 
+time we include an additional lag, which leaves us with less variation 
+in crop loss because the first round had a higher proportion of crop 
+losses than the subsequent rounds.
 
-If crop losses last more than one period and there is a cost 
-associated with further postponement, the effect of both shocks 
-on their own should be same and the interaction the opposite. 
-Broadly speaking that appears to hold. 
-The upshot---as far as we can tell---is that what we observe is 
-postponement and not an overall limitation in fertility. 
-This makes sense given results for Guatemala, where long-term risk 
-affects the desire fertility \emph{level}, whereas shocks experienced 
-affect the \emph{timing} of fertility. 
-We have added a discussion of this in the Conclusion.
+Even if crop losses' impact are limited to only one period, there 
+are two additional concerns.
+First, the compensatory effect may be spread over multiple 
+periods because of the uncertainty associated with waiting time
+to conception, i.e. a couple may be trying to conceive in the
+period after the crop loss but fail.
+This makes a positive compensatory effect harder to 
+identify.
+Second, it is possible that households may respond differently to a 
+shock if they are hit repeatedly.
+There are two competing factors at work.
+On one hand, repeated shocks may exacerbate the negative impact on the
+household, making it \emph{more} likely that fertility is postponed.
+On the other hand, even longer spacing between births can be costly 
+because economies of scale are harder to realize and because the 
+mother may have to  be out of the labor force for longer if she 
+already have young children to care for, which will make it
+\emph{less} likely that fertility is postponed.
+To capture the idea that repeated shocks have an additional
+impact we therefore also include the interaction between 
+crop losses.
+
+With the data limitations and additional concerns,
+it is difficult to draw strong predictions on what the coefficients
+should be (except that the initial impact should lead to postponement 
+of fertility).
+Including only the additional lag of crop loss it appears that 
+there is a weak positive association between crop loss in the prior
+period and the likelihood of pregnancy and birth, although the 
+coefficients are small and statistically insignificant.
+There is still a negative effect of the immediate crop loss
+variable, although these are also not statistically significant.
+These results are broadly consistent with a compensatory response
+where crop loss longer occur leads to higher fertility now.
+However, if we include the interaction between crop losses we 
+instead find negative association for both immediate and lagged 
+crop loss with a positive interaction coefficient.
+Even with the positive interaction coefficient the estimated 
+impact of crop losses in the two prior periods is negative.
+For contraceptive use we see similar patterns (with the signs 
+reversed).
+
+Although these results should be interpreted with caution, they
+are not inconsistent with a compensatory/postponement effect, 
+as long as some crop losses' impact last for more than one
+period after its occurrence and/or the compensatory effect is
+spread over more than one period.
+In either specification the coefficients on lagged crop loss
+are small relative to the immediate impact.
+In sum, we cannot establish with any degree of confidence from
+these results whether the response to shocks lead to only postponement 
+of shocks or a permanent reduction.
+Given the findings from Guatemala that shocks only 
+have a negative impact on completed fertility if they occur
+late in a woman's reproductive life and that completed 
+fertility is mainly driven by the \emph{risk} of shocks,
+it is most likely that what we observe are, indeed, postponement
+rather than a reduction in completed fertility \citep{Portner2014}.
 
@@
 table to the Appendix, which further subdivides the results in Table 6.
-Table A-13 [TK] show, not surprisingly, that abstinence is the most 
+Appendix Table A-8 shows, not surprisingly, that abstinence is the most 
 effective traditional contraception, although only in combination 
@@
 is not statistically significant.
-We discuss these results at the end of the section on 
+We discuss these results toward the end of the section on 
 ``Conscious Decision or Unintended Consequences?''.

### 2017-10-10 paper/incomeShocks-jde-r1.tex
Respond to R1 on traditional contraceptives

@@
 
-[TK This last part is crap!]
-
-Finally, the Appendix discuss in detail whether the results 
+Finally, in the Appendix we discuss in detail whether the results 
 vary by age group, wealth, and education levels.
@@
 characteristics.
-The Appendix also discuss how including additional lags
-of crop loss affects the results.
-In principle, if shocks lead to a postponement of fertility we 
-should see longer lags be associated with higher fertility, 
-whereas if shocks lead to a permanent reduction in fertility we
-should not see any compensatory effect.
-There are, however, a number of issues with this approach.
-First, it is possible that the effect of a crop loss is felt
-in subsequent period and not only in the period in which it
-was observed.
-Second, longer spacing between births can be costly because economies
-of scale are harder to realize and because the mother may have to 
-be out of the labor force for longer if she already have young
-children to care for.
-This implies that a household that are hit twice in a row by
-a crop loss is likely to behave differently from a household that
-is hit once.
-Both of these factors make it difficult to draw strong 
-predictions on what the effects should be (except that the
-initial impact should lead to postponement of fertility).
+Similarly, we discuss in the Appendix whether we can 
+distinguish between a postponement of fertility and a
+permanent reduction as a result of crop losses.
+Unfortunately, because of data limitations this is difficult
+to show with any degree of confidence, but we believe that 
+postponement of fertility is more probable than a permanent 
+reduction, which is more likely to be affected by the risk 
+of a shock \citep{Portner2014}.
+
+% The Appendix also discuss how including additional lags
+% of crop loss affects the results.
+% In principle, if shocks lead to a postponement of fertility we 
+% should see longer lags be associated with higher fertility, 
+% whereas if shocks lead to a permanent reduction in fertility we
+% should not see any compensatory effect.
+% There are, however, a number of issues with this approach.
+% First, it is possible that the effect of a crop loss is felt
+% in subsequent period and not only in the period in which it
+% was observed.
+% Second, longer spacing between births can be costly because economies
+% of scale are harder to realize and because the mother may have to 
+% be out of the labor force for longer if she already have young
+% children to care for.
+% This implies that a household that are hit twice in a row by
+% a crop loss is likely to behave differently from a household that
+% is hit once.
+% Both of these factors make it difficult to draw strong 
+% predictions on what the effects should be (except that the
+% initial impact should lead to postponement of fertility).
 
@@
 % [Alternative lead]
-What implications do our results have for our understanding of 
-long-term fertility?
+What implications, if any, do our results then have for our understanding 
+of long-term fertility?
 % What we do not know is whether the psychic cost of using traditional 
@@
 
-In this section we discuss how including additional lags
-of crop loss affects the results.
-In principle, if shocks lead to a postponement of fertility we 
-should see longer lags be associated with higher fertility, 
+When discussing our results in the paper, we treat the 
+lower likelihoods of pregnancy and birth as a postponement 
+of fertility.
+In this section, to try to disentangle whether the lower 
+likelihoods reflect a postponement or a more permanent reduction 
+in fertility, we discuss how including additional lags
+of crop loss affects the results.%
+\footnote{
+We are grateful to an anonymous reviewer for suggesting 
+this approach.
+}
+
+In principle, if shocks lead to a postponement of fertility,
+crop losses that occurred further in the past should be
+associated with an \emph{increased} likelihood of current pregnancy 
+or birth (an initial reduction in fertility should be followed
+by a compensatory increase later on), 
 whereas if shocks lead to a permanent reduction in fertility we
 should not see any compensatory effect.
-There are, however, a number of issues with this approach.
-First, it is possible that the effect of a crop loss is felt
-in subsequent periods and not only in the period in which it
-was observed.
-Second, we lose 1/4 of observations each time we include an 
-additional lag, which leaves us with less variation because
-the first round had a higher proportion of crop losses than
-the subsequent rounds.
-This is especially a concern for contraception use because
-of the higher likelihood of measurement error discussed in the
-paper.
-
-Furthermore, even if shocks only had an impact in one period, 
-there are two competing factors at work here.
-First, a negative shocks will make people more likely to postpone 
-their pregnancy/birth because of the factors discussed in the
-paper.
-Second, longer spacing between births can be costly because economies
-of scale are harder to realize and because the mother may have to 
-be out of the labor force for longer if she already have young
-children to care for.
-This implies that a household that are hit twice in a row by
-a crop loss is likely to behave differently from a household that
-is hit once.
-Both of these factors make it difficult to draw strong 
-predictions on what the effects should be (except that the
-initial impact should lead to postponement of fertility).
-We have therefore also added an interaction between crop loss for 
-different periods with the idea that there is a cost to continually 
-postponing births.
+
+Unfortunately, because of data limitation, this approach does not 
+provide a clean test for whether households 
+postpone or reduce their fertility in response to shocks.
+First, given the short duration between surveys and the lack of 
+information about when the shocks occurred, it is possible and
+even likely that the effects of a crop loss last for more than one period.
+Hence, if we find a negative effect of more distant crop losses
+this may be the result of the negative effects of the crop losses
+still affecting the household.
+Second, we lose 1/4 of the original number of observations each 
+time we include an additional lag, which leaves us with less variation 
+in crop loss because the first round had a higher proportion of crop 
+losses than the subsequent rounds.
+
+Even if crop losses' impact are limited to only one period, there 
+are two additional concerns.
+First, the compensatory effect may be spread over multiple 
+periods because of the uncertainty associated with waiting time
+to conception, i.e. a couple may be trying to conceive in the
+period after the crop loss but fail.
+This makes a positive compensatory effect harder to 
+identify.
+Second, it is possible that households may respond differently to a 
+shock if they are hit repeatedly.
+There are two competing factors at work.
+On one hand, repeated shocks may exacerbate the negative impact on the
+household, making it \emph{more} likely that fertility is postponed.
+On the other hand, even longer spacing between births can be costly 
+because economies of scale are harder to realize and because the 
+mother may have to  be out of the labor force for longer if she 
+already have young children to care for, which will make it
+\emph{less} likely that fertility is postponed.
+To capture the impact of repeated shocks we therefore also 
+include the interaction between crop losses.
 
@@
 
+With the data limitations and additional concerns,
+it is difficult to draw strong predictions on what the coefficients
+should be (except that the initial impact should lead to postponement 
+of fertility).
 Tables \ref{tab:postpone_birth} and \ref{tab:postpone_contraceptive}
@@
 between crop loss and the additional lag in crop loss are included.
+
 Including only the additional lag of crop loss it appears that 
@@
 variable, although these are also not statistically significant.
-If we include the interaction between crop losses we instead
-find negative association for both immediate and lagged crop
-loss with a positive interaction coefficient.
-For contraceptive use we see a similar pattern (of course,
-with the signs reversed).
-
-Although these results should be interpreted with caution the
-results appear most in line with the following conclusions.
-First, crop losses tend to affect not only this period but also 
-(weakly) the period after as indicated by the negative coefficient 
-on the lagged crop loss.
-Second, experiencing two crop losses in a row makes it \emph{less}
-likely for people to postpone fertility (although the effect 
-of the crop losses is still negative), most likely because of
-the cost associated with longer spacing.
+These results are broadly consistent with a compensatory response
+where crop loss longer occur leads to higher fertility now.
+However, if we include the interaction between crop losses we 
+instead find negative association for both immediate and lagged 
+crop loss with a positive interaction coefficient.
+Even with the positive interaction coefficient the estimated 
+impact of crop losses in the two prior periods is negative.
+For contraceptive use we see similar patterns (with the signs 
+reversed).
+
+Although these results should be interpreted with caution, they
+are not inconsistent with a compensatory/postponement effect, 
+as long as some crop losses' impact last for more than one
+period after its occurrence and/or the compensatory effect is
+spread over more than one period.
+In either specification the coefficients on lagged crop loss
+are small relative to the immediate impact.
+In sum, we cannot establish with any degree of confidence from
+these results whether the response to shocks lead to only postponement 
+of shocks or a permanent reduction.
+Given the findings from Guatemala that shocks only 
+have a negative impact on completed fertility if they occur
+late in a woman's reproductive life and that completed 
+fertility is mainly driven by the \emph{risk} of shocks,
+it is most likely that what we observe are, indeed, postponement
+rather than a reduction in completed fertility \citep{Portner2014}.
+
+% First, crop losses tend to affect not only this period but also 
+% (weakly) the period after as indicated by the negative coefficient 
+% on the lagged crop loss.
+% Second, experiencing two crop losses in a row makes it \emph{less}
+% likely for people to postpone fertility (although the effect 
+% of the crop losses is still negative), possibly because of
+% the cost associated with longer spacing.
 

### 2017-10-10 response/response-jde-1.tex
Respond to R1 on marriage paper
Respond to R1 on cut-off

@@
 First, there are already two papers that examine this question, 
-albeit for other countries \citet{mbiti08,Hoogeveen2011}.
+albeit for other countries \citep{mbiti08,Hoogeveen2011}.
 Second, as those two papers show, the length of our paper would have
@@
 
-In many ways, the log specification in Table A- TK is more theoretically 
+In many ways, the log specification in Appendix Table A-10 is more theoretically 
 appealing because it captures that we should expect a given increase
@@
 Crop losses lead to statistically significant reductions in the likelihoods 
-of pregnancy and birth with crop losses, and statistically significant 
-increases in the likelihood of using contraceptives overall and
-for the use of traditional contraceptives.
+of pregnancy and birth, and statistically significant 
+increases in the likelihood of using any contraceptives and using 
+traditional contraceptives.
 The differences from the dummy results are that the pregnancy and
@@
 To show that the specific cut-off chosen does not have substantial effects on
-the conclusions drawn we have added Table A-TK, which shows the main results
+the conclusions drawn, we have added Appendix Table A-9, which shows the main results
 for different cut-offs together with the associated percentages of observations
 that would be counted as having a crop loss.
-Except for any crop loss (cut-off of above zero) the results
+Except for any crop loss (cut-off at zero TZS) the results
 are consistent across all other specifications.
@@
 As we increase the cut-off there are more and more households 
-that have experienced a crop loss but are not counted as having a
+that have experienced a crop loss, but are not counted as having a
 crop loss, which will tend to bias results towards zero if there

### 2017-10-10 response/response-jde-1.tex
Respond to R1 on causality

@@
 We have tried to highlight the factors that can affect our identification 
-strategy, but if we need to tone down the causality claims further
-we are happy to do so.
+strategy and reduce the use of causal language, but if we need to tone down 
+the causality claims further we can do so.
 We list below our responses to the individual comments and suggestions.
@@
 
-[Matches with ADF's question on rainfall and with R2's first question 
-on what drives crop losses]
+We have reduce the use of causal language throughout the paper and
+added this as an additional caveat to our results in the Conclusion.
+Below we discuss factors that may explain the distribution of
+crop loss.
 

### 2017-10-10 paper/incomeShocks-jde-r1.tex
Respond to R1 on causality

@@
 a crop shock.
-Furthermore, women significantly increase their contraception use in response 
-to crop losses. 
+Furthermore, women have significantly higher contraception use after 
+crop losses. 
 The increase in contraceptive use comes almost entirely from traditional 
@@
 fertility and contraceptive use.
-We show that income shocks decrease the likelihood of pregnancies and births.
-At the same time, shocks significantly increase use of contraceptives.
+We show that income shocks are associated with a decrease the likelihood of 
+pregnancies and births.
+At the same time, use of contraceptives significantly increase after a shock.
 This increase mainly occurs through the use of traditional contraceptives, 
@@
 We argue that the postponement of fertility in households experiencing income 
-shocks is the result of the increased contraception use rather than a biological 
-response or separation of spouses.
+shocks is likely the result of the increased contraception use rather than a 
+biological response or separation of spouses.
 
@@
 Our preferred estimates---woman level fixed effects---show that experiencing 
-crop loss decreases the likelihood of being pregnant by 10 percentage points 
-and the likelihood of a birth by 17 percentage points.
-These effects are substantial; on average across the four survey rounds, approximately 
+crop loss is associated with a 10 percentage points decrease in the likelihood 
+of being pregnant and a 17 percentage points reduction in the likelihood of a birth.
+These estimates are substantial; on average across the four survey rounds, approximately 
 17\% of women were pregnant and 17\% had given birth in the preceding period. 
@@
 is so strong when there is little use of modern contraceptives.
-In Table \ref{tab:effectiveness}, we examine to what extent shocks cause
-a delay in pregnancy through the increased use of contraceptives by
-estimating the following: 
+In Table \ref{tab:effectiveness}, we examine to what extent shocks are
+associated with a delay in pregnancy through the increased use of 
+contraceptives by estimating the following: 
 the effects of crop loss between the prior survey and the
@@
 characteristics and both shocks and fertility decisions, we employ a fixed effect model.
-Our estimates show that both the likelihood of pregnancy and birth decrease in
-response to a crop loss and that contraceptive use, especially use of traditional 
+Our estimates show that both the likelihood of pregnancy and birth decrease 
+after a crop loss and that contraceptive use, especially use of traditional 
 methods, such as abstinence and rhythm, significantly increases.
 
-We examine a number of causes that may explain the reduction 
+We examine a number of potential explanations for the reduction 
 in fertility following a crop loss.
@@
 
-Our main conclusion is that households consciously delay 
-fertility through contraceptive use to cope with income shocks. 
+Although we cannot entirely rule out that other, unobserved
+factors play a role, our main conclusion is that households appear
+to consciously delay fertility through contraceptive use to cope 
+with income shocks. 
 This is especially of interest since this is accomplished 
@@
 
-There are two potential caveats to our results.
+There are three potential caveats to our results.
 First, the sample size is small and the survey rounds
@@
 }
+Third, although we cover a number of possible explanations
+for what determines crop loss, it is possible that there may 
+be other unobserved, time-varying household characteristics 
+that influence whether a household experience a crop loss.
+If that is the case the causality from crop to fertility decisions
+may not hold.
 

### 2017-10-10 response/response-jde-1.tex
Respond to R1 on time-varying variables

@@
 Other time-varying variables are illness and health variables, which 
-we already show that these are unlikely to explain the results we find.
+we already show are unlikely to explain the results we find.
 
@@
 
-In Table TK we show the main results when including a dummy for livestock
+In Appendix Table A-14 we show the main results when including a dummy for livestock
 loss and the price index included with the Kagera data.
@@
 crop loss question in that the livestock question also includes 
-any livestock given away by the household and that is not indication which
-proportion falls into actual loss and which proportion is gifts.
+any livestock given away by the household without specifying how much 
+is actual loss and how much is gifts.
 Second, the prices used for the price index are measured at the time
 of the survey.
-We tried lagging the price index, but it does not change the findings.
+We tried lagging the price index, but that does not change the findings.
 

### 2017-10-10 response/response-jde-1.tex
Respond to R1 on coefficients
Respond to R1 on contraceptive use interacted with crop loss

@@
 the coefficients.
-See, for example, Table A-8 on the effects of log crop loss.
+See, for example, Table A-10 on the effects of log crop loss.
 
@@
 difficult to measure correctly.
-[something on the panel nature of the data and that that is important
-for getting at the incentives and the use over time; not perfect but
-at least is for a shorter period than other (cross-sectional) surveys]
+% [something on the panel nature of the data and that that is important
+% for getting at the incentives and the use over time; not perfect but
+% at least is for a shorter period than other (cross-sectional) surveys]
 The table is important because it provides us a way to get at this
@@
 
-[also speaks to whether this is a conscious decision or not] 
+% [also speaks to whether this is a conscious decision or not] 
 The table also supports our conjecture that the estimated effects 
-[might need a less causal word] 
+% [might need a less causal word] 
 of crop loss capture a conscious decision. 

### 2017-10-10 response/response-jde-1.tex
Minor edits

@@
 suggestions,
-which we believe has improved the paper.
+which we believe have improved the paper.
 
@@
 \item A number of new Appendix tables have been added in response
-to various comments and other edited.
+to various comments.
 \end{enumerate}
@@
 These results are broadly consistent with a compensatory response
-where crop loss longer occur leads to higher fertility now.
+where crop loss longer ago leads to higher fertility now.
 However, if we include the interaction between crop losses we 
@@
 for pregnancy and birth, 
-The contraceptive use results have the correct sign but the direct
+The contraceptive use results have the correct sign, but the direct
 effect is not statistically significant, presumably because of the 

### 2017-10-10 response/response-jde-1.tex
Respond to R2 on omitted variables

@@
 
-[see response to R1's q7 on time-varying covariates]
-
-[TK we need to tone down the causality part; explain the growing season and
-the crop loss question]
+% [see response to R1's q7 on time-varying covariates]
+% 
+% [TK we need to tone down the causality part; explain the growing season and
+% the crop loss question]
 
@@
 it is not possible to verify what factors played a role in the crop loss.
+We have therefore reduced the use of causal language as detailed in our
+response to Reviewer 1 above.
+
 As you suggest, weather is a potential factor.
@@
 of period lengths and we therefore focus on the growing season approach.
-Table TK shows individual level fixed effects estimates of the 
+Appendix Table A-15 shows individual level fixed effects estimates of the 
 effects of average monthly rainfall in centimeters on the likelihood

### 2017-10-10 response/response-jde-1.tex
Respond to R2 on spillovers

@@
 
-We have added an additional table to the Appendix to address this
+We have added Appendix Table A-13 to address this
 concern and discuss the results in the Robustness section. 
@@
 
-We already include a version of this in the paper.
-Table A-17 TK include the fraction of households in the community 
+We already include a version of this in the paper, but the 
+results were not well explained.
+Appendix Table A-16 include the fraction of households in the community 
 who have experienced a crop loss (excluding the household itself).
@@
 particular type of shock.
+
+Finally, as detailed in our response to Reviewer 1 on potentially 
+time-varying covariates we specifically discuss that prices do not appear 
+to have any association with fertility decisions and including 
+a price index does not the change the main results.
 

### 2017-10-10 paper/incomeShocks-jde-r1.tex
Minor edits
Minor edits to shorten paper

@@
 % incomeShocks-jde-r1.tex
-% Begun.: 2017-07-13
-% Edited: 2017-07-13
-
-% Revisions based on Demography referee reports
+
+% Revisions based on Journal of Development Economics referee reports
 
@@
 
-\date{September 2017}
+\date{October 2017}
 
@@
 on timing of fertility \citep[see, for example,][]{Barreca2015}.
-%  and TK Josh paper.
 In a different vein, a monthlong blackout in Zanzibar increased 
@@
 focusing on the number of children born \citep{Portner2018}.
-% [Little evidence in general on timing of fertility in developing countries;
-% most of the research has focused on completed fertility / children-ever-born] 
+% Little evidence in general on timing of fertility in developing countries;
+% most of the research has focused on completed fertility / children-ever-born
 % Furthermore, standard models of fertility focus mainly on the effects of time
@@
 
-Observing a decline in pregnancies and birth rates following a shock is, however, 
+A decline in pregnancies and birth rates following a shock is, however, 
 not evidence of a conscious decision to change the timing of fertility.
-A shock may, for example, lead one or more members of the household to 
+A shock may, for example, lead one or more household members to 
 migrate in search of better economic opportunities, and if either the husband 
@@
 
-Most of the prior literature on the impact of economic and other shocks on fertility 
+Most of the literature on the impact of economic and other shocks on fertility 
 has used historic data on what are now developed countries.
@@
 showed no change \citep{bengtsson06}.
-Using present-day data, there is evidence that famine and domestic/regional 
-military attacks in Ethiopia led to short-term decreases in the likelihood of 
-conception \citep{lindstrom99}.
+Using present-day data, famine and domestic/regional military attacks in Ethiopia 
+led to short-term decreases in the likelihood of conception \citep{lindstrom99}.
 Closer to our paper, positive rainfall shocks in Senegal increased 
@@
 the possible correlation between the probability of experiencing 
-an adverse event and unobserved individual and household characteristics may 
-bias the estimates, and the cross-sectional data allows for neither household 
+an adverse event and unobserved characteristics may bias the estimates, 
+and the cross-sectional data allow for neither household 
 nor individual level fixed effects estimates.
@@
 
-On average across the four surveys, 17\% of women report being pregnant 
-at the time of the interviews and 16\% of women report giving birth since the 
-prior survey round.
+On average, 17\% of women report being pregnant at the time of the interviews 
+and 16\% of women report giving birth since the prior survey round.
 In total, just below 70\% of women in the sample report giving birth in
@@
 
-% [Descriptive statistics and discussion here]
+% Descriptive statistics and discussion here
 \input{../tables/desstat1.tex}
@@
 primary education, and 44\% completed at least primary education.
-The high reported education level is most likely the result of the 1974 
+The high education level is most likely the result of the 1974 
 Universal Primary Education Movement,
@@
 }
-Hence, it is unclear to what extent reported education levels reflect 
+Hence, it is unclear whether reported education levels reflect 
 women's actual human capital.
@@
 
-% [BASELINE estimations]
+% BASELINE estimations
 
@@
 Above 1,000 TZS the effects on pregnancy and birth become smaller and 
-eventually statistically significant insignificant.
+eventually statistically insignificant.
 This is not surprising.
@@
 and the coefficient on crop loss does not change.
-[TK bla, bla - next sentences need a rewrite]
 Although this does not completely rule out the possibility
@@
 come about.
-We present suggestive evidence that show that only those households
+We present suggestive evidence that shows that only those households
 with a strong incentive, i.e. those exposed to crop loss, are 
@@
 that influence whether a household experience a crop loss.
-If that is the case the causality from crop to fertility decisions
-may not hold.
+If that is the case the causality from crop loss to fertility 
+decisions may not hold.
 
@@
 
-% [Lead with this or the next sentence?]
-% Our result that people are able to postpone fertility using mainly 
-% traditional contraceptive do, however, have implications for our 
-% understanding of long-term fertility.
-% [Alternative lead]
-What implications, if any, do our results then have for our understanding 
+What implications do our results have for our understanding 
 of long-term fertility?
-% What we do not know is whether the psychic cost of using traditional 
-% contraceptives is so high that although households are willing to use 
-% them to deal with short-term shocks, they become difficult to use for 
-% long-run moderation of fertility.
 The economic literature on fertility transition in Europe and the US 

### 2017-10-10 paper/incomeShocks-jde-r1.tex
Add highlights

@@
 \newpage
+\setcounter{page}{0}
+\thispagestyle{empty}
+
+\section*{Highlights}
+
+\begin{enumerate}
+\item In Tanzania, likelihood of pregnancy and birth is lower after an income shock.
+\item Contraception use, especially traditional, are higher after an income shock.
+\item Postponement of fertility in response to shock appears to be conscious decision.
+\item Traditional contraceptives have high efficacy after income shock. 
+\end{enumerate}
+
+\newpage
 

### 2017-10-10 response/response-jde-1.tex
Minor edits of comments to ourself
Minor edits

@@
 
-Although we cannot measure directly measure stress, we do have information 
+Although we cannot directly measure stress, we do have information 
 on illnesses and their durations.
@@
 are consistent across all other specifications.
-As the cut-off becomes very large statistically significance is lower
+As the cut-off becomes very large significance is lower
 for pregnancy and births and eventually also for contraceptive use.
@@
 
-We have reduce the use of causal language throughout the paper and
+We have reduced the use of causal language throughout the paper and
 added this as an additional caveat to our results in the Conclusion.
@@
 % 
-% [TK we need to tone down the causality part; explain the growing season and
+% [we need to tone down the causality part; explain the growing season and
 % the crop loss question]
@@
 We use two specification, one with the last prior growing season and 
-one with prior growing for waves 2 through 4 and the last two prior 
-growing seasons for wave 1.
+one that use the prior growing for waves 2 through 4 and the last two prior 
+growing seasons for wave 1 (given that wave 1 has a recall period 
+of 12 months).
 Higher rainfall is associated with a lower likelihood of a crop loss. 
